### 2021-02-18 Notes.Rmd
back up data exploration

---
title: "Notes"
output: html_notebook
author: Juan Rocha
---

To-do list:

1. Compute the import / export ratio per country
2. Same at commodity level: are there commodities that are purely trade driven?
3. Explore Paula's file: rank countries per number of companies on high risk
4. Extract main commodities
5. Characterize missing values: where are they missing (countries) and what vars are not very useful?
6. Construct networks between countries and corporations weighted by subsidiaries?

### 2021-02-18 01-data_import.R

@@ -1,0 +1,40 @@
+library(tidyverse)
+
+df1 <- readxl::read_xlsx(
+    path = "data/Drivers_Increasing_Risk_ED-2.xlsx",
+    sheet = 1) %>% 
+    janitor::clean_names()
+
+df2 <- readxl::read_xlsx(
+    path = "data/Drivers_Increasing_Risk_ED-2.xlsx",
+    sheet = 2,
+    na = "n.d") %>% 
+    janitor::clean_names()
+
+df1 %>% skimr::skim()
+
+df1 %>%
+    ggplot(aes(country)) + 
+    geom_bar() +
+    coord_flip()
+    
+df1 %>%
+    filter(country == "USA") %>% 
+    pull(driver_of_change) %>% 
+    unique()
+
+df2 %>% # how far is it coded
+    filter(!is.na(access_date)) %>% 
+    skimr::skim()
+
+## Fix missing values
+df2 <- df2 %>% 
+    mutate(
+        across(where(is.character), ~str_remove(., pattern = "n.d."))) %>% 
+    mutate(
+        percent_certificated_areas = as.numeric(percent_certificated_areas) ,
+        area_ha, no_workers =  as.numeric(area_ha),
+        turnover_mll_usd = as.numeric(turnover_mll_usd))
+
+df2 
+

### 2021-03-10 FAO_stats.Rmd
updates on data exploration with latest file from Paula

---
title: "Exploring FAO stats"
author: "Juan Rocha"
date: "2/21/2021"
output: 
  html_notebook:
    toc: false
    code_folding: hide
  html_document:
    toc: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This notebook is based on the script draft `99-data_exploration.R`. The central question of this preliminary analysis is: **how do we prioritize areas of the world and which commodities to further focus our analysis?**

```{r}
library(tidyverse)
library(plotly)
```

## Import data

```{r}
## FAO data on crops
file <- "~/Documents/Projects/DATA/FAOSTAT/CommodityBalances_Crops_E_All_Data_(Normalized).csv"

dat <- read_csv(file) %>% 
  janitor::clean_names()
```

Some summary stats:
```{r}
## descriptive 
skimr::skim(data = dat)
```

FAOStats has production data for `r dat %>% pull(area) %>% unique() %>% length()` areas, of which 182 are countries. It keeps track of `r dat %>% pull(item) %>% unique() %>% length()` agricultural commodities (excluding animals). For every country, year and commodity, it keeps record of: _`r dat %>% pull(element) %>% unique() `_. This dataset has three flags (you can read more [here](http://www.fao.org/faostat/en/#definitions)):

- S: stanrdardized data
- SA: Aggregate, may include official, semi-official, estimated or calculated data
- SD: statistical discrepancy

The commodities available are:

```{r}
dat %>% select(item) %>% unique()
```

In order to prioritize areas of the world and commodities to focus our analysis, we need to know:

1. If the emergence of zoonotic diseases is driven by deforestation and land cover change (Allen 2017), then we need to figure out where has those rates being higher. The EU Copernicus Land use data can be of use (30m resolution, 10+ years), but one needs to be careful on how it is calculated -- e.g. % versus area; and it's computationally expensive.
2. Alternatively, without land cover data, one can use FAOStats to investigate which commodities have increased production faster, and what proportion of that production is dedicated to the international market. As such, one can identify places where international coorportations might have played an important role in that agricultural expantion (in production, not area).
3. The ratio of exports / production can inform us how much export oriented has a country become (per commodity). Note that production can increase because there is more population to feed, so one probably calculate per capita values to account for increase of internal demand. Then we need

### Example

Let's look at Colombia's coffee production:

```{r}
dat %>% 
  select(-element_code, -year_code, -flag) %>% 
  filter(area == "Colombia") %>% 
  filter(item == "Coffee and products") %>% 
  pivot_wider(names_from = element, values_from = value) %>%
  mutate(
    total = `Domestic supply quantity` + `Export Quantity`,
    total2 = total - `Stock Variation` - `Import Quantity`,
    differ = total2 - Production
  ) %>%
  pivot_longer(cols = Production:differ, names_to = "element", values_to = "value") %>%
  # filter(element != "total") %>%
  #group_by(item, year) %>% 
  ggplot(aes(year, value)) +
  geom_line(aes(group = element, color = element)) +
  scale_y_continuous(labels = scales::label_scientific()) +
  facet_wrap(~element) +
  theme_light(base_size = 8)
```

The variables that start on capital letters are the original from the data, the ones with small caps are variables I re-calculated to understand what they mean. Roughly speaking, `Production` is equal to `Domestic supply quantity` + `Export Quantity` - `Stock Variation` - `Import Quantity`. Thus, there are a few options for the ratio that indicates whether a country produces mainly for the international market or for domestic consumption:

- **Exports / Production**: Would be the % of production that is exported. The problem is if production is zero (errors in the data). If no errors, it should always be a number between 0-1.
- **Exports / Domestic**: This is a true ratio, if > 1 it means the country is export oriented, if < 1 domestic oriented. The amount of what is exported does not matter, only relative to what is domestically consumed.

Note that both production and domestic can be increasing simply because population and hence domestic demand is growing. But if that's true, it is already accounted for in the denominator. Dividing by population in the numerator and denominator would render the same ratio, so it's unnecessary.

Let's calculated in the full dataset:

```{r}
dat %>% 
  ggplot(aes(value)) +
  geom_density() +
  scale_x_log10() +
  facet_wrap(~element, scales = "free") +
  theme_light(base_size = 8)

```
Notice that there will be some weird cases. Just by looking at the distributions (note on log scale!), there is the possibility of division by zero when a country exports but does not produce (e.g. all products comes from inports, some processing is done and then re-exported again), or when a country does not consume domestically.

```{r}
# get list of countries
countries <- dat %>% pull(area) %>% unique()

ratios <- dat %>% 
  select(-element_code, -year_code, -item_code, -unit) %>% 
  filter(area %in% countries[1:182]) %>% 
  pivot_wider(names_from = element, values_from = value) %>% 
  # avoid divisions by zero:
  filter(
    !is.na(`Export Quantity`),  
    Production != 0,
    !is.na(`Domestic supply quantity`),
    `Domestic supply quantity` != 0
  ) %>%  # reduces the dataset from 1M obs to 0.5
  group_by(year, area, item) %>% 
  mutate(
    export_ratio = `Export Quantity` / `Domestic supply quantity`,
    export_prop = `Export Quantity` / Production
  ) 
```

```{r}
ratios_summary <- ratios %>% 
  ungroup() %>% 
  group_by(area,item) %>% 
  summarize(
    export_ratio = median(export_ratio, na.rm = TRUE),
    export_prop = median(export_prop, na.rm = TRUE)
  )

ratios_summary %>% arrange(desc(export_ratio))
```
```{r}
ratios_summary %>% arrange(desc(export_prop))
```
```{r fig.width=6, fig.height=7}
keyvar <- sym("export_prop")

g1 <- ratios_summary %>% ungroup() %>% 
  filter(area %in% countries[1:182]) %>% 
  filter(eval(keyvar) < 10 , eval(keyvar) > -0.1) %>% 
  # group_by(area) %>% 
  mutate(area = as_factor(area),
         item = as_factor(item)) %>%
  mutate(area = fct_reorder(area, .x = eval(keyvar), .fun = median, .desc = TRUE),
         item = fct_reorder(item, .x = eval(keyvar), .fun = median, .desc = TRUE)) %>% 
  ggplot(aes(item, area)) +
  geom_tile(aes(fill = eval(keyvar))) +
  scale_fill_gradient2("Stat", mid = "gray85", midpoint = 0) +
  theme_light(base_size = 8) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "top")

g1
```

### 2021-03-10 companies.Rmd
updates on data exploration with latest file from Paula

---
title: "Exploring companies data"
author: "Juan Rocha"
date: "21-03-10"
output: 
  html_notebook:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(plotly)
library(patchwork)

theme_set(theme_light())
```

Import data:

```{r message=FALSE, warning=FALSE, error=FALSE}
df1 <- readxl::read_xlsx(
    path = "data/Drivers_Increasing_Risk_ED-2.xlsx",
    sheet = 1,
    na = "NA") %>% 
    janitor::clean_names()

df2 <- readxl::read_xlsx(
    path = "data/Drivers_Increasing_Risk_ED-2.xlsx",
    sheet = 2, 
    na = "NA") %>% 
    janitor::clean_names()
```

Currently the dataset has `r df1 %>% nrow()` companies, in `r df1$country %>% unique() %>% length()` countires, dealing with `r df1$commodity %>% unique() %>% length()` commodities (including `NA`).

The top countries are:
```{r}
g1 <- df1 %>%
    group_by(country) %>% 
    summarize(count = n()) %>%
    arrange(desc(count)) %>% 
    ggplot(aes(count)) +
    geom_density() +
    geom_vline(xintercept = 20, color = "orange") +
    labs(title = "Companies per country") +
    theme_light(base_size = 8)

g2 <- df1 %>%
    group_by(country) %>% 
    summarize(count = n()) %>%
    arrange(desc(count)) %>% 
    filter(count > 20) %>% 
    mutate(country = as_factor(country) %>% fct_reorder(count)) %>% 
    ggplot(aes(country, count)) +
    geom_col() +
    coord_flip() +
    labs(title = "Country")+
    theme_light(base_size = 8)

g3 <- df1 %>%
    group_by(origin) %>% 
    summarize(count = n()) %>%
    arrange(desc(count)) %>% 
    filter(count > 20) %>% 
    mutate(origin = as_factor(origin) %>% fct_reorder(count)) %>% 
    ggplot(aes(origin, count)) +
    geom_col() +
    coord_flip() +
    labs(title = "Origin")+
    theme_light(base_size = 8)

g1 + g2 + g3
```

Top companies are:

```{r}
g1 <- df1 %>%
    group_by(company) %>% 
    summarize(count = n()) %>%
    arrange(desc(count)) %>% 
    ggplot(aes(count)) +
    geom_density() +
    geom_vline(xintercept = 10, color = "orange")

g2 <- df1 %>%
    group_by(company) %>% 
    summarize(count = n()) %>%
    arrange(desc(count)) %>% 
    filter(count > 10) %>% 
    mutate(company = as_factor(company) %>% fct_reorder(count)) %>% 
    ggplot(aes(company, count)) +
    geom_col() +
    coord_flip() +
    theme(axis.text.y = element_text(size = 6))

g2 + inset_element(g1, left = 0.25, bottom = 0.05, right = 0.95, top = 0.65)

```

And the most importat commodities:

```{r}
g1 <- df1 %>%
    group_by(commodity) %>% 
    summarize(count = n()) %>%
    arrange(desc(count)) %>% 
    ggplot(aes(count)) +
    geom_density() +
    geom_vline(xintercept = 30, color = "orange")

g2 <- df1 %>%
    group_by(commodity) %>% 
    summarize(count = n()) %>%
    arrange(desc(count)) %>% 
    filter(count > 0) %>% 
    mutate(commodity = as_factor(commodity) %>% fct_reorder(count)) %>% 
    ggplot(aes(commodity, count)) +
    geom_col() +
    geom_hline(yintercept = 30, color = "orange") +
    coord_flip() +
    theme(axis.text.y = element_text(size = 6))

g2 + inset_element(g1, left = 0.25, bottom = 0.05, right = 0.95, top = 0.65)
```

Note that for commodities going for the big numbers might not be wise. Many zoonotic diseases are related to animal exploitation (pork, pigs, lamb) which might not be as big when compare to other commodities, but are more prone to diseases that can jump to humans.

```{r}
df1 %>%
    group_by(commodity) %>% 
    summarize(count = n()) %>%
    arrange(desc(count)) %>% 
    filter(count > 0) %>% 
    mutate(commodity = as_factor(commodity) %>% fct_reorder(count)) %>% 
    ggplot(aes(commodity, count)) +
    geom_col() +
    geom_hline(yintercept = 30, color = "orange") +
    coord_flip() + 
    facet_wrap()
    theme(axis.text.y = element_text(size = 6))

```

## Networks

Here are some ideas for further analysis:

1. Bipartite network of countries and companies
    - One mode network of countries connected by number of companies
    - One mode network of companies who operate on the same countries
    - Can one figure out cooperation / competition for resources?
2. For a selection of countries we could zoom into subnational networks (e.g. USA)
3. Shareholder - Conglomerate - Company - Subsidiaries | Which layers are relevant for our question? Bear in mind the missing values:

```{r}
df1 %>% 
    select(1:4) %>% 
    skimr::skim()
```

```{r}
df1 %>% 
    select(company, origin, country) %>% 
    group_by(origin, country) %>% 
    unique() %>% 
    summarize(companies = n()) %>% 
    ggplot(aes(origin, country)) +
    geom_tile(aes(fill = companies)) +
    scale_fill_viridis_c() +
    theme_light(base_size = 6) +
    theme(axis.text.x = element_text(angle = 90))

```
Hard to read, maybe better on table form:

```{r}
df1 %>% 
    select(company, origin, country) %>% 
    group_by(origin, country) %>% unique() %>% 
    summarize(companies = n()) %>% 
    arrange(desc(companies))
```

### Errors

- `DriversED` rows 102 onwards have shifted column order for columns L, M. Needs to be corrected to be able to do network analysis
- There are typos on countries names e.g. Brazil and Brasil

```{r}
df1 %>% 
    filter(company == "NESTLE S.A.")
```

### 2021-03-10 01-data_import.R

@@ -1,5 +1,9 @@
 library(tidyverse)
 
+url <- "https://stockholmuniversity.app.box.com/file/777762238977?s=8u6vl01c1efiwn7s4hv688o44e53nwdm"
+
+""
+
 df1 <- readxl::read_xlsx(
     path = "data/Drivers_Increasing_Risk_ED-2.xlsx",
     sheet = 1) %>% 
@@ -11,22 +15,27 @@ df2 <- readxl::read_xlsx(
     na = "n.d") %>% 
     janitor::clean_names()
 
+
+
+
 df1 %>% skimr::skim()
 
-df1 %>%
+df2 %>%
     ggplot(aes(country)) + 
     geom_bar() +
     coord_flip()
     
-df1 %>%
+df2 %>%
     filter(country == "USA") %>% 
     pull(driver_of_change) %>% 
     unique()
 
-df2 %>% # how far is it coded
+df1 %>% # how far is it coded
     filter(!is.na(access_date)) %>% 
     skimr::skim()
 
+
+df1 %>% skimr::skim()
 ## Fix missing values
 df2 <- df2 %>% 
     mutate(

### 2021-03-10 99-data_exploration.R

@@ -1,16 +1,135 @@
 library(tidyverse)
 library(plotly)
+library(tictoc)
 
-## FAO data on crops
+#### FAO data on crops ####
 file <- "~/Documents/Projects/DATA/FAOSTAT/CommodityBalances_Crops_E_All_Data_(Normalized).csv"
 
-dat <- read_csv(file)
+dat <- read_csv(file) %>% 
+  janitor::clean_names()
+
+
+## descriptive 
+tic()
+skimr::skim(data = dat)
+toc() #28 s
+
+
+
+countries <- dat %>% pull(area) %>% unique() # 216 areas of which 182 are countries
+dat %>% pull(item) %>% unique() # 112 foods and other ag products (no animals)
+dat %>% pull(item_code) %>% unique() # are the codes standard across UN?
+dat %>% pull(element) %>% unique() # interesting, it has production (total?), import and exports, domestic supply, losses, food supply...
+dat %>% pull(flag) %>% unique() # S, A, SD?
+# S: standardized data
+# A: Aggregate, may include official, semi-official, estimated or calculated data 
+# SD: statistical discrepancy
+# 
+
+
+dat %>% 
+  select(-element_code, -year_code, -flag) %>% 
+  filter(area == "Colombia") %>% 
+  filter(item == "Coffee and products") %>% 
+  pivot_wider(names_from = element, values_from = value) %>%
+  mutate(
+    total = `Domestic supply quantity` + `Export Quantity`
+  ) %>%
+  pivot_longer(cols = Production:total, names_to = "element", values_to = "value") %>%
+  # filter(element != "total") %>%
+  #group_by(item, year) %>% 
+  ggplot(aes(year, value)) +
+  geom_line(aes(group = element, color = element)) +
+  scale_y_continuous(labels = scales::label_scientific())
+
+## calcualte ratios
+tic()
+ratios <- dat %>% 
+  select(-element_code, -year_code, -item_code, -unit) %>% 
+  filter(area %in% countries[1:182])
+  pivot_wider(names_from = element, values_from = value) %>% 
+  # avoid divisions by zero:
+  filter(
+    !is.na(`Export Quantity`),  
+    Production != 0,
+    !is.na(`Domestic supply quantity`),
+    `Domestic supply quantity` != 0
+  ) %>%  # reduces the dataset from 1M obs to 0.5
+  group_by(year, area, item) %>% 
+  mutate(
[…]

### 2021-06-10 companies.Rmd
visualizing latest data cleaned by Paula

@@
 ```
+
+## Ideas to explore:
+
+- Hotspots of interest within countries
+- Commodities within the hotspots driven by land use (high priority cases)
+- Explore Allen data and see if one can make linkages to specific diseases
+- One can use risk areas to link companies -> hotspots -> disease risk
+- Once a subset of companies is identified, go up to shareholders: which ones are more exposed?

### 2021-06-10 networks.Rmd
visualizing latest data cleaned by Paula

---
title: "Networks of financial actors"
author: "Juan Rocha"
date: "6/10/2021"
output: 
  html_notebook:
      theme:
          code_font:
            google: Fira Code
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE, message=FALSE, warning = FALSE, error = FALSE,
    fig.width = 6, fig.height = 5, tidy = TRUE, dpi  = 300)
```

```{r}
library(tidyverse)
library(plotly)
library(patchwork)
library(network)

theme_set(theme_light(base_size = 10))
```

Read data:

```{r message=FALSE, warning=FALSE, error=FALSE}
df1 <- readxl::read_xlsx(
    path = "data/Paulas_files/001-Drivers_Increasing_Risk_ED.xlsx",
    sheet = 2,
    na = "NA") %>% 
    janitor::clean_names()

```

Currently the data set has `r df1 %>% nrow()` companies, in `r df1$country %>% unique() %>% length()` countries, dealing with `r df1$commodity %>% unique() %>% length()` commodities (including `NA`).

The data is very incomplete with respect to country locations

```{r}
df1 %>% 
    select(company, origin, country) %>% 
    skimr::skim()
```

```{r}
library(sna)
net_countries <-  df1 %>% 
    select(company, origin, country) %>% 
    group_by(origin, country) %>% 
    # unique() %>% 
    count(name = "channel") %>% # a channel is acountry * commodities
    filter(!is.na(origin), !is.na(country)) %>% 
    network(directed = TRUE, bipartite = FALSE, matrix.type = "edgelist", ignore.eval=FALSE, loops = FALSE)

net_countries %v% "indegree" <- sna::degree(
    net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE, cmode = "indegree")
net_countries %v% "outdegree" <- sna::degree(
    net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE, cmode = "outdegree")
net_countries %v% "degree" <- sna::degree(net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE)

gplot(
    net_countries,
    vertex.col = alpha("purple", 0.25),
    label = network.vertex.names(net_countries),
    label.cex = 0.5, label.col = "grey50",
    vertex.border = 0, 
    vertex.cex = sqrt(0.2 + net_countries %v% "outdegree"),
    label.pos = 5,
    edge.col = alpha("grey50", 0.1),
    edge.lwd = log1p(net_countries %e% "channel"),
    displayisolates = TRUE, pad = 0.5,
    mode = "mds"
    )

# networkD3::simpleNetwork(
#     df1 %>%
#         select(company, origin, country) %>%
#         group_by(origin, country) %>%
#         # unique() %>%
#         count(name = "channel") %>% 
#         filter(!is.na(origin), !is.na(country)) ,
#     Source = "origin", Target = "country", zoom = TRUE,
#     linkColour = "grey", nodeColour = "blue")

```
Maybe fit a PCA to the data with network stats, commodities and so forth per country, and use the coordinates for visualization. Currently using a multi-dimensional scaling projection. Another way of "seeing" the network is by ranking which countries are the most influencial given their links to other countries (_outdegree_), once self-links are removed (companies whose country of origin and operation is the same):

```{r}
df1 %>% 
    select(company, origin, country) %>% 
    group_by(origin, country) %>% 
    # unique() %>% 
    count(name = "channel") %>% # a channel is acountry * commodities
    filter(!is.na(origin), !is.na(country)) %>% 
    filter(origin != country) %>% 
    arrange(desc(channel))
```

The dataset has `r unique(df1$company) %>% length()` unique companies and `r unique(df1$commodity) %>% length()` commodities. Below a table that summarizes in how many different places the same company extract the same resource:

```{r}
df1 %>% 
    select(company, commodity) %>% 
    group_by(company, commodity) %>% 
    count(name = "geographies") %>% 
    arrange(desc(geographies))
```

```{r}
df1 %>% 
    select(company, commodity) %>% 
    group_by(company, commodity) %>% 
    count(name = "geographies") %>% 
    arrange(desc(geographies)) %>% 
    ggplot(aes(company, commodity)) +
    geom_tile(aes(fill = geographies)) +
    scale_fill_viridis_c() +
    theme_light(base_size = 6) +
    theme(axis.text.x = element_text(angle = 90)) 
```

### 2021-06-10 box.R

@@ -1,0 +1,23 @@
+library(tidyverse)
+library(httr)
+library(boxr)
+
+
+box_auth(client_id = "jemswi6nx6bie6xep36kgrcvibbgpw71", 
+         client_secret = "2AHfw8XXGPU6I37a5KYQ7aF2a10quWXM")
+
+### failed attempts with httr
+myapp <- oauth_app(
+    "box", 
+    key = "6a15rlvm5w3au3l3p09o5o68soo8slhd", 
+    secret = "8khpqPzalF6M0WpvMOg60xYTvGT2lbCS")
+
+box_endpoint <- oauth_endpoint(
+    request = NULL,
+    authorize = "https://account.box.com/api/oauth2/authorize?client_id=214593079&response_type=cod",
+    access = "https://account.box.com/api/oauth2/authorize?client_id=214593079&response_type=cod", 
+    base_url = "https://account.box.com/api/oauth2/authorize"
+)
+
+
+box_token <- oauth2.0_token(box_endpoint, myapp)

### 2021-06-10 disease_spots.R

@@ -1,0 +1,21 @@
+library(tidyverse)
+library(sf)
+# Allen data
+
+dat <- readxl::read_xlsx(
+    path = "~/Documents/Projects/DATA/Allen_2005_DataMaps_Unprojected/0_Metadata_MOESM4_ESM.xlsx",
+    sheet = 1, skip = 1
+)
+
+path <- "~/Documents/Projects/DATA/Allen_2005_DataMaps_Unprojected/"
+
+files <- paste0(path, dat$filename)
+
+length(files)
+
+df_shp <- read_sf(files[1])
+
+df_shp %>% names()
+df_shp %>% 
+    ggplot() + 
+    geom_sf()

### 2021-09-14 Allen-data.Rmd
exploring Allen zoonotic disease raw data

---
title: "Allen data"
author: "Juan Rocha"
date: "8/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

The purpose of this script is to extract the locations of the map produced in Allen et al, match with a suitable subnational political division and use the polygon to approximate location of companies within countries.

```{r}
# Allen data
# dat <- readxl::read_xlsx(
#     path = "~/Documents/Projects/DATA/Allen_2005_DataMaps_Unprojected/0_Metadata_MOESM4_ESM.xlsx",
#     sheet = 1, skip = 1
# )
```

I did it on a separate script `disease_spots.R` the resulting figure is:

![Fig 1. Allen raw data](figures/Allen_data.png)

The scale at which the observations are coded make it difficult to join with the company data. As the map shows, some diseases were recorded in very specific areas, below the 2nd level admin unit. Others are at the level of states or subnational political units. But most of them are actually country scale. It probably means we need to do the network reconstruction based on country data, and zoom in into second level admin unit for only a few countries (e.g. USA, China, Brazil).

### 2021-09-14 disease_spots.R

@@ -1,10 +1,10 @@
 library(tidyverse)
 library(sf)
-# Allen data
 
+# Allen data
 dat <- readxl::read_xlsx(
-    path = "~/Documents/Projects/DATA/Allen_2005_DataMaps_Unprojected/0_Metadata_MOESM4_ESM.xlsx",
-    sheet = 1, skip = 1
+  path = "~/Documents/Projects/DATA/Allen_2005_DataMaps_Unprojected/0_Metadata_MOESM4_ESM.xlsx",
+  sheet = 1, skip = 1
 )
 
 path <- "~/Documents/Projects/DATA/Allen_2005_DataMaps_Unprojected/"
@@ -13,9 +13,44 @@ files <- paste0(path, dat$filename)
 
 length(files)
 
-df_shp <- read_sf(files[1])
 
-df_shp %>% names()
-df_shp %>% 
-    ggplot() + 
+df_shp <- map(files, read_sf)
+
+rbind(df_shp[[1]], df_shp[[3]]) ## it works
+
+df_all <- map_df(df_shp, rbind)
+
+df_all <- left_join(df_all, dat, by = c("SppName" = "eid_name"))
+
+df_all %>% 
+  ggplot() + 
+  geom_sf(aes(fill = path_disease), alpha = 0.5, size = 0.4) + 
+  xlim(-180,180) +
+  labs(title = "Observations of emergent zoonotic diseases", 
+       subtitle = "149 dieseases recorded from 1970 to 2003",
+       caption = "Data source: Allen et al 2017") +
+  scale_fill_discrete( 
+    guide = guide_legend(
+      title = "Diseases", title.position = "top", title.hjust = 0.5,
+      ncol = 3, keywidth = 0.5, keyheight = 0.5
+    )) +
+  theme_light(base_size = 7) + theme(legend.position = "bottom")
+
+ggsave(
+  filename = "Allen_data.png",
+  plot = last_plot(),
+  device = "png", 
+  path = "figures/",
+  width = 7, height = 4, dpi = 400,
+  bg = "white"
+)
+
+## World maps for comparison:
+## This world map is over ~2Gb, filter only level1 admin unit 
+gam <- read_sf("~/Documents/Projects/DATA/GADM_maps/gadm36_shp/gadm36.shp")
+
+gam %>% 
+  select(starts_with("NAME")) %>% 
+  filter(is.na(NAME_2)) %>% 
+    ggplot() +
     geom_sf()

### 2021-09-23 networks.Rmd
collecting geographical coordinates

@@
 author: "Juan Rocha"
-date: "6/10/2021"
+date: "Last update `r Sys.Date()`"
 output: 
@@
 library(network)
+library(sna)
 
@@
 Currently the data set has `r df1 %>% nrow()` companies, in `r df1$country %>% unique() %>% length()` countries, dealing with `r df1$commodity %>% unique() %>% length()` commodities (including `NA`).
-
-The data is very incomplete with respect to country locations
+Below some overview of the missing values:
 
@@
 ```{r}
-library(sna)
+
 net_countries <-  df1 %>% 
@@
     filter(!is.na(origin), !is.na(country)) %>% 
-    network(directed = TRUE, bipartite = FALSE, matrix.type = "edgelist", ignore.eval=FALSE, loops = FALSE)
+    network(directed = TRUE, bipartite = FALSE, matrix.type = "edgelist", ignore.eval=FALSE, loops = TRUE)
 
@@
     # unique() %>% 
-    count(name = "channel") %>% # a channel is acountry * commodities
+    count(name = "channel") %>% # a channel is a country * commodities
     filter(!is.na(origin), !is.na(country)) %>% 
@@
 
-```{r}
+```{r fig.width = 6, fig.height = 5}
 df1 %>% 
@@
   arrange(desc(geographies)) %>% 
-    ggplot(aes(company, commodity)) +
-    geom_tile(aes(fill = geographies)) +
-    scale_fill_viridis_c() +
-    theme_light(base_size = 6) +
-    theme(axis.text.x = element_text(angle = 90)) 
+  filter(geographies > 3) %>%  ungroup() %>% 
+  mutate(company = as_factor(company)) %>%
+  ## the data needs to be ungrouped for fct_reorder to work!
+  mutate(company = fct_reorder(
+    .f = company, .x = geographies, .fun = sum, .desc = FALSE)) %>% 
+  #pull(company) %>% levels()
+  ggplot(aes(y = company)) +
+  geom_bar(aes(weight = geographies, fill = commodity),
+           position = "stack", show.legend = TRUE, alpha = 0.5) +
+  scale_fill_discrete(
+    "Commodities", guide = guide_legend(
+      ncol = 4, title.position = "top", title.hjust = 0.5,
+      keywidth = 0.5, keyheight = 0.5
+    )
+  ) +
+  labs(title = "Companies ranking by number of countries and commodities",
+       subtitle = "A geography is a combination of `country * commodity` as a proxy of geographical influence of a company.\nOnly companies with >3 geographies are shown", caption = "Data curated by the project, source Orbis database", x = "Geographies", y = "Top companies") +
+  theme_light(base_size = 10) + theme(legend.position = c(0.7, 0.1))
+
 ```
+
+Note there is a relatively low diversity of specialization within companies, meaning that a lot of them are putting their eggs on the same basket.
+
+Another key network representation is by companies given the countries where they together operate, and by the common geographies (meaning exporting the same commodity from the same country).
+
+```{r}
+df1 %>% 
+  select(company, origin, country) %>% 
+  group_by(origin, company) %>% 
+  # Assumption: if country is NA, it means the company operates in its country of origin:
+  mutate(country = case_when(is.na(country) ~ origin, !is.na(country) ~ country)) %>% 
+  # unique() %>% 
+  count(name = "channel") %>% # a channel is acountry * commodities
+  ungroup() %>%
+  pivot_wider(values_from = geographies, names_from = companies, values_fill = 0)
+
+   # filter(!is.na(origin), !is.na(company)) %>% 
+  network(directed = TRUE, bipartite = FALSE, matrix.type = "edgelist", ignore.eval=FALSE, loops = TRUE)
+
+net_companies %v% "indegree" <- sna::degree(
+    net_companies, gmode = "digraph", diag = TRUE, rescale = FALSE, cmode = "indegree")
+net_companies %v% "outdegree" <- sna::degree(
+    net_companies, gmode = "digraph", diag = TRUE, rescale = FALSE, cmode = "outdegree")
+net_companies %v% "degree" <- sna::degree(net_companies, gmode = "digraph", diag = TRUE, rescale = FALSE)
+
+gplot(
+    net_companies,
+    vertex.col = alpha("orange", 0.25),
+    label = network.vertex.names(net_companies),
+    label.cex = 0.5, label.col = "grey50",
+    vertex.border = 0, 
+    vertex.cex = sqrt(0.2 + net_companies %v% "outdegree"),
+    label.pos = 5,
+    edge.col = alpha("grey50", 0.1),
+    edge.lwd = log1p(net_companies %e% "channel"),
+    displayisolates = TRUE, pad = 0.5,
+    mode = "mds"
+    )
+
+```

### 2021-09-23 coords.R

@@ -1,0 +1,59 @@
+# A script to collect coordinates and geographical units that help combine datasets
+# within the project
+# Juan Rocha
+# 20210914
+library(tidyverse)
+library(ggmap)
+
+# register key for the session:
+# register_google(key = "<insert key here>") # available on your google console
+
+# import data:
+df1 <- readxl::read_xlsx(
+    path = "data/Paulas_files/001-Drivers_Increasing_Risk_ED.xlsx",
+    sheet = 2,
+    na = "NA") %>% 
+    janitor::clean_names()
+
+df1 <- df1 %>% 
+    mutate(place = case_when(
+        !is.na(subregion) & !is.na(country) ~ str_c(subregion, country, sep = ", "),
+        !is.na(subregion) & is.na(country) ~ str_c(subregion, origin, sep = ", "),
+        is.na(subregion) & !is.na(country) ~ country ,
+        is.na(subregion) & is.na(country) ~ origin
+    )) 
+
+# extract the unique locations
+
+dat <- select(df1, place) %>% unique()
+
+# interact with GoogleMaps API to recover coordinates:
+
+res <- dat %>% 
+    mutate_geocode(location = place)
+
+## Some records resulted in errors:
+probs <- res %>%  filter(is.na(lon))
+
+probs <- probs %>% 
+    mutate(place = case_when(
+        place == "Southeastern region, China" ~ "Guangzhou, China", 
+        place == "Mato Grosso, Bolivia" ~ "Mato Grosso, Brazil",
+        place == "South region, Congo" ~ "Kinshasa, Congo",
+        place == "Central region, Argentina" ~ "Cordoba, Argentina",
+        place == "Southeastern region, Gabon" ~ "Ogooué-Maritime, Gabon",
+    ))
+
+res2 <- probs %>% 
+    mutate_geocode(location = place)
+
+res <- res2 %>% 
+    rename(lon = `lon...4`, lat = `lat...5`) %>% 
+    select(place, lon, lat) %>% 
+    bind_rows(res)
+
+save(res, file = "data/coords.RData")
+
+
+## TODO: now all coords are calculated, I need to correct and standardize names 
+## on the dataset to be able to merge them.

### 2021-09-23 box.R
delete box.R

@@ -1,23 +1,0 @@
-library(tidyverse)
-library(httr)
-library(boxr)
-
-
-box_auth(client_id = "jemswi6nx6bie6xep36kgrcvibbgpw71", 
-         client_secret = "2AHfw8XXGPU6I37a5KYQ7aF2a10quWXM")
-
-### failed attempts with httr
-myapp <- oauth_app(
-    "box", 
-    key = "6a15rlvm5w3au3l3p09o5o68soo8slhd", 
-    secret = "8khpqPzalF6M0WpvMOg60xYTvGT2lbCS")
-
-box_endpoint <- oauth_endpoint(
-    request = NULL,
-    authorize = "https://account.box.com/api/oauth2/authorize?client_id=214593079&response_type=cod",
-    access = "https://account.box.com/api/oauth2/authorize?client_id=214593079&response_type=cod", 
-    base_url = "https://account.box.com/api/oauth2/authorize"
-)
-
-
-box_token <- oauth2.0_token(box_endpoint, myapp)

### 2021-11-08 companies.Rmd
updating cleaning data routine to avoid double counts

--- 2021-03-10 companies.Rmd
@@
 author: "Juan Rocha"
-date: "21-03-10"
+date: "Last update `r Sys.Date()`"
 output:
   html_notebook:
-    code_folding: hide
+    self_contained: true
+  html_document:
+    self_contained: false
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, tidy = TRUE)
 ```
@@
 library(tidyverse)
-library(plotly)
+# library(plotly)
 library(patchwork)
@@
 
-Import data:
+## Import data
 
@@
 df1 <- readxl::read_xlsx(
-    path = "data/Drivers_Increasing_Risk_ED-2.xlsx",
+    path = "data/Paulas_files/001-Drivers_Increasing_Risk_ED.xlsx",
     sheet = 1,
@@
 df2 <- readxl::read_xlsx(
-    path = "data/Drivers_Increasing_Risk_ED-2.xlsx",
+    path = "data/Paulas_files/001-Drivers_Increasing_Risk_ED.xlsx",
     sheet = 2, 
@@
 ```{r}
-g1 <- df1 %>%
+g1 <- df2 %>%
     group_by(country) %>% 
@@
     labs(title = "Companies per country") +
-    theme_light(base_size = 8)
+    theme_light(base_size = 7)
 
-g2 <- df1 %>%
+g2 <- df2 %>%
     group_by(country) %>% 
@@
     coord_flip() +
-    labs(title = "Country")+
-    theme_light(base_size = 8)
+    labs(title = "Country of operations")+
+    theme_light(base_size = 7)
 
-g3 <- df1 %>%
+g3 <- df2 %>%
     group_by(origin) %>% 
@@
     coord_flip() +
-    labs(title = "Origin")+
-    theme_light(base_size = 8)
+    labs(title = "Country of origin")+
+    theme_light(base_size = 7)
 
 g1 + g2 + g3
+
+# ggsave(
+#     filename = "country_stats.png",
+#     plot = (g1 + g2 + g3),
+#     device = "png",
+#     path = "figures/",
+#     width = 6, height = 2,
+#     bg = "white", dpi = 400
+# )
 ```
 
-Top companies are:
+Top companies are ranked by geographies where a geography is a combination of countries * commodities that they produce. Only the heavy tail of the distribution is shown:
 
 ```{r}
-g1 <- df1 %>%
+g1 <- df2 %>%
   group_by(company) %>% 
@@
   geom_density() +
-    geom_vline(xintercept = 10, color = "orange")
+  geom_vline(xintercept = 10, color = "orange") +
+  theme_light(base_size = 6)
 
-g2 <- df1 %>%
+g2 <- df2 %>%
   group_by(company) %>% 
@@
   coord_flip() +
-    theme(axis.text.y = element_text(size = 6))
+  labs(x= "Top companies", y = "Geographies") +
+  theme(axis.text.y = element_text(size = 6)) +
+  theme_light(base_size = 6)
 
-g2 + inset_element(g1, left = 0.25, bottom = 0.05, right = 0.95, top = 0.65)
+g2 + inset_element(g1, left = 0.45, bottom = 0.05, right = 0.99, top = 0.45)
+
+# ggsave(
+#     filename = "top_companies.png",
+#     plot = last_plot(),
+#     device = "png",
+#     path = "figures/",
+#     width = 6, height = 3,
+#     bg = "white", dpi = 400
+# )
 
@@
 
-And the most importat commodities:
+Similarly, the most important commodities by number of companies * places where they are exploited:
 
 ```{r}
-g1 <- df1 %>%
+g1 <- df2 %>%
     group_by(commodity) %>% 
@@
     geom_density() +
-    geom_vline(xintercept = 30, color = "orange")
+    geom_vline(xintercept = 30, color = "orange") +
+    theme_light(base_size = 6)
 
-g2 <- df1 %>%
+g2 <- df2 %>%
     group_by(commodity) %>% 
@@
     coord_flip() +
-    theme(axis.text.y = element_text(size = 6))
+    theme(axis.text.y = element_text(size = 6)) +
+    theme_light(base_size = 6)
 
 g2 + inset_element(g1, left = 0.25, bottom = 0.05, right = 0.95, top = 0.65)
+
+# ggsave(
+#     filename = "top_commodities.png",
+#     plot = last_plot(),
+#     device = "png",
+#     path = "figures/",
+#     width = 4, height = 3,
+#     bg = "white", dpi = 400
+# )
+
 ```
@@
 ```{r}
-df1 %>%
-    group_by(commodity) %>% 
-    summarize(count = n()) %>%
-    arrange(desc(count)) %>% 
-    filter(count > 0) %>% 
-    mutate(commodity = as_factor(commodity) %>% fct_reorder(count)) %>% 
-    ggplot(aes(commodity, count)) +
+df3 <- read_csv(
+    file = "data/Paulas_files/004-Financial_summary.csv") %>% 
+    janitor::clean_names()
+
+g1 <- df3 |> 
+    arrange(desc(companies)) |> 
+    top_n(10) |> 
+    mutate(guo_final = as_factor(guo_final) |> 
+               fct_reorder( companies, sort)) |> 
+    ggplot(aes(companies, guo_final)) +
     geom_col() +
-    geom_hline(yintercept = 30, color = "orange") +
-    coord_flip() + 
-    facet_wrap()
-    theme(axis.text.y = element_text(size = 6))
+    labs(title = "", x = "Number of companies", y ="") +
+    theme_light(base_size = 7)
 
+g2 <- df3 |> 
+    arrange(desc(holdings)) |> 
+    top_n(10) |> 
+    mutate(guo_final = as_factor(guo_final) |> 
+               fct_reorder( holdings, sort)) |> 
+    ggplot(aes(holdings, guo_final)) +
+    geom_col() +
+    labs(title = "", x = "Number of holdings", y ="")+
+    theme_light(base_size = 7)
+
+g3 <- df3 |> 
+    arrange(desc(size_ownership)) |> 
+    top_n(10) |> 
+    mutate(guo_final = as_factor(guo_final) |> 
+               fct_reorder( size_ownership, sort)) |> 
+    ggplot(aes(size_ownership, guo_final)) +
+    geom_col() +
+    labs(title = "", x = "Millions of dollars", y ="")+
+    theme_light(base_size = 7)
+
+g1 +g2 + g3
+
+ggsave(
+    filename = "top_shareholders.png",
+    plot = (g1 +g2 + g3),
+    device = "png",
+    path = "figures/",
+    width = 8, height = 3,
+    bg = "white", dpi = 400
+)
 ```
@@
 1. Bipartite network of countries and companies
-    - One mode network of countries connected by number of companies
-    - One mode network of companies who operate on the same countries
-    - Can one figure out cooperation / competition for resources?
+    - ~~One mode network of countries connected by number of companies~~ [done]
+    - ~~One mode network of companies who operate on the same countries~~ [done]
+    - **Can one figure out cooperation / competition for resources?**
 2. For a selection of countries we could zoom into subnational networks (e.g. USA)
@@
 ```{r}
-df1 %>% 
+df2 %>% 
     select(company, origin, country) %>% 
@@
     unique() %>% 
-    summarize(companies = n()) %>% 
+    summarize(companies = n()) %>% #arrange(desc(companies))
     ggplot(aes(origin, country)) +
@@
 ```
-Hard to read, maybe better on table form:
+Hard to read, maybe better on table form below. Note that the higher numbers of companies are for local ones, they are registered and exploiting resources within the same country:
 
 ```{r}
-df1 %>% 
+df2 %>% 
     select(company, origin, country) %>% 
@@
 ```
+Be aware of missing values on the contry column.
 
-### Errors
+## Ideas to explore:
 
-- `DriversED` rows 102 onwards have shifted column order for columns L, M. Needs to be corrected to be able to do network analysis
-- There are typos on countries names e.g. Brazil and Brasil
-
-```{r}
-df1 %>% 
-    filter(company == "NESTLE S.A.")
-```
+- Hot spots of interest within countries
+- Commodities within the hotspots driven by land use (high priority cases)
+- ~~Explore Allen data and see if one can make linkages to specific diseases~~ [done]
+- One can use risk areas to link companies -> hotspots -> disease risk
+- Once a subset of companies is identified, go up to shareholders: which ones are more exposed? [tripartite networks]

### 2021-11-08 networks.Rmd
updating cleaning data routine to avoid double counts

--- 2021-06-10 networks.Rmd
@@
 author: "Juan Rocha"
-date: "6/10/2021"
-output: 
-  html_notebook:
-      theme:
-          code_font:
-            google: Fira Code
+date: "Last update `r Sys.Date()`"
+
+output: html_notebook
 ---
@@
     echo = TRUE, message=FALSE, warning = FALSE, error = FALSE,
-    fig.width = 6, fig.height = 5, tidy = TRUE, dpi  = 300)
+    fig.width = 6, fig.height = 5, tidy = TRUE)
 ```
 
-```{r}
+```{r message = FALSE, warning = FALSE, error = FALSE}
 library(tidyverse)
@@
 
-Read data:
+## Read data
 
@@
 
+## Networks
+
+### Countries
+
+Connected by companies and resource flow
+
 ```{r}
-library(sna)
+
 net_countries <-  df1 %>% 
@@
   # unique() %>% 
-    count(name = "channel") %>% # a channel is acountry * commodities
+  count(name = "channel") %>% # a channel is a country * commodities
   filter(!is.na(origin), !is.na(country)) %>% 
-    network(directed = TRUE, bipartite = FALSE, matrix.type = "edgelist", ignore.eval=FALSE, loops = FALSE)
+  network(directed = TRUE, bipartite = FALSE, matrix.type = "edgelist", ignore.eval=FALSE, loops = TRUE)
 
 net_countries %v% "indegree" <- sna::degree(
-    net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE, cmode = "indegree")
+  net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE, 
+  cmode = "indegree")
+
 net_countries %v% "outdegree" <- sna::degree(
-    net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE, cmode = "outdegree")
-net_countries %v% "degree" <- sna::degree(net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE)
+    net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE, 
+    cmode = "outdegree")
+
+net_countries %v% "degree" <- sna::degree(
+  net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE)
 
@@
     label = network.vertex.names(net_countries),
-    label.cex = 0.5, label.col = "grey50",
+    label.cex = 0.75, label.col = "grey50",
     vertex.border = 0, 
@@
 
+```
+
+```{r include = FALSE}
+# Alternative visualizations:
 # networkD3::simpleNetwork(
@@
 #     linkColour = "grey", nodeColour = "blue")
+library(ggnetwork)
+
+g1 <- ggplot(ggnetwork(net_countries, arrow.gap = 0.01), 
+       aes(x = x, y = y, xend = xend, yend = yend)) +
+    geom_edges(
+        aes(color = channel), size = 0.15, alpha = .5,
+        arrow = arrow(length = unit(2, "pt"), type = "closed")) +
+    geom_nodes(aes(size = outdegree), color = "red", alpha = 0.4) +
+    geom_nodes(aes(size = indegree), color = "blue", alpha = 0.4) +
+    geom_nodetext(aes(label = vertex.names, size = degree/4)) +
+    scico::scale_color_scico(
+        "Channel", palette = "batlow", direction = -1,
+        guide = guide_colorbar(
+            barwidth = unit(3, "mm"), barheight = unit(1.5, "cm"))) +
+    scale_size(name = "Degree", range = c(1,2)) +
+    labs(title = "Network of countries connected by companies",
+         subtitle = "A channel is a company times commodities. Red nodes are influencers while blue are influenced per\nnumber of connections", caption = "Data source: Orbis database") +
+    theme_blank(base_size = 6) 
+# 
+# ggsave(
+#     filename = "figures/countries_net.png",
+#     plot = g1,
+#     device = "png",
+#     width = 4, height = 3,
+#     bg = "white"
+# )
 
 ```
+
 Maybe fit a PCA to the data with network stats, commodities and so forth per country, and use the coordinates for visualization.
 Currently using a multi-dimensional scaling projection.
-Another way of "seeing" the network is by ranking which countries are the most influencial given their links to other countries (_outdegree_), once self-links are removed (companies whose country of origin and operation is the same):
+Another way of "seeing" the network is by ranking which countries are the most influential given their links to other countries (_outdegree_), once self-links are removed (companies whose country of origin and operation is the same):
 
@@
 
-```{r}
-df1 %>% 
+```{r fig.width = 6, fig.height = 5}
+g2 <- df1 %>% 
   select(company, commodity) %>% 
@@
   arrange(desc(geographies)) %>% 
-    ggplot(aes(company, commodity)) +
-    geom_tile(aes(fill = geographies)) +
-    scale_fill_viridis_c() +
-    theme_light(base_size = 6) +
-    theme(axis.text.x = element_text(angle = 90)) 
+  filter(geographies > 3) %>%  ungroup() %>% 
+  mutate(company = as_factor(company)) %>%
+  ## the data needs to be ungrouped for fct_reorder to work!
+  mutate(company = fct_reorder(
+    .f = company, .x = geographies, .fun = sum, .desc = FALSE)) %>% 
+  #pull(company) %>% levels()
+  ggplot(aes(y = company)) +
+  geom_bar(aes(weight = geographies, fill = commodity),
+           position = "stack", show.legend = TRUE, alpha = 0.5) +
+  scale_fill_discrete(
+    "Commodities", guide = guide_legend(
+      ncol = 4, title.position = "top", title.hjust = 0.5,
+      keywidth = 0.5, keyheight = 0.5
+    )
+  ) +
+  labs(title = "Companies ranking by number of countries and commodities",
+       subtitle = "A geography is a combination of `country * commodity` as a proxy of geographical influence of a company.\nOnly companies with >3 geographies are shown", caption = "Data curated by the project, source Orbis database", x = "Geographies", y = "Top companies") +
+  theme_light(base_size = 5) + theme(legend.position = c(0.7, 0.1))
+
+# ggsave(
+#     filename = "figures/companies_commodity.png",
+#     plot = g2,
+#     device = "png",
+#     width = 6, height = 5,
+#     bg = "white"
+# )
+
+g2
 ```
+
+Note there is a relatively low diversity of specialization within companies, meaning that a lot of them are putting their eggs on the same basket.
+This is not surprising.
+The knowledge and know-how of a company is limited and can only produce a set of products that are themselves similar in terms of the tech, knowledge and institutions.
+
+### Companies
+
+Companies connected by shared countries*resources.
+
+Another key network representation is by companies given the countries where they together operate, and by the common geographies (meaning exporting the same commodity from the same country).
+
+```{r}
+bip_net <- df1 %>% 
+  select(company, origin, country) %>% 
+  # Assumption: if country is NA, it means the company operates in
+  #  its country of origin:
+  mutate(country = case_when(
+    is.na(country) ~ origin, !is.na(country) ~ country))  %>% 
+  group_by(company, country) %>% 
+  # a channel can be either different commodities, or different places within the same country
+  count(name = "channel") %>% 
+  ungroup() |> 
+  network(
+    directed = TRUE, bipartite = TRUE, matrix.type = "edgelist", 
+    ignore.eval=FALSE, loops = FALSE)
+
+bip_mat <- bip_net |> 
+  as.matrix.network.adjacency()
+```
+
+The bipartite matrix has `r dim(bip_mat)[1]` companies and `r dim(bip_mat)[2]` countries.
+
+```{r}
+net_comp <- network(
+  bip_mat %*% t(bip_mat), 
+  directed = FALSE)
+
+net_comp %v% "degree" <- sna::degree(
+  net_comp, gmode = "graph", rescale = FALSE)
+net_comp %v% "betw" <- sna::betweenness(
+  net_comp, gmode = "graph", rescale = TRUE)
+
+df_comps <- tibble(
+  companies = network.vertex.names(net_comp),
+  degree = net_comp %v% "degree",
+  betweenness = net_comp %v% "betw"
+)
+
+df_comps |> 
+  arrange(desc(degree))
+```
+
+```{r fig.cap="Network of companies working on the same countries regardless of commodity. Isolates has been excluded"}
+plot.network(
+  net_comp,
+  edge.col = alpha("grey50", 0.3),
+  edge.lwd = 0.1,
+  vertex.lty = 0,
+  vertex.cex = 0.2,
+  arrowhead.cex = 0.3,
+  displayisolates = FALSE)
+
+g3 <- ggplot(ggnetwork(net_comp, arrow.gap = 0.01),
+    aes(x = x, y = y, xend = xend, yend = yend)) +
+    geom_edges(color = "grey40", size = 0.15, alpha = .5,
+        arrow = arrow(length = unit(2, "pt"), type = "closed")) +
+    geom_nodes(aes(size = degree), color = "red", alpha = 0.4) +
+    geom_nodes(aes(size = betw), color = "blue", alpha = 0.4) +
+    #geom_nodetext(aes(label = vertex.names, size = degree/4)) +
+    scico::scale_color_scico(
+        "Channel", palette = "batlow", direction = -1,
+        guide = guide_colorbar(
+            barwidth = unit(3, "mm"), barheight = unit(1.5, "cm"))) +
+    scale_size(name = "Degree", range = c(1,5)) +
+    labs(title = "Network of companies",
+         subtitle = "Companies are connected if they share geographies", caption = "Data source: Orbis database") +
+    theme_blank(base_size = 6)
+
+# ggsave(
+#     filename = "figures/companies_net.png",
+#     plot = g3,
+#     device = "png",
+#     width = 4, height = 3,
+#     bg = "white"
+# )
+```
+
+```{r}
+bip_net <- df1 %>% 
+  select(company, origin, country) %>% 
+  # Assumption: if country is NA, it means the company operates in
+  #  its country of origin:
+  mutate(country = case_when(
+    is.na(country) ~ origin, !is.na(country) ~ country))  %>% 
+  group_by(company, country) %>% 
+  # a channel can be either different commodities, or different places within the same country
+  count(name = "channel") %>% 
+  ungroup() |> 
+  network(
+    directed = TRUE, bipartite = TRUE, matrix.type = "edgelist", 
+    ignore.eval=FALSE, loops = FALSE)
+
+bip_mat <- bip_net |> 
+  as.matrix.network.adjacency()
+```
+
+### Financial actors
+
+Investors linked by investments on the same companies:
+
+```{r}
+fls <- list.files(path = "data/Paulas_files/Financial_actors/")
+fls <- paste0("data/Paulas_files/Financial_actors/", fls)
+head(fls)
+```
+
+```{r financial-actors}
+dat <- map(
+  fls,
+  function(x) {
+    x |> 
+      read_delim(
+        delim = ";", col_types = "cccccccccc", name_repair = "minimal",
+        col_names = c("company", "name", "pas", "type", "direct_percent", "total_percent", "source", "date", "op_revenue", "employees"),
+        skip = 1, na = c("n.d.", "-")) |> 
+    janitor::clean_names() }
+)
+
+dat <- dat |> 
+  bind_rows()
+
+dat <- dat |> 
+  filter(company != "") |> # 2 empty rows
+  mutate(
+    company = str_to_title(company),
+    name = str_to_title(name),
+    total_percent = str_remove_all(total_percent, pattern = "±")
+  ) |> 
+  mutate(
+    total_percent = as.numeric(total_percent),
+    direct_percent = as.numeric(direct_percent), 
+    date = lubridate::my(date),
+    op_revenue = as.numeric(op_revenue),
+    employees = as.numeric(employees)
+  )
+
+skimr::skim(dat)
+
+# save(dat, file = "data/investors_cleaned.RData")
+```
+
+The dataset contains `r dat |> pull(name) |> unique() |> length()` investors and `r dat |> pull(company) |> unique() |> length()` companies extracting resources in areas prone to the emergence of zoonotic dieseases.
+
+```{r}
+# note there are companies that are investors in others, so it's not purely bipartite
+net_companies <- dat |> 
+  select(investors = name, company, total_percent) |> 
+  network(
+    directed = TRUE, bipartite = FALSE, matrix.type = "edgelist", 
+    ignore.eval=FALSE, loops = FALSE)
+
+net_companies %v% "outdegree" <- sna::degree(
+    net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE, 
+    cmode = "outdegree")
+```
+
+```{r}
+plot.network(
+  net_companies,
+  edge.col = alpha("grey50", 0.3),
+  edge.lwd = 0.1,
+  vertex.lty = 0,
+  vertex.cex = 0.2,
+  arrowhead.cex = 0.3
+  )
+
+g4 <- ggplot(ggnetwork(net_companies, arrow.gap = 0.01),
+    aes(x = x, y = y, xend = xend, yend = yend)) +
+    geom_edges(color = "grey40", size = 0.15, alpha = .5,
+        arrow = arrow(length = unit(2, "pt"), type = "closed")) +
+    geom_nodes(aes(size = outdegree), color = "orange", alpha = 0.4) +
+    #geom_nodes(aes(size = betw), color = "blue", alpha = 0.4) +
+    #geom_nodetext(aes(label = vertex.names, size = degree/4)) +
+    # scico::scale_color_scico(
+    #     "Channel", palette = "batlow", direction = -1,
+    #     guide = guide_colorbar(
+    #         barwidth = unit(3, "mm"), barheight = unit(1.5, "cm"))) +
+    scale_size(name = "Outdegree", range = c(0.2,4)) +
+    labs(title = "Network of shareholders",
+         subtitle = "Shareholders are connected if they have shares on the same companies", caption = "Data source: Orbis database") +
+    theme_blank(base_size = 6)
+
+# ggsave(
+#     filename = "figures/shareholders_net.png",
+#     plot = g4,
+#     device = "png",
+#     width = 4, height = 3,
+#     bg = "white"
+# )
+```
+
+```{r}
+out_df <- tibble(
+  actor = network.vertex.names(net_companies),
+  outdegree = net_companies %v% "outdegree"
+)
+
+out_df |> 
+  ggplot(aes(outdegree)) +
+  geom_density()
+```
+
+There is a few investors with a disproportionate number of investments in the network.
+
+```{r}
+out_df |> 
+  arrange(desc(outdegree))
+```

### 2021-11-08 02-actors_descriptive_stats.R

@@ -1,0 +1,202 @@
+## Extracting descriptive stats of financial actors
+## Juan Rocha
+## 210929
+
+library(tidyverse)
+library(fs)
+library(here)
+
+# The cleaning replicates code from the networks.Rmd file:
+fls <- dir_ls(path = "data/Paulas_files/Financial_actors/")
+
+
+dat <- map(
+    fls,
+    function(x) {
+        x |> 
+            read_delim(
+                delim = ";", col_types = "cccccccccc", name_repair = "minimal",
+                col_names = c("company", "shareholder", "country", "type", "direct_percent", "total_percent", "source", "date", "op_revenue", "employees"),
+                skip = 1, na = c("n.d.", "-"),
+                locale = locale(encoding = "latin1")) |> 
+            janitor::clean_names() }
+) 
+
+
+dat <- dat |> 
+    bind_rows() 
+
+dat <- dat |>
+    filter(company != "") |> # 2 empty rows
+    mutate(
+        company = str_to_title(company),
+        shareholder = str_remove_all(shareholder, " via its funds"),
+        #shareholder = str_to_title(shareholder),
+        total_percent = str_remove_all(total_percent, pattern = ">|±|<"),
+        direct_percent = str_remove_all(direct_percent, pattern = ">|±|<"),
+        direct_percent = case_when(
+            direct_percent == "NG" ~ "0.01",
+            direct_percent == "MO" ~ "50",
+            direct_percent == "FME" ~ "NA",
+            TRUE ~ direct_percent
+        ), 
+        total_percent = case_when(
+            str_detect(total_percent,"84.94") ~ "84.94", # correct the encoding
+            TRUE ~ total_percent
+        ),
+        op_revenue = str_replace_all(op_revenue, ",", ".")
+    ) |> #pull(direct_percent) |> unique() |> as.numeric()
+    mutate(
+        total_percent = as.numeric(total_percent),
+        direct_percent = as.numeric(direct_percent), 
+        date = lubridate::my(date),
+        op_revenue = as.numeric(op_revenue),
+        employees = as.numeric(employees)
[…]

### 2021-11-08 FinancialActors_analysis.R

@@ -1,0 +1,272 @@
+## Title: Financial Actors Analysis
+## Author: Paula A. Sanchez
+## Modified by Juan
+## date: 4/26/2021 and 20210928
+
+
+library(tidyverse)
+
+
+companies  <- read.csv("data/Paulas_files/002-PublicCompanies.csv", header = TRUE, sep = ";")
+|> unique.shareholders <- read.csv("data/Paulas_files/003-list_shareholders_v2.csv", header = TRUE, sep = ";", fileEncoding = "latin1")
+
+companies <- companies %>%  filter(Type == "Public")
+
+colnames(unique.shareholders) <- c("Shareholder", "GUO.orbis", "GUO.final")
+unique.shareholders <- unique.shareholders %>% select(-c(GUO.orbis))
+  
+
+names.files <- dir("data/Paulas_files/Financial_actors/")
+
+## Create an empty vector for merging all files 
+
+shareholders <- data.frame(matrix(ncol=10, nrow=0))
+colnames(shareholders) <- c("Company", "Shareholder", "Country", "Type", "Direct", "Total", "Source", "Date", "Revenue", "Employees")
+
+## Using a loop to create a unique data frame containing all shareholders 
+
+for(i in 1:length(names.files)){
+  
+  scrfile <- paste("data/Paulas_files/Financial_actors/", names.files[i], sep="")
+  data <- read.csv(scrfile, header = TRUE, fileEncoding = "latin1", sep=";")
+  
+  colnames(data) <- c("Company", "Shareholder", "Country", 
+                      "Type", "Direct", "Total", "Source", 
+                      "Date", "Revenue", "Employees")
+  data$Shareholder <- gsub(" via its funds", "", data$Shareholder)
+  
+  
+  shareholders <- rbind(shareholders, data)
+  
+}
+
+## Replacing non-numeric values in data frame and converting Direct and Total ownership into numeric classes
+
+shareholders$Direct <- gsub("±", "", shareholders$Direct)
+shareholders$Direct <- gsub("<", "", shareholders$Direct)
+shareholders$Direct <- gsub(">", "", shareholders$Direct)
+
+shareholders$Total <- gsub("±", "", shareholders$Total)
+shareholders$Total <- gsub("<", "", shareholders$Total)
+shareholders$Total <- gsub(">", "", shareholders$Total)
+
+## J210928: This needs explanation: What is FME, NG, MO?
[…]

### 2021-11-15 Notes.Rmd
update cleaning data + viz started draft ergms for national networks

@@
 2. Same at commodity level: are there commodities that are purely trade driven?
-3. Explore Paula's file: rank countries per number of companies on high risk
-4. Extract main commodities
+3. ~~Explore Paula's file: rank countries per number of companies on high risk~~
+4. ~~Extract main commodities~~
 5. Characterize missing values: where are they missing (countries) and what vars are not very useful?
-6. Construct networks between countries and corporations weighted by subsidiaries?
+6. ~~Construct networks between countries and corporations weighted by subsidiaries?~~
+
+Exponential random graph models:
+- Include world governance indicators
+- Inequality
+- Triangles? Feedforward triangles -> theory on instability?
+- Valued ergm with number of companes count on link weight?

### 2021-11-15 02-actors_descriptive_stats.R

@@ -5,6 +5,7 @@
 library(tidyverse)
 library(fs)
 library(here)
+library(patchwork)
 
 # The cleaning replicates code from the networks.Rmd file:
 fls <- dir_ls(path = "data/Paulas_files/Financial_actors/")
@@ -200,3 +201,56 @@ dat |>
 
 
 # save(dat, file = "data/investors_cleaned.RData")
+
+### Visualizations
+load("data/investors_cleaned.RData")
+
+dat |> ggplot(aes(holdings)) + geom_density()
+
+dat <- dat |>
+    mutate(holdings = ownership * 0.2)
+
+g1 <- dat |> 
+    group_by(guo_final) |> 
+    summarise(companies = n()) |> 
+    arrange(desc(companies)) |> 
+    top_n(10) |> 
+    mutate(guo_final = as_factor(guo_final) |> 
+               fct_reorder( companies, sort)) |> 
+    ggplot(aes(companies, guo_final)) +
+    geom_col() +
+    labs(title = "", x = "Number of companies", y ="Global Unique Owner") +
+    theme_light(base_size = 10)
+
+g2 <- dat |> 
+    group_by(guo_final) |> 
+    summarise(holdings = sum(holdings)) |>  
+    arrange(desc(holdings)) |> 
+    top_n(10) |> 
+    mutate(guo_final = as_factor(guo_final) |> 
+               fct_reorder( holdings, sort)) |> 
+    ggplot(aes(holdings, guo_final)) +
+    geom_col() +
+    labs(title = "", x = "Number of holdings", y ="")+
+    theme_light(base_size = 10)
+
+# g3 <- df3 |> 
+#     arrange(desc(size_ownership)) |> 
+#     top_n(10) |> 
+#     mutate(guo_final = as_factor(guo_final) |> 
+#                fct_reorder( size_ownership, sort)) |> 
+#     ggplot(aes(size_ownership, guo_final)) +
+#     geom_col() +
+#     labs(title = "", x = "Millions of dollars", y ="")+
+#     theme_light(base_size = 7)
+
+g1 + g2 
+
+ggsave(
+    filename = "top_shareholders.png",
+    plot = (g1 +g2),
+    device = "png",
+    path = "figures/",
+    width = 8, height = 3,
+    bg = "white", dpi = 400
+)

### 2021-11-15 03-governance.R

@@ -1,0 +1,45 @@
+## exploring World Governance Index dataset from the World Bank
+## Juan Rocha
+## 211108
+## Source: https://datacatalog.worldbank.org/search/dataset/0038026/Worldwide-Governance-Indicators
+## Background info: http://info.worldbank.org/governance/wgi/
+
+library(tidyverse)
+
+wgi <- read_csv(file = "~/Documents/Projects/DATA/WorldBank/WGI_csv/WGIData.csv") |> 
+    janitor::clean_names()
+
+wgi |> 
+    pull(indicator_name) |> 
+    unique()
+
+
+## Some notes:
+## Estimate	Estimate of governance (ranges from approximately -2.5 (weak) to 2.5 (strong) governance performance)
+# StdErr	Standard error reflects variability around the point estimate of governance.
+# NumSrc	Number of data sources on which estimate is based
+# Rank	Percentile rank among all countries (ranges from 0 (lowest) to 100 (highest) rank)
+# Lower	Lower bound of 90% confidence interval for governance, in percentile rank terms
+# Upper	Upper bound of 90% confidence interval for governance, in percentile rank terms
+
+## Work only with estimates for now:
+
+wgi <- wgi |> 
+    filter(str_detect(indicator_name, "Estimate")) |> 
+    select(-x27) |> 
+    pivot_longer(cols = starts_with("x"), names_to = "year", values_to = "value") |> 
+    mutate(year = str_remove(year, "x")) |> 
+    mutate(year = as.numeric(year))
+
+
+wgi |> 
+    ggplot(aes(year, value, group = country_code)) +
+    geom_path(alpha = 0.2, size = 0.5) +
+    facet_wrap(~indicator_name)
+
+df_wgi <- wgi |> 
+    group_by(country_name, indicator_name) |> 
+    summarize(estimate = mean(value, na.rm = TRUE)) |> 
+    mutate(indicator_name = str_remove_all(indicator_name, ": Estimate"))
+
+save(df_wgi, file = "data/governance_index_cleaned.RData")

### 2021-11-15 04-ergms.R

@@ -1,0 +1,181 @@
+## exponential random graph models of financial actors
+library(tidyverse)
+library(sna)
+library(ergm)
+library(broom)
+
+#### Load data ####
+df1 <- readxl::read_xlsx(
+    path = "data/Paulas_files/001-Drivers_Increasing_Risk_ED.xlsx",
+    sheet = 2,
+    na = "NA") %>% 
+    janitor::clean_names()
+
+#### Network of countries ###
+# This first analysis merge all commodities into "geographies". Subsequent analysis 
+# could separate by commodity and see if it holds
+# Exclude nodes that are not countries:
+no_country <- c("Central Europe", "Eastern Europe", "Middle East", "Southern Europe", "Europe", "Fejér")
+net_countries <-  df1 %>% 
+    select(company, origin, country) %>% 
+    group_by(origin, country) %>% 
+    # unique() %>% 
+    count(name = "channel") %>% # a channel is a country * commodities
+    filter(!is.na(origin), !is.na(country)) %>% 
+    # Get rid of destinations that are not countries
+    filter(!(origin %in% no_country), !(country %in% no_country)) |> 
+    # remove "Liechtenstein" to include inequality metrics:
+    filter(origin != "Liechtenstein" ) |> 
+    ## standardize names to combine later with wgi dataset
+    mutate(
+        origin = case_when(
+            origin == "Croacia" ~ "Croatia",
+            origin == "Hong Kong" ~ "Hong Kong SAR, China" ,
+            origin == "Ivory Coast" ~ "Cote d'Ivoire" ,
+            origin == "New Zeland" ~ "New Zealand",
+            origin == "Russia" ~ "Russian Federation",
+            origin == "Yemen" ~ "Yemen, Rep.",
+            origin == "South Korea" ~ "Korea, Rep." ,
+            origin == "UK" ~ "United Kingdom"  ,
+            origin == "USA" ~ "United States", 
+            origin == "Slovakia" ~ "Slovak Republic" ,
+            origin == "Congo" ~ "Congo, Rep.",
+            origin == "Republic of Sao Tome" ~ "Sao Tome and Principe" ,
+            origin == "Venezuela" ~ "Venezuela, RB" ,
+            origin == "Laos" ~ "Lao PDR",
+            origin == "Bosnia" ~ "Bosnia and Herzegovina" ,
[…]

### 2022-01-31 02-actors_descriptive_stats.R
update on ergms for several cases start buidling investor networks

@@ -206,7 +206,6 @@ dat |>
 load("data/investors_cleaned.RData")
 
 dat |> ggplot(aes(holdings)) + geom_density()
-
 dat <- dat |>
     mutate(holdings = ownership * 0.2)
 

### 2022-01-31 04-ergms.R

@@ -22,7 +22,7 @@ library(patchwork)
 #alternatively: it contains the case study column needed for spliting.
 load("data/casestudies.RData")
 
-#### Network of countries ###
+#### Network of countries ####
 # This first analysis merge all commodities into "geographies". Subsequent analysis 
 # could separate by commodity and see if it holds
 # Exclude nodes that are not countries:

### 2022-01-31 05-investors.R

@@ -1,0 +1,76 @@
+# Extract key investors per case study
+library(tidyverse)
+
+# load data
+load("data/investors_cleaned.RData")
+load("data/casestudies.RData")
+
+
+dat <- dat |>
+    mutate(holdings = ownership * 0.2)
+
+dat |> ggplot(aes(holdings)) + geom_density()
+
+dat
+df1
+
+
+# Exclude nodes that are not countries:
+no_country <- c("Central Europe", "Eastern Europe", "Middle East", "Southern Europe", "Europe", "Fejér", "Taiwan")
+case_list <-  df1 %>% 
+    select(company, origin, country, casestudy) %>% 
+    group_by(origin, country, casestudy) %>% 
+    ## standardize names to combine later with wgi dataset
+    mutate(
+        origin = case_when(
+            origin == "Croacia" ~ "Croatia",
+            origin == "Hong Kong" ~ "Hong Kong SAR, China" ,
+            origin == "Ivory Coast" ~ "Cote d'Ivoire" ,
+            origin == "New Zeland" ~ "New Zealand",
+            origin == "Russia" ~ "Russian Federation",
+            origin == "Yemen" ~ "Yemen, Rep.",
+            origin == "South Korea" ~ "Korea, Rep." ,
+            origin == "UK" ~ "United Kingdom"  ,
+            origin == "USA" ~ "United States", 
+            origin == "Slovakia" ~ "Slovak Republic" ,
+            origin == "Congo" ~ "Congo, Rep.",
+            origin == "Republic of Sao Tome" ~ "Sao Tome and Principe" ,
+            origin == "Venezuela" ~ "Venezuela, RB" ,
+            origin == "Laos" ~ "Lao PDR",
+            origin == "Bosnia" ~ "Bosnia and Herzegovina" ,
+            origin == "Egypt" ~ "Egypt, Arab Rep.", 
+            origin == "Macedonia" ~ "North Macedonia" , 
+            origin == "Republic of Congo" ~ "Congo, Rep.",
+            TRUE ~ origin),
+        country = case_when(
+            country == "Croacia" ~ "Croatia",
+            country == "Hong Kong" ~ "Hong Kong SAR, China" ,
+            country == "Ivory Coast" ~ "Cote d'Ivoire" ,
+            country == "New Zeland" ~ "New Zealand",
+            country == "Russia" ~ "Russian Federation",
+            country == "Yemen" ~ "Yemen, Rep.",
+            country == "South Korea" ~ "Korea, Rep." ,
[…]

### 2022-03-11 02-actors_descriptive_stats.R
update some figures and prelim results

@@ -64,9 +64,10 @@ skimr::skim(dat)
 
 ## load the unique shareholders list: it was curated manually by Paula and Alice
 shareholders <- read_csv2(
-    file = "data/Paulas_files/003-list_shareholders_v2.csv", 
+    file = "data/Paulas_files/003-list_shareholders.csv", 
     locale = locale(encoding = "latin1")) |> 
-    janitor::clean_names()
+    janitor::clean_names() |> 
+    select(shareholder:guo_final)
 
 # shareholders <- shareholders |> 
 #     mutate(across(everything(), str_to_title))
@@ -125,8 +126,8 @@ dat <- dat |> mutate(
     )
 ) 
 
-dat |> filter(is.na(guo_orbis))
-
+dat |> filter(is.na(guo_orbis)) |> select(-c(source:employees))
+dat |> filter(is.na(guo_final)) |> select(-c(source:employees))
 #### J211010: This dataset still has some names that correspond to private persons, 
 #### not public companies. There were errors on Paula's list of GUOs, I corrected 
 #### them manually with the look up table: 003-list_shareholders_v2
@@ -146,7 +147,7 @@ dat |>
     add_count() |> 
     filter(n > 1) |> 
     arrange(desc(shareholder)) |> 
-    print(n = 34) 
+    print(n = 45) 
     
 
 dat <- dat |> 
@@ -158,30 +159,33 @@ dat <- dat |>
     mutate(ownership = pmax(direct_percent, total_percent, na.rm = TRUE),
            is_max = (ownership == max(ownership, na.rm = TRUE)))
 
-dat |> filter(count >1) |>arrange(desc(shareholder)) |> print(n = 34)
+dat |> filter(count >1) |>arrange(desc(shareholder)) |> print(n = 45)
 dat |> filter(count == 1, is.na(is_max)) # there is 298 records with no ownership (NA in direct and total %)
 
 dat <- dat |> 
-    filter(is_max == TRUE) # reduces the dataset to 3805 obs
+    filter(is_max == TRUE) # reduces the dataset to 5683obs
 
 ## Still errors: why? the final guo is also reported as shareholder, so double counted
 dat |> 
     ungroup() |> 
     group_by(company, guo_final) |> 
     summarize(total_own = sum(ownership)) |> 
-    arrange(desc(total_own))
+    arrange(desc(total_own)) |> 
+    print(n=10)
 
 ## the two problematic cases:
[…]

### 2022-03-11 05-investors.R

@@ -1,15 +1,28 @@
 # Extract key investors per case study
 library(tidyverse)
-
+library(sna)
+library(ggnetwork)
+library(sf)
+library(spData)
+library(scico)
+library(ggnewscale)
 # load data
 load("data/investors_cleaned.RData")
 load("data/casestudies.RData")
 
 
 dat <- dat |>
-    mutate(holdings = ownership * 0.2)
+    mutate(holdings = ownership * 0.05)
 
-dat |> ggplot(aes(holdings)) + geom_density()
+dat |> 
+    ggplot(aes(ownership)) + 
+    geom_density() +
+    scale_x_log10() +
+    theme_light(base_size = 6)
+
+ggsave(filename = "figures/ownership_distribution.png", 
+       device = "png", width = 2, height = 2)
+
 
 dat
 df1
@@ -17,9 +30,9 @@ df1
 
 # Exclude nodes that are not countries:
 no_country <- c("Central Europe", "Eastern Europe", "Middle East", "Southern Europe", "Europe", "Fejér", "Taiwan")
-case_list <-  df1 %>% 
-    select(company, origin, country, casestudy) %>% 
-    group_by(origin, country, casestudy) %>% 
+case_df <-  df1 %>% unique() |> 
+    select(company, commodity, casestudy, country, origin, type) %>% 
+    #group_by(origin, country, casestudy) %>% 
     ## standardize names to combine later with wgi dataset
     mutate(
         origin = case_when(
@@ -64,13 +77,268 @@ case_list <-  df1 %>%
             TRUE ~ country
         )
     ) |>
-    # unique() %>% 
-    count(name = "channel") %>% # a channel is a country * commodities
     filter(!is.na(origin), !is.na(country)) %>% 
     # Get rid of destinations that are not countries
     filter(!(origin %in% no_country), !(country %in% no_country)) |> 
     # remove "Liechtenstein" to include inequality metrics:
     filter(origin != "Liechtenstein" ) |> 
     ungroup() |>
-    select(origin,country, channel, casestudy) |> 
-    split(~casestudy)
+    filter(casestudy != "Other", type != "Private") |> 
+    # unique is needed because some companies operate in different subregions 
+    # within a country
+    unique() |> 
+    select(-country, -origin, -type) |> 
[…]

### 2022-04-04 Notes.Rmd
update on ergms for several cases start buidling investor networks
update some figures and prelim results
some figs updates

@@
 - Valued ergm with number of companes count on link weight?
+- Create network of investors combined for all cases
+- Create investor - country (where investors are based)
+- The later two should provide descriptive stats.
+
+Experiment with the HH-Index: https://en.wikipedia.org/wiki/Herfindahl–Hirschman_index
+
+Cannot make it bipartite because:
+The following vertices appear in both columns:
+	- Bnp Paribas
+	- Oji Holdings Corporation

### 2022-04-04 05-investors.R

@@ -10,8 +10,13 @@ library(ggnewscale)
 load("data/investors_cleaned.RData")
 load("data/casestudies.RData")
 
+dat ## companies - shareholders dataset
+df1 # company - commodities dataset
+df0 # case studies classification
 
 dat <- dat |>
+    ## Filtering ownership > 0.01 %: reduces obs from 4609 to 3621
+    filter(ownership > 0.01) |> 
     mutate(holdings = ownership * 0.05)
 
 dat |> 
@@ -22,10 +27,6 @@ dat |>
 
 ggsave(filename = "figures/ownership_distribution.png", 
        device = "png", width = 2, height = 2)
-
-
-dat
-df1
 
 
 # Exclude nodes that are not countries:
@@ -87,7 +88,7 @@ case_df <-  df1 %>% unique() |>
     # unique is needed because some companies operate in different subregions 
     # within a country
     unique() |> 
-    select(-country, -origin, -type) |> 
+    select(-country, -origin) |> 
     ## correct names for merging with investors data:
     mutate(
         company = case_when(
@@ -97,6 +98,8 @@ case_df <-  df1 %>% unique() |>
         )
     ) |> unique()
 
+case_df
+
 case_df$company [!case_df$company %in% dat$company] |> unique() # 4 in case_df but not in dat
 dat$company[!dat$company %in% case_df$company] |> unique() # 45 in dat but not in case_df
 
@@ -104,7 +107,7 @@ case_df |> pull(company) |> unique()  # 58 companies
 dat |> pull(company) |> unique() # 100 companies
 
 case_df <- case_df |> 
-    left_join(dat |> rename(shr_country = country)) ## 6.3k rows with countries; 3.9k without
+    left_join(dat |> rename(shr_country = country, shr_type = type)) ## 6.3k rows with countries; 3.9k without
 
 case_df |> filter(is.na(guo_final)) |>  pull(company) |> unique() 
 
@@ -117,12 +120,19 @@ case_df <- case_df |>
 case_df |> 
     group_by(shareholder, casestudy) |> 
     summarize(companies = n()) |> 
-    arrange(desc(companies)) |> 
-    ungroup() |> # group_by(casestudy) |> 
-    top_n(25) |> #print(n=200)
+    filter(companies > 1) |> 
+    arrange(casestudy, desc(companies)) |> 
+    ungroup() |> group_by(casestudy) |> 
[…]

### 2022-04-18 01-data_import.R
network viz updates (bipartite + one-mode)

@@ -5,12 +5,12 @@ url <- "https://stockholmuniversity.app.box.com/file/777762238977?s=8u6vl01c1efi
 ""
 ### Old datasets
 df1 <- readxl::read_xlsx(
-    path = "data/Drivers_Increasing_Risk_ED-2.xlsx",
+    path = "data/Drivers_Increasing_Risk_ED-3.xlsx",
     sheet = 1) %>% 
     janitor::clean_names()
 
 df2 <- readxl::read_xlsx(
-    path = "data/Drivers_Increasing_Risk_ED-2.xlsx",
+    path = "data/Drivers_Increasing_Risk_ED-3.xlsx",
     sheet = 2,
     na = "n.d") %>% 
     janitor::clean_names()

### 2022-04-18 02-actors_descriptive_stats.R

@@ -208,7 +208,7 @@ dat <- dat |> unique()
 ## remove individual persons from the analysis
 dat <- dat |> filter(type != "I")
 
-# save(dat, file = "data/investors_cleaned.RData")
+#save(dat, file = "data/investors_cleaned.RData")
 
 ### Visualizations
 load("data/investors_cleaned.RData")

### 2022-04-18 05-investors.R

@@ -6,6 +6,7 @@ library(sf)
 library(spData)
 library(scico)
 library(ggnewscale)
+library(ggrepel)
 # load data
 load("data/investors_cleaned.RData")
 load("data/casestudies.RData")
@@ -16,7 +17,7 @@ df0 # case studies classification
 
 dat <- dat |>
     ## Filtering ownership > 0.01 %: reduces obs from 4609 to 3621
-    filter(ownership > 0.01) |> 
+    #filter(ownership > 0.01) |> 
     mutate(holdings = ownership * 0.05)
 
 dat |> 
@@ -134,6 +135,15 @@ case_df |>
     facet_wrap(~casestudy, scales = "free_y") +
     theme_light(base_size = 6)
 
+ggsave(
+    filename = "top_investors_per_case.png",
+    plot = last_plot(),
+    device = "png",
+    path = "figures/",
+    width = 6, height = 5,
+    bg = "white", dpi = 400
+)
+
 case_df |> 
     group_by(shareholder, casestudy) |> 
     summarize(companies = n()) |> 
@@ -150,13 +160,44 @@ case_df |>
           legend.key.width = unit(3, "mm"))
 
 ggsave(
-    filename = "top_investors_higher001.png",
+    filename = "top_investors_higher001_by_type.png",
     plot = last_plot(),
     device = "png",
     path = "figures/",
     width = 3, height = 3,
     bg = "white", dpi = 400
 )
+
+case_df |> 
+    group_by(shareholder, casestudy, shr_type) |> 
+    mutate(shr_type = case_when(
+        is.na(shr_type) ~ "Company",
+        shr_type == "A" ~ "Insurance company",
+        shr_type == "B" ~ "Bank",
+        shr_type == "C" ~ "Corporate companies",
+        shr_type == "E" ~ "Mutual and pension fund",
+        shr_type == "F" ~ "Financial company",
+        shr_type == "J" ~ "Foundation, research institute",
+        shr_type == "L" ~ "Unnamed shareholders",
+        shr_type == "P" ~ "Private equity firms",
+        shr_type == "S" ~ "Public authorities, states, government",
+        shr_type == "V" ~ "Venture capital",
+        shr_type == "Y" ~ "Hedge fund",
+        shr_type == "" ~ "I"
+    )) |> filter(shr_type != "I") |> 
+    summarize(companies = n()) |> 
+    arrange(desc(companies)) |> 
[…]

### 2022-04-20 02-actors_descriptive_stats.R
re-make figures with latest datasets (April 19)

@@ -177,14 +177,21 @@ dat |>
 dat |> filter(company == "Bio Pappel S.a.b") ## It seems that it double counts when accounting for individual persons
 dat |> filter(company == "Th Plantations Berhad") |> #print(n=32)
     filter(direct_percent> 70 | total_percent > 70) |> 
-    select(shareholder) ## It seem32)here was a big move of shares ~70% from one owner to another. Since we don't know who sold, we only keep the most recent data for this case.
+    select(shareholder) ## It seem32)here was a big move of shares ~70% from one owner to another. Since we don't know who sold, we only keep the most recent data for this case.\
+## three
+dat |> filter(company == "Nine Dragons Paper (Holdings) Limited") 
+    
+
 
 ## Solution: filter out the strange cases
 dat <- dat |> 
     ungroup() |> 
     filter(company != "Bio Pappel S.a.b" | type != "I") |> # get's rid of the individual
     # Lembaga Tabung Haji is a bank that owned >70% of the company that a year later is owned on similar % by the Government of Malasya. It get rid of two obs
-    filter(company != "Th Plantations Berhad" | shareholder != "Lembaga Tabung Haji") 
+    filter(company != "Th Plantations Berhad" | shareholder != "Lembaga Tabung Haji") |> 
+    ## same problem, keep only the most recent transaction
+    filter(company != "Nine Dragons Paper (Holdings) Limited" | shareholder != "Tfs Trust And Fiduciary Services Sa" ) |> 
+    filter(company != "Nine Dragons Paper (Holdings) Limited" | shareholder != "Bnp Paribas" )
 ## Reduces the data set to 3825 observations
 ## And the test is passed
 dat |> 
@@ -208,7 +215,7 @@ dat <- dat |> unique()
 ## remove individual persons from the analysis
 dat <- dat |> filter(type != "I")
 
-#save(dat, file = "data/investors_cleaned.RData")
+# save(dat, file = "data/investors_cleaned.RData")
 
 ### Visualizations
 load("data/investors_cleaned.RData")

### 2022-04-20 05-investors.R

@@ -7,6 +7,7 @@ library(spData)
 library(scico)
 library(ggnewscale)
 library(ggrepel)
+library(patchwork)
 # load data
 load("data/investors_cleaned.RData")
 load("data/casestudies.RData")
@@ -160,7 +161,7 @@ case_df |>
           legend.key.width = unit(3, "mm"))
 
 ggsave(
-    filename = "top_investors_higher001_by_type.png",
+    filename = "top_investors_type.png",
     plot = last_plot(),
     device = "png",
     path = "figures/",
@@ -233,7 +234,7 @@ case_df |>
     theme_light(base_size = 5)
 
 ggsave(
-    filename = "top_ownership_higher001_combined.png",
+    filename = "top_ownership.png",
     plot = last_plot(),
     device = "png",
     path = "figures/",
@@ -274,11 +275,12 @@ case_df |>
     theme_light(base_size = 5)
 
 ggsave(
-    filename = "top_holdings_higher001.png",
+    filename = "top_holdings_percase.png",
     plot = last_plot(),
     device = "png",
     path = "figures/",
-    width = 4.5, height = 4,
+    width = 7, height = 5,
+    #width = 4.5, height = 4,
     bg = "white", dpi = 400
 )
 
@@ -296,16 +298,17 @@ publicomp <- publicomp |>
 case_df |> 
     left_join(publicomp) |> 
     mutate(size_own = ownership * market_cap_mll) |> # skimr::skim() #induces 11 NAs
-    group_by(shareholder, casestudy) |> 
-    summarize(sum_own = sum(size_own, na.rm = TRUE)) |> unique() |> 
-    slice_max(order_by = sum_own, n = 5) |> 
+    group_by(shareholder) |> unique() |> 
+    summarize(sum_own = sum(size_own, na.rm = TRUE)) |> 
+    ungroup() |> #group_by(casestudy) |> 
+    slice_max(order_by = sum_own, n = 25) |> 
     mutate(shareholder = as_factor(shareholder)) |> 
-    #mutate(shareholder = fct_reorder(shareholder, sum_own, sort)) |> 
+    mutate(shareholder = fct_reorder(shareholder, sum_own, sort)) |> 
     ggplot(aes(sum_own, shareholder)) +
     #geom_col(alpha = 0.5, fill = "goldenrod") +
     geom_col() +
     labs(x = "Size of ownership in US$", y = "Top 25 shareholders") +
[…]

### 2022-05-05 02-actors_descriptive_stats.R
Improving network models and prelim graphs for paper

@@ -215,7 +215,7 @@ dat <- dat |> unique()
 ## remove individual persons from the analysis
 dat <- dat |> filter(type != "I")
 
-# save(dat, file = "data/investors_cleaned.RData")
+#save(dat, file = "data/investors_cleaned.RData")
 
 ### Visualizations
 load("data/investors_cleaned.RData")

### 2022-05-05 04-ergms.R

@@ -7,22 +7,13 @@ library(ggnetwork)
 library(patchwork)
 
 #### Load data ####
-# df1 <- readxl::read_xlsx(
-#     path = "data/Paulas_files/001-Drivers_Increasing_Risk_ED.xlsx",
-#     sheet = 2,
-#     na = "NA") %>% 
-#     janitor::clean_names()
-# 
-# df0 <- readxl::read_xlsx(
-#     path = "data/Paulas_files/001-Drivers_Increasing_Risk_ED.xlsx",
-#     sheet = 1,
-#     na = "NA") %>% 
-#     janitor::clean_names()
 
 #alternatively: it contains the case study column needed for spliting.
+# load data
+load("data/investors_cleaned.RData")
 load("data/casestudies.RData")
 
-#### Network of countries ####
+#### Network of countries per case study ####
 # This first analysis merge all commodities into "geographies". Subsequent analysis 
 # could separate by commodity and see if it holds
 # Exclude nodes that are not countries:
@@ -294,7 +285,7 @@ ggsave(
     dpi = 400, bg = "white"
 )
 
-#### One single network ####
+#### Network countries sharing companies: One single network ####
 net_countries <-  df1 %>% 
     select(company, origin, country) %>% 
     ## standardize names to combine later with wgi dataset
@@ -563,6 +554,299 @@ ggsave(
     bg = "white", dpi = 400
 )              
 
+### Network of countries given common shareholders
+## I need to recompile again the list of case studies with country of destination in it. It is a slighthly larged dataset that the used previously, because companies can appear multiple times within a country if they expoit different subnational regions. 
+
+# Exclude nodes that are not countries:
+no_country <- c("Central Europe", "Eastern Europe", "Middle East", "Southern Europe", "Europe", "Fejér", "Taiwan")
+case_df <-  df1 %>% unique() |> 
+    select(company, commodity, casestudy, country, origin, type) %>% 
+    #group_by(origin, country, casestudy) %>% 
+    ## standardize names to combine later with wgi dataset
+    mutate(
+        origin = case_when(
+            origin == "Croacia" ~ "Croatia",
+            origin == "Hong Kong" ~ "Hong Kong SAR, China" ,
[…]

### 2022-05-05 05-investors.R

@@ -96,6 +96,7 @@ case_df <-  df1 %>% unique() |>
         company = case_when(
             company == "Golden Energy And Resources Ltd" ~ "Golden Energy And Resources Limited" ,
             company == "Pt Astra Agro Lestari Tbk" ~ "Pt Astra Agro Lestari Tbkê",
+            company == "Nine Dragons Paper Holdings" ~ "Nine Dragons Paper (Holdings) Limited",
             TRUE ~ company
         )
     ) |> unique()
@@ -105,8 +106,8 @@ case_df
 case_df$company [!case_df$company %in% dat$company] |> unique() # 4 in case_df but not in dat
 dat$company[!dat$company %in% case_df$company] |> unique() # 45 in dat but not in case_df
 
-case_df |> pull(company) |> unique()  # 58 companies
-dat |> pull(company) |> unique() # 100 companies
+case_df |> pull(company) |> unique()  # 54 companies
+dat |> pull(company) |> unique() # 99 companies
 
 case_df <- case_df |> 
     left_join(dat |> rename(shr_country = country, shr_type = type)) ## 6.3k rows with countries; 3.9k without
@@ -118,6 +119,7 @@ case_df <- case_df |>
     filter(!is.na(guo_final)) # 
     #filter(direct_percent > 0.01) # 0.01 is Neglegible (NG) in Orbis language
 
+case_df
 
 case_df |> 
     group_by(shareholder, casestudy) |> 
@@ -291,7 +293,7 @@ publicomp <- read_csv2(
     janitor::clean_names()
 
 publicomp <- publicomp |> 
-    filter(type == "Public") |>
+    #filter(type == "Public") |>
     mutate(company = str_to_title(company)) |> 
     select(company, market_cap_mll) |> unique()
 

### 2022-06-09 Notes.Rmd
network viz updates (bipartite + one-mode)
updates and figures

@@
 	- Oji Holdings Corporation
+
+One could force it to be bipartite by changing name according to category.
+Once bipartite, use the one mode projections to calculate network statistics more relevant for influence.
+For example a weighted betweenness or something like that.
+
+### EGRMs
+
+- use constraints on degree: keep degrees constant. But not sure how to code it
+- Play with control features:
+        - MCMCLE.maxit = < 60 which is the current default (how do I get the defaults?)

### 2022-06-09 04-ergms.R

@@ -5,8 +5,8 @@ library(ergm)
 library(broom)
 library(ggnetwork)
 library(patchwork)
+library(ggnewscale)
 
-#### Load data ####
 
 #alternatively: it contains the case study column needed for spliting.
 # load data
@@ -286,7 +286,7 @@ ggsave(
 )
 
 #### Network countries sharing companies: One single network ####
-net_countries <-  df1 %>% 
+net_edgelist <-  df1 %>% 
     select(company, origin, country) %>% 
     ## standardize names to combine later with wgi dataset
     mutate(
@@ -340,7 +340,9 @@ net_countries <-  df1 %>%
     filter(!(origin %in% no_country), !(country %in% no_country)) |> 
     # remove "Liechtenstein" to include inequality metrics:
     filter(origin != "Liechtenstein" ) |> 
-    ungroup() |> #unique() |> 
+    ungroup() 
+
+net_countries <- net_edgelist |> #unique() |> 
     network(directed = TRUE, bipartite = FALSE, matrix.type = "edgelist", 
             ignore.eval=FALSE, loops = TRUE)
 
@@ -406,38 +408,113 @@ net_countries %v% "regulatory_qual" <- df_attr$`Regulatory Quality`
 net_countries %v% "rule_law" <- df_attr$`Rule of Law`
 net_countries %v% "voice_acc" <- df_attr$`Voice and Accountability`
 net_countries %v% "gini" <- df_attr$gini_mean
+net_countries %e% "chanlog" <- log1p(net_countries %e% "channel")
 
-
-#### ergms ####
-f0 <- ergm(net_countries ~ edges) # null model
-# Governance models
+#### ergm: binary ####
+# null model
+f0 <- ergm(net_countries ~ edges)
+# governance model
+tic()
 f1 <- ergm(
-    net_countries ~ edges + absdiff("corruption") + absdiff("gov_effectiveness") +
-        absdiff("pol_stability") + absdiff("regulatory_qual") + 
-        absdiff("rule_law") + absdiff("voice_acc"))
+    net_countries ~ edges + diff("corruption") + diff("gini") +
+        diff("gov_effectiveness") +  diff("pol_stability") +
+        diff("regulatory_qual") + diff("rule_law") + diff("voice_acc"))
+toc()
+# governance + inequality
+tic()
+f2 <- ergm(
+    net_countries ~ edges + diff("corruption") + diff("gini") +
[…]

### 2022-06-09 05-investors.R

@@ -21,6 +21,7 @@ dat <- dat |>
     #filter(ownership > 0.01) |> 
     mutate(holdings = ownership * 0.05)
 
+
 dat |> 
     ggplot(aes(ownership)) + 
     geom_density() +
@@ -106,7 +107,7 @@ case_df
 case_df$company [!case_df$company %in% dat$company] |> unique() # 4 in case_df but not in dat
 dat$company[!dat$company %in% case_df$company] |> unique() # 45 in dat but not in case_df
 
-case_df |> pull(company) |> unique()  # 54 companies
+case_df |> pull(company) |> unique()  # 57 companies
 dat |> pull(company) |> unique() # 99 companies
 
 case_df <- case_df |> 
@@ -147,7 +148,7 @@ ggsave(
     bg = "white", dpi = 400
 )
 
-case_df |> 
+a <- case_df |> 
     group_by(shareholder, casestudy) |> 
     summarize(companies = n()) |> 
     arrange(desc(companies)) |> 
@@ -158,6 +159,8 @@ case_df |>
         .f = shareholder, .x = companies, .fun = sum, .desc = FALSE)) |> #pull(shareholder) |> levels()
     ggplot(aes(companies, shareholder)) +
     geom_col(aes(fill = casestudy)) +
+    scale_fill_brewer("Case study", palette = "Set2") +
+    labs(tag = "A", y = "Top shareholders", x = "Number of companies") + 
     theme_light(base_size = 6) +
     theme(legend.position = c(0.7, 0.2), legend.key.height = unit(3,"mm"),
           legend.key.width = unit(3, "mm"))
@@ -297,9 +300,11 @@ publicomp <- publicomp |>
     mutate(company = str_to_title(company)) |> 
     select(company, market_cap_mll) |> unique()
 
-case_df |> 
+b <- case_df |> 
+    select(company, shareholder, ownership) |> 
+    unique() |> 
     left_join(publicomp) |> 
-    mutate(size_own = ownership * market_cap_mll) |> # skimr::skim() #induces 11 NAs
+    mutate(size_own = (ownership/100) * market_cap_mll) |> # skimr::skim() #induces 11 NAs
     group_by(shareholder) |> unique() |> 
     summarize(sum_own = sum(size_own, na.rm = TRUE)) |> 
     ungroup() |> #group_by(casestudy) |> 
@@ -309,9 +314,9 @@ case_df |>
[…]

### 2022-10-13 ocean_disclosure_project.Rmd
quick check of companies for fisheries study

---
title: "Exploring fishing data"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Ocean disclosure project

```{r}
library(tidyverse)

# ocean disclosure project:
odp <- readxl::read_excel("data/005-Ocean_disclosure_database.xlsx", sheet = 2) |> 
    janitor::clean_names()
```

Summary:

```{r}
odp |> skimr::skim()
```
Data exist for the years `r odp$reporting_year |> as.numeric() |> range(na.rm = TRUE)`. It covers the following species <font color = "blue">`r odp$species |> unique()` </font> captured in:

```{r}
odp |> pull(catching_area) |> unique()
```

A preliminary list of companies fishing in "Gulf of Alaska":

```{r}
odp |> 
    filter(str_detect(catching_area, "Gulf of Alaska")) |> 
    pull(company) |> unique()
```

A preliminary list of companies fishing anchoveta in South America:

```{r}
odp |> 
    filter(species %in% c("Anchoveta", "Pacific anchoveta")) |> 
    pull(company) |> unique()
```

A preliminary list of companies fishing tuna:

```{r}
odp |> 
    filter(str_detect(species, "tuna")) |> 
    pull(company) |> unique()
```

Although not all of them are based in the Eastern Pacific.

```{r}
odp |> 
    filter(str_detect(species, "tuna")) 
```

```{r}
odp
```

## Peruvian anchoveta and sardine

```{r}
anch <- readxl::read_excel(
    "data/16_cuota_anchoveta-sardina_comun_v-x_2022_v20220908.xlsx",
    sheet = 2, range = "D6:L33") |> janitor::clean_names()
sard <- readxl::read_excel(
    "data/16_cuota_anchoveta-sardina_comun_v-x_2022_v20220908.xlsx",
    sheet = 2, range = "D39:L64") |> janitor::clean_names()

anch

```

```{r}
sard
```
Sardina española: It can be a mixture of companies, individual people, and coops

```{r}
sardesp <- readxl::read_excel("data/14_cuota_anchoveta-sardina_espanola_xv-iv_2022_v20220908.xlsx", sheet = "Publicacion web") |> 
    janitor::clean_names()

sardesp |> 
    select(organizacion_titular_area) |> 
    unique() |> 
    filter(!str_detect(organizacion_titular_area, "TOTAL ASIGNATARIO"))
```
Industrial actors only:

```{r}
sardesp <- readxl::read_excel("data/14_cuota_anchoveta-sardina_espanola_xv-iv_2022_v20220908.xlsx", sheet = 6, range = "B10:P47") |> 
    janitor::clean_names()

sardesp |> 
    select(titular_de_cuota_ltp) |> 
    filter(!is.na(titular_de_cuota_ltp)) |> 
    unique() 
```

## Alaskan pollock

```{r}
apo <- readxl::read_excel("data/002-Allocation_quotas_2020.xlsx", sheet = 2) |> 
    janitor::clean_names()

apo |> select(permit_holder) |> unique()
```

## Pacific tuna

Nothing on the folder

## Feedback from team

- ODP is a voluntary database, not complete. Missing Japan, China, etc.
- Clean up retailers? base on suppliers
- check Peter pan... data for pollock incomplete
- expect a lot of companies to be private
- skip orbis this time and instead of equity looking at debt
- Explore the database in the library...

## Next steps

1. Juan puts together a list of companies
2. Paula checks costs of licences at Library
3. Learning session with Paula. 
4. Check pollock info in Alaska (first NOAA, and Kevin)
5. Publish in Science

Global fishing watch:
 - IMO unique identifier of vessels: track over time
 - Paper Carmine 2020 OneEarth ? (Only for Tuna?) Check with Frida: how much we can trace back companies, and how much do we miss?

### 2022-10-13 04-ergms.R

@@ -346,7 +346,7 @@ df_attr <- df_attr |>
 
 fig5b <- df_attr |> 
     ggplot(aes(indegree, outdegree)) +
-    geom_point(aes(color = `Government Effectiveness`), size = 3) +
+    geom_point(aes(color = `Government Effectiveness`), size = 4, alpha = 0.9) +
     geom_text(aes(label = isoa2), color = "white", size = 2) + 
     geom_abline(slope = 1, intercept = 0, color = "red") +
     scico::scale_color_scico(
@@ -634,7 +634,7 @@ c <- ggplot(df_stats2, aes(AIC, model)) +
 
 fig6b <- df_attr2 |> 
     ggplot(aes(indegree, outdegree)) +
-    geom_point(aes(color = `Government Effectiveness`), size = 3) +
+    geom_point(aes(color = `Government Effectiveness`), size = 4, alpha = 0.9) +
     geom_text(aes(label = nodes), color = "white", size = 2) + 
     geom_abline(slope = 1, intercept = 0, color = "red") +
     scico::scale_color_scico(

### 2022-10-13 05-investors.R

@@ -121,7 +121,7 @@ case_df <- case_df |>
     filter(!is.na(guo_final)) # 
     #filter(direct_percent > 0.01) # 0.01 is Neglegible (NG) in Orbis language
 
-unique(case_df$shareholder) %in% shr_class$shareholder
+unique(case_df$shareholder) %in% shr_class$shareholder |> sum()
 
 
 
@@ -215,7 +215,7 @@ case_df |> #filter(shr_type != "I") |>
         .f = shareholder, .x = companies, .fun = sum, .desc = FALSE)) |> #pull(shareholder) |> levels()
     ggplot(aes(companies, shareholder)) +
     geom_col(aes(fill = shr_type)) +
-    scale_fill_brewer("Shareholder type",palette = "Set2") +
+    scale_fill_brewer("Shareholder type",palette = "Set2", na.value = "grey50") +
     theme_light(base_size = 5) +
     theme(legend.position = c(0.7, 0.2), legend.key.height = unit(2,"mm"),
           legend.key.width = unit(2, "mm"))
@@ -232,7 +232,7 @@ case_df |>
     ggplot(aes(sum_own, shareholder)) +
     geom_col(aes(fill = casestudy)) +
     labs(y = "Top 25 shareholders", x = "Sum of ownership") +
-    scale_fill_brewer("Case studies", palette = "Set3") +
+    scale_fill_brewer("Case studies", palette = "Set3", na.value = "grey50") +
     theme_light(base_size = 6) +
     theme(legend.position = c(0.8, 0.25))
 
@@ -523,6 +523,12 @@ shr_stats <- case_df |>
     summarize(mean_own = mean(ownership),
               sd_own = sd(ownership, na.rm = TRUE)) |># arrange(desc(sd_own)) |> print(n=100)
     right_join(shr_stats) 
+
+## problems:
+shr_stats |> filter(is.na(shr_type)) |> 
+    select(shareholder, degree) |> 
+    arrange(desc(degree)) |> 
+    print(n=61)
 
 p1 <- shr_stats |> filter(shr_type != "I") |> 
     mutate(label = ifelse( betw > 200 | degree > 180 | mean_own > 60, shareholder, "")) |> 

### 2022-10-13 99-fisheries.R

@@ -1,0 +1,10 @@
+## Exploring fisheries data
+
+library(tidyverse)
+
+# ocean disclosure project:
+odp <- readxl::read_excel("data/005-Ocean_disclosure_database.xlsx", sheet = 2) |> 
+    janitor::clean_names()
+
+odp |> skimr::skim()
+    

### 2022-12-22 ocean_disclosure_project.Rmd
back ups: graphs for paper, some notes, and data collection script

@@
 ```
+```{r}
+## from Alaskan pollock permits:
+## files ending in cp are catcher processors, endin gin cv are catcher vessels
+fls <- dir_ls("data/Alaska_pollock/")
+
+dat <- fls |> str_subset("afa") |> 
+    map(.f = function(x) read_csv(x, skip=1) |> janitor::clean_names()) |> 
+    map(select, owner_name) |> 
+    bind_rows() |> 
+    unique()
+dat
+
+write_csv(dat, file = "fishing_companies_alaskan-pollock.csv")
+```
+
+### Extract companies
+
+```{r}
+
+companies <- bind_rows(
+    anch |> select(company = titular_de_cuota_ltp) |> unique(),
+    sard |> select(company = titular_de_cuota_ltp) |> unique(),
+    sardesp |> select(company = titular_de_cuota_ltp) |> unique()
+) |> unique()
+
+companies 
+```
+
+```{r}
+write_csv(companies, file = "fishing_comps_sardina-anchoveta.csv")
+```
 

### 2022-12-22 05-investors.R

@@ -159,7 +159,8 @@ a <- case_df |>
     ungroup() |> 
     mutate(shareholder = as_factor(shareholder)) |> 
     mutate(shareholder = fct_reorder(
-        .f = shareholder, .x = companies, .fun = sum, .desc = FALSE)) |> #pull(shareholder) |> levels()
+        .f = shareholder, .x = companies, .fun = sum, .desc = FALSE)) |> 
+    #pull(shareholder) |> levels()
     ggplot(aes(companies, shareholder)) +
     geom_col(aes(fill = casestudy)) +
     scale_fill_brewer("Case study", palette = "Set2") +
@@ -180,27 +181,50 @@ a <- case_df |>
 case_df <- case_df |> 
     left_join(shr_class) |> 
     #group_by(shareholder, casestudy, shr_type) |> 
-    mutate(shr_type = str_sub(class, start = 1L, end = 1L)) |> 
+    mutate(shr_type2 = str_sub(class, start = 1L, end = 1L)) |> 
+    mutate(shr_type2 = case_when(
+        # this is the NACE classification
+        #is.na(shr_type2) ~ "Missing",
+        shr_type2 == "A" ~ "Agriculture, forestry and fishing",
+        shr_type2 == "B" ~ "Mining and quarrying",
+        shr_type2 == "C" ~ "Manufacturing",
+        shr_type2 == "D" ~ "Electricity, gas, steam and air conditioning supply",
+        shr_type2 == "F" ~ "Construction",
+        shr_type2 == "G" ~ "Wholesale and retail trade",
+        shr_type2 == "H" ~ "Transportation and storage",
+        shr_type2 == "I" ~ "Accommodation and food service activities",
+        shr_type2 == "J" ~ "Information and communication",
+        shr_type2 == "K" ~ "Financial and insurance activities",
+        shr_type2 == "L" ~ "Real estate activities",
+        shr_type2 == "M" ~ "Professional, scientific and technical activities",
+        shr_type2 == "N" ~ "Administrative and support service activities",
+        shr_type2 == "O" ~ "Public administration and defence",
+        shr_type2 == "P" ~ "Education",
+        shr_type2 == "Q" ~ "Human health and social work activities",
+        shr_type2 == "R" ~ "Arts, entertainment and recreation",
+        shr_type2 == "S" ~ "Other service activities"
+        #shr_type2 == "" ~ "I"
+    )) |> # below the orginal Orbis classification
     mutate(shr_type = case_when(
[…]

### 2022-12-22 99-orbis.R

@@ -1,0 +1,205 @@
+library(tidyverse)
+library(RSelenium)
+library(keyring)
+library(rvest)
+library(tictoc)
+
+
+## todo:
+## - implement the search in such a way that always picks up the first company (most likely result)
+## - check before going into shareholders if there are shareholders on the first result page: solved
+## - If not, create an error / message and record that as response: solved | it does not create an error message but return a df saying there is no companies.
+## - if multiple shareholders, loop through pages
+## - Create a loop or walking function that go through all of them (safely?)
+## 
+## - download the sardina-anchoveta data and tuna; for tuna use the open ocean project
+
+dat <- read_csv(file = "data/fishing_companies_alaskan-pollock.csv") # pollock
+dat <- read_csv(file = "data/fishing_comps_sardina-anchoveta.csv") # sardine-anchovies
+load("data/shr_class.RData")
+
+## Initialize remote driver
+d <- rsDriver(
+    port = 4444L, browser = "firefox", extraCapabilities = list(acceptInsecureCerts = TRUE)) # should open a chrome
+remDr <- d[["client"]]
+#remDr$open()
+tic()
+remDr$navigate(url = "https://ezp.sub.su.se/login?url=https://orbis4.bvdinfo.com/ip")
+toc() # >40s
+## Note: working with VPN skips authentification steps! but takes longer to load
+Sys.sleep(3)
+
+## Authentication
+# me <- remDr$findElement("id", "username")
+# # me$clearElement()
+# me$sendKeysToElement(list("juro7132"))
+# pswd <- remDr$findElement("xpath", '//*[(@id = "password")]')
+# pswd$clearElement()
+# pswd$sendKeysToElement(list(keyring::key_get("sub", "juro7132"), key = "enter"))
+
+#### for loop should start here
+## create list to catch results
+revenues <- list()
+guo_list <- list()
+shr_list <- list()
+
+## re-start the search: here just for quick restart
+remDr$goBack()
+# You need to do the search manually (2-3 loops) until it gives you the option
+# of restarting search again 
+tic()
+for (i in seq_along(dat$company)){
+    ## Search one company:
+    srch <- remDr$findElement("id", "search")
+    
+    ## repeating the search allows me to clean up the previous search, that is why this
+    ## line is repeated, but second time with "enter"
[…]

### 2022-12-23 11-cleaning_data.R
cleaning data

@@ -1,0 +1,84 @@
+library(tidyverse)
+
+
+load("data/alaskan_pollock_orbis.RData")
+dat <- read_csv("data/fishing_companies_alaskan-pollock.csv")
+
+dat
+
+## Recovering IDs
+guo_list[[1]][1][[1]][2,1]
+
+dat$ids <-  guo_list |> 
+    map(function(x) x[[1]][[1]][2]) |> 
+    map(function(x) str_remove_all(x,pattern = "\n|\t|Orbis ID: ")) |> 
+    map(function(x) str_remove_all(x, "ActiveBvD ID: |Active \\(dormant\\)BvD ID: "))
+
+dat <- dat |> 
+    unnest(ids, keep_empty = TRUE) |> 
+    mutate(ids = str_replace(ids, "[:space:]", "_")) |> 
+    separate(col = ids, into = c("BvD_id", "orbis_id"), sep = "_")
+
+## We also have company address if needed
+## Recover GUO:
+
+guo_list[[1]][[1]][2,3]
+
+dat$guo <- guo_list |> 
+    map(function(x) x[[1]][2,3]) |> 
+    # recover GUOs
+    map(function(x) str_remove_all(x, pattern = "Private The Global Ultimate Owner of this controlled subsidiary is ")) |> 
+    map(function(x) str_replace(x, "Private This company is the Global Ultimate Owner of the corporate group", "self-GUO")) |> 
+    map(function(x) str_replace_all(x, "Private This company is Independent \\(but not the Global Ultimate Owner of a Corporate Group\\)", "independent")) |> 
+    map(function(x) str_replace_all(x, "Private This company is a Single location", "single")) |> 
+    map(function(x) str_remove_all(x, "Private This entity is a branch of "))
+
+dat <- dat |> 
+    unnest(guo, keep_empty = TRUE)
+
+## Recover revenues
+revenues[[1]][[1]][1,2] # units
+revenues[[1]][[1]][3,2] # revenue
+revenues[[1]][[1]][4,2] # employees
+
+dat$rev_units <- revenues |> map(function(x) x[[1]][2][[1]][1])
+dat$revenue <- revenues |> map(function(x) x[[1]][2][[1]][3])
+dat$employees <- revenues |> map(function(x) x[[1]][2][[1]][4])
+
+dat <- dat |> 
+    unnest(cols = rev_units:employees, keep_empty = TRUE) |> 
[…]

### 2023-02-23 Notes.Rmd
back ups: graphs for paper, some notes, and data collection script
update sst analysis

@@
 author: Juan Rocha
+editor_options: 
+  chunk_output_type: inline
 ---
+
+# Notes
+
+*20230202*
+
+- Download the Fishing effort dataset used in Krodsma y Carmine.
+It contain GPS traces of >70k vessels with some information about ownership.
+Carmine has addional owners identified but only for high seas vessels.
+- Problem: we only have ocean disclosure project data for pacific tuna, which are mainly retailiers and very few companies.
+Excludes China and all the key actors of that area.
+Go with the fine grane data, much better.
+
+- create a spatial mask to filter the vessels of interest: this is, regions of the world with exposure to tipping points: my map + heatwaves papers.
 
@@
         - MCMCLE.maxit = < 60 which is the current default (how do I get the defaults?)
+
+### Lloyds
+
+- P & I club is insurance
+- Download Carmine SM it has fishing vessel numbers and owners.
+opitions AHS markets: IMO global record of fishing vessles,. globalrecord.fao.org/vessels
+
+- Overlay management areas RFMO regional fisheries managmenet organisations maps...regulatory body that groups the relevant countries that need to cooperate to handle potential shifts in the fisheries.
+
+## Readings
+
+*Carmine 2020 OneEarth*
+- Combining satellite data with other public databases, our analysis showed 1,120 corporate actors owned 2,482 high seas fishing vessels in 2018 ($2/3 of total detected ves- sels).
+The industry is consolidated with 100 companies representing 36% of high seas fishing effort.
+- China, the fishing entity of Taiwan, Japan, Indonesia, Spain, and the Republic of Korea accounting for 77% of the identified global high seas fishing fleet and 80% of all high seas fishing effort for 2016.
+- Global Fishing Watch (GFW): gps records per vessel, 3.5k vessels in 2018
+- One of the main challenges in this analysis was determining the ultimate parent company. [we target that challenge]
+- At present, about 55,000 vessels have been identified with AIS as actively fishing in EEZs in 2018.
+- The fishing vessels with AIS data represent a fraction of the world’s total motorized fishing fleets of length greater than 12 m, estimated to be 300,000 vessels,
+
+Method:
+-  To avoid counting fishing hours associated with falsified AIS signals, we excluded, from this analysis, vessels whose supposedly unique AIS identities were simultaneously shared between two or more vessels over 72 h in total during 2018.
+- GFW’s vessel database attributed owner information to about 1,199 vessels relying mostly on publicly available data including IMO’s shipping information system.
+
+*kroodsma 2018 Science*
+
+- more than 70,000 identified fishing vessels that are 6 to 146 m in length. 
+- Global hot spots of fishing effort were seen in the northeast Atlantic (Europe) and northwest Pacific (China, Japan, and Russia) and in upwelling regions off South America and West Africa (Fig. 1A).
+- most nations fished predominantly within
+their own EEZ, with five flag states (China, Spain, Taiwan, Japan, and South Korea) accounting for more than 85% of observed fishing effort on the high seas
+- modern fishing is like other forms of mass production that are partially insulated from natural cycles and are instead shaped by policy and culture.
+
+**Frölicher et al 2018 Nature**
+
+- "Between 1982 and 2016, we detect a doubling in the number of MHW days, and this number is projected to factor of 16 for global warming of 1.5 degrees Celsius relative to preindustrial levels and by a factor 23 for global warming of 2.0 degrees Celsius"...3.1C will be 41x times more MHW.
+- The modeling results when compared with satellite observations gives confidence to the spatial extend of predictions, less so to their duration, cumulative mean intensity or absolute spatial extend.
+- at 3.5C warming, a one-in-a-hundred-days event will become a one-in-three-days event!
+- Excellent paper!

### 2023-02-23 manuscript.Rmd
update sst analysis

---
title: "Identifying companies and financial actors exposed to marine tipping points"
author: |
    | \small Juan C. Rocha^1^, Jean-Baptiste Jouffray^1^, Paula A. Sánchez^1,2^, Victor Galaz^1^
    | \footnotesize ^1^Stockholm Resilience Centre, Stockholm University
    | \footnotesize ^2^Germany?
    | \small \texttt{\href{mailto:juan.rocha@su.se}{\nolinkurl{juan.rocha@su.se}}}
output:
  pdf_document:
    dev: pdf
    keep_tex: yes
    toc: no
    latex_engine: xelatex
  rtf_document:
    toc: no
  word_document:
csl: /Users/juanrocha/Documents/styles/apa.csl
#bibliography: refs.bib
citecolor: blue
documentclass: article
font: Arial
fontsize: 10pt
linkcolor: blue
urlcolor: blue
header-includes:
- \usepackage{dcolumn, rotating, longtable, lineno, float, array, tabularx, inputenc}
- \setmainfont{Arial}
- \renewcommand{\abstractname}{}
- \graphicspath{{figures/}}
- \usepackage[margin=2.5cm]{geometry}
- \linenumbers
- \usepackage{setspace}
---

# Notes

- Author order tbd
- Target journal: Nature family?, GEC, Finance journal? ~ 4000w, 4 figs, 40refs
- Word budget
    - Intro: 1000w
    - Method: 800w
    - Results: 1000w
    - Discussion: 800w
    - Conclusions: 400w

\pagebreak

\begin{abstract}

\end{abstract}

# Introduction

- Marine ecosystems are expected to undergo large biological reorganization and regime shifts that can affect primary productivity and other ecosystem services.
- Exposure to such regime shifts are mediated by pressures on marine ecosystems, from climate change, fishing pressure, or pollution.
    - Marine heatwaves
    - Destabilization of upwellings
- Companies and financial actors are part of the problem by being partly responsible for the causes of ecological degradation. These actors are in turn exposed to the risk of regime shifts and decline of ecosystem services.
- RQ: **Which financial actors are exposed to regime shift risks, and thus have the agency to avoid them or can play a role in navigating transformations?**

# Methods

_Case studies:_ To that end, we first identified areas of the oceans exposed to potential regime shifts. Marine heatwaves are expected to increase in frequency and intensity [@Frolicher:2018dc; @Frolicher:2018cm], with stark consequences for the economy of companies, regions and nations depending on the ocean productivity [@Smith_2021]. For example, the Gulf of Alaska experienced a heat anomaly known as the Blob in 2014-2019, impacting a fishery valued at US\$103M / year, with a estimated loss of US\$24.4M, while harmful algae blooms costed US\$40M loss in tourist spending in the coast of Washington in 2015 [@Smith_2021]. Other areas with high vulnerability to heatwaves include the tropical Pacific ocean, the Caribbean sea, the North coasts of Australia, the Eastern China sea and the West Pacific ocean [@Smith_2021]. Proxies of resilience loss in the oceans have been calculated using early warnings to the proximity of tipping points in time series data of primary productivity [@Rocha_2022]. Areas showing strong signals or resilience loss include the Easter Indo-Pacific and the tropical Easter Pacific oceans.

- Maybe a figure Map with the case studies

_Companies_: Based on the geographic exposure to regime shifts, we then identified fishing companies operating in these regions. We used permits and licences issued in Alaska by the National Oceanic and Atmospheric Administration in the United States ([NOAA](https://www.fisheries.noaa.gov/alaska/commercial-fishing/permits-and-licenses-issued-alaska)) to list companies with permits to fish Alaskan pollock, one of the most profitable resources in the region. We used fishing permits issued by the _Servicio nacional de Pesca y Acuicultura_ or the National Fishing and Aquaculture authority of Chile, to companies with the purpose of harvesting anchovies and sardines.

- check with Paula, we should have Peru as well
- what to do with tuna? we only have O

_Financial actors_: Using the [Orbis](https://login.bvdinfo.com/R0/Orbis) database we extracted the shareholder information for publicly listed companies (names and approximate ratio of ownership), their number of employees, and annual revenues. We complemented the data when possible with information on debt and liquidity from the [Eikon refinitiv](https://eikon.refinitiv.com) database. Unfortunately these data sources are not open access and their terms of service forbids sharing accessed data with third parties. Both of them were accessed through research subscriptions by the home institution.

_Systemic risk exposure_:

# Results

# Discussion

- Limitation: publicly listed companies, and legal fisheries.

# Conclusion

## Acknowledgements

# References {.unnumbered}

::: {#refs}
:::

\pagebreak

# Supplementary Material {#sec:SM}

### 2023-02-23 processing_sst.Rmd
update sst analysis

---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

This notebook takes all `.nc` files from NOAA with sea surface temperature data and merge them into one big file. It follows the rationale of this [tutorial](https://robwschlegel.github.io/heatwaveR/articles/Download_SST_v2.html).
Note that `cdo` needs to be previously installed. If not available, follow instructions [here](https://code.mpimet.mpg.de/projects/cdo/embedded/index.html#x1-30001.1).

Working locally on my mac, the command was `brew install cdo` on the terminal.

```{bash}
cd ~/Documents/Projects/DATA/OISST
time cdo mergetime *.nc OISST_combined.nc
```

A few files do not have the expected size (~1.6Mb), instead they have a few Kb. Need to be re-downloaded manually or programatically. Less than 10, doing manual.s

J230223: Output created, file moved now under the root directory of the project/data/

### 2023-02-23 12-GFW_data.R

@@ -1,0 +1,13 @@
+## Exploring the global fish watch data
+
+library(tidyverse)
+
+
+dat <- read_csv("~/Documents/Projects/DATA/GlobalFishWatch/9c3d6bc0-a3c1-11ed-a61d-7ba19957371b/fishing-vessels-v2.csv")
+
+dat |> names()
+dat |> skimr::skim()
+dat |> pull(registries_listed) |> unique()
+dat |> pull(vessel_class_registry) |> unique() ## identifies tuna as target
+dat
+

### 2023-02-23 heatwaves.R

@@ -1,0 +1,64 @@
+library(tidyverse)
+library(heatwaveR)
+library(fs)
+library(here)
+library(raster)
+library(tictoc)
+
+# Analyze the sea surface temperature data to identify places where they can happen
+
+fls <- dir_ls("~/Documents/Projects/DATA/OISST/")
+# a dataframe with the dates will make easier to filter out 
+df_fls <- tibble(file = fls)
+df_fls <- df_fls |> 
+    mutate(date = file |>
+               str_remove("/Users/juanrocha/Documents/Projects/DATA/OISST/oisst-avhrr-v02r01.")) |> 
+    mutate(date = str_remove(date, ".nc"))
+
+
+# one file: can be use to extract relevant coordinates
+dat <- raster(fls |> str_subset("20071022"), varname = "sst") # only one file
+dat
+
+## create a mask without NAs
+coords <- as.data.frame(dat, xy = TRUE) |> 
+    as_tibble()|> 
+    filter(!is.na(Daily.sea.surface.temperature)) |> 
+    dplyr::rename(lon = x, lat = y, sst = Daily.sea.surface.temperature)
+# creates a unique id of the pixels that need to be studied
+coords <- coords |> 
+    unite(col = "id", lon, lat, sep = "_", remove = FALSE)
+
+## Multiple files: one month - test the statistics.
+tic()
+stk <- stack(fls |> str_subset("202212"), quick = TRUE, varname = "sst")
+toc() # 0.6s for one month,
+
+stk_df <- as.data.frame(stk, xy = TRUE) |> 
+    as_tibble() # 261MB World
+
+
+tic()
+stk_df |> 
+    pivot_longer(cols = starts_with("Daily"), names_to = "day", values_to = "temp") |> 
+    mutate(day = str_remove(day, "Daily.sea.surface.temperature.")) |> 
+    group_by(x, y) |> 
+    filter(all(is.na(temp)))
+toc() ## takes too long, rather pre-compute pixels that should not be used.
+
+raster::plot(dat)
+
+
+
+
+#### leftover ####
+# example code to analyze one ts
+
+head(heatwaveR::sst_WA)
+ts <- ts2clm(sst_WA, climatologyPeriod = c("1982-01-01", "2011-12-31"))
+mhw <- detect_event(ts, categories = TRUE, climatology = TRUE)
+mhw$event |> names()
+
+mhw$climatology |> dplyr::filter(!is.na(category))
+mhw$event |> 
+    dplyr::select(category) |> table()

### 2023-03-02 heatwaves.R
detecting marine heatwaves

@@ -2,9 +2,113 @@ library(tidyverse)
 library(heatwaveR)
 library(fs)
 library(here)
-library(raster)
+library(tidync)
 library(tictoc)
+library(doParallel)
+registerDoParallel(cores = 10)
+#library(data.table)
 
+#### Heatwaves on big ncdf file ####
+# J230301: `processing_sst.Rmd` was used to combine all .nc files from NOAA into
+# a unique file that can be queried. Instead of reading files multiple times one
+# reads the big one once and call slices at the time.
+# Following code from the package vignette also available at:
+# https://robwschlegel.github.io/heatwaveR/articles/Download_SST_v2.html
+
+nc_file <- dir_ls("data/") |> str_subset("combined.nc") # one file
+datadir <- "data/processed_MHW" # processed marine heatwaves
+
+## Function to read data (from tutorial)
+OISST_load <- function(file_name, lon1, lon2) {
+    OISST_dat <- tidync(file_name) %>%
+        hyper_filter(lon = between(lon, lon1, lon2)) %>%
+        hyper_tibble(select_var = "sst", force = TRUE, drop = TRUE) %>%
+        dplyr::select(-zlev) %>%
+        dplyr::rename(t = time, temp = sst) %>%
+        mutate(t = as.Date(t, origin = "1978-01-01"))
+    return(OISST_dat)
+    rm(OISST_dat)
+}
+
+# Event detection function (from tutorial)
+event_only <- function(df) {
+    # first calculate the climatologies
+    clim <- ts2clm(data = df, climatologyPeriod = c("1989-01-01", "2019-12-31"))
+    # then the events
+    event <- detect_event(data = clim)
+    rm(clim)
+    # return only the event metric dataframe of results
+    return(event$event)
+    rm(event)
+}
+
+# Define the slices
+# 5° longitude slices seem to work fine on
+# my MacBook Pro with 32Gb RAM and 12 cores
+slice_df <- tibble(lon1 = seq(0, 355, 5),
+                   lon2 = seq(5, 360, 5))
+
+# test on one slide
+# tic()
+# sst <- OISST_load(nc_file, lon1 = slice_df$lon1[1], lon2 = slice_df$lon2[1]) |> 
+#     filter(t < "2020-02-01") ## added line to address problem below
+# toc() # 60s, including filter 74s
+
+## I get an error with the following:
+## Error in do.ply(i) :
+## task 1 failed - "The specified end date follows the last day of series, which is 2020-02-27"
[…]

### 2023-03-02 map_heatwaves.R
mapping heatwave results

@@ -1,0 +1,36 @@
+library(tidyverse)
+library(fs)
+library(here)
+library(tictoc)
+
+
+fls <- dir_ls("data/processed_MHW/")
+
+## function to summarize each file
+mhw_summary <- function(file){
+    tic()
+    dat <- read_csv(file)
+    x <- dat |> 
+        group_by(lon, lat) |> 
+        summarize(mean_intensity = mean(intensity_mean))
+    return(x)
+    rm(dat, x)
+    toc()
+}
+
+
+out <- map(fls, mhw_summary)
+
+out <- out |>
+    bind_rows()
+
+out |> 
+    ggplot(aes(lon, lat)) +
+    geom_tile(aes(fill = mean_intensity)) +
+    scale_fill_viridis_c("Mean intensity") + 
+    theme_void(base_size = 6) +
+    theme(legend.position = "bottom")
+    
+
+# I forgot to calculate the severity categories on `heatwaves.R`. It takes over 10hrs
+# to compute, so only do it if reviewers ask for it.

### 2023-03-04 Notes.Rmd
analyzing heatwaves

@@
 
+- Heatwaves are events when SST exceed the 99th percentile with respect to daily satalite observations for a 30yr period. 
 - "Between 1982 and 2016, we detect a doubling in the number of MHW days, and this number is projected to factor of 16 for global warming of 1.5 degrees Celsius relative to preindustrial levels and by a factor 23 for global warming of 2.0 degrees Celsius"...3.1C will be 41x times more MHW.
@@
 - Excellent paper!
+
+**Smith et al 2021 Science**
+
+- " MHW events exceed US\$800 million in direct losses or >US\$3.1 billion in indirect losses of ecosystem services for multiple years"
+- "discrete warming events, marine heatwaves (MHWs), have increased substantially, with a rise in global annual MHW days of 54% over the past century"
+- " an event is considered a MHW if it lasts for 5 or more days, with tem- peratures warmer than the 90th percentile based on a 30-year historical climatology (41)."

### 2023-03-04 heatwaves.R

@@ -35,7 +35,7 @@ event_only <- function(df) {
     # first calculate the climatologies
     clim <- ts2clm(data = df, climatologyPeriod = c("1989-01-01", "2019-12-31"))
     # then the events
-    event <- detect_event(data = clim)
+    event <- detect_event(data = clim, categories = TRUE, climatology = TRUE)
     rm(clim)
     # return only the event metric dataframe of results
     return(event$event)
@@ -89,14 +89,15 @@ system.time(
         # save results to disk
         cat(noquote("  > 3. saving events to csv\n"))
         data.table::fwrite(MHW, file = paste0(datadir, "/MHW_slice_", i, "_",
-                                  slice_df$lon1[i], "-", slice_df$lon2[i], ".csv"))
+                                  slice_df$lon1[i], "-", slice_df$lon2[i], ".csv"),
+                           showProgress = TRUE)
         rm(MHW)
         cat(noquote("SLICE DONE!\n"))
         toc()
         cat(sep="\n\n")
     }
 )
-toc() #38993.427 s = 10.8hrs
+toc() #38993.427 s = 10.8hrs; 14.6hrs with event category detection
 
 
 ## Results are saved on folder data/processed_MHW in csv files per slice.

### 2023-03-04 map_heatwaves.R

@@ -8,26 +8,62 @@ fls <- dir_ls("data/processed_MHW/")
 
 ## function to summarize each file
 mhw_summary <- function(file){
-    tic()
     dat <- read_csv(file)
     x <- dat |> 
         group_by(lon, lat) |> 
-        summarize(mean_intensity = mean(intensity_mean))
+        summarize(
+            mean_intensity = mean(intensity_mean),
+            n_events = n(),
+            tot_days = sum(duration))
     return(x)
     rm(dat, x)
-    toc()
+}
+# create a function that slice only one year to check if results align with other papers
+mhw_year <- function(file, year){
+    dat 
 }
 
+mhw_categories <- function(file){
+    dat <- read_csv(file)
+    x <- dat |> 
+        group_by(lon, lat, category) |>
+        add_column(n = 1) |> 
+        summarize(tot = sum(n)) |> 
+        pivot_wider(names_from = category, values_from = tot)
+    return(x)
+    rm(dat, x)
+}
+# create a function that extracts the proportion of the events above / below the 
+# categories thresholds.
+dat <- read_csv(fls[1])
 
+
+tic()
 out <- map(fls, mhw_summary)
+toc() # 1.3s
 
 out <- out |>
     bind_rows()
+# 
+# out |> 
+#     pivot_longer(cols = starts_with("I"), names_to = "category", values_to = "events" ) |> 
+#     filter(category != "I Moderate") |> 
+#     ggplot(aes(lon, lat)) +
+#     geom_tile(aes(fill = events)) +
+#     scale_fill_viridis_c(option = "D") + 
+#     facet_wrap(~category) +
+#     theme_void(base_size = 6) +
+#     theme(legend.position = "bottom")
+
+out |> ggplot(aes(mean_intensity)) + geom_density()
+out$mean_intensity |> summary()
 
 out |> 
-    ggplot(aes(lon, lat)) +
-    geom_tile(aes(fill = mean_intensity)) +
-    scale_fill_viridis_c("Mean intensity") + 
+    ungroup() |> 
+    ggplot(aes(lon, lat)) + 
+    geom_tile(aes(fill = mean_intensity, alpha = (mean_intensity > 1.45))) +
+    #geom_tile(aes(fill = mean_intensity, alpha = n_events)) +
+    scale_fill_viridis_c( option = "D") + 
     theme_void(base_size = 6) +
     theme(legend.position = "bottom")
     

### 2023-08-16 Notes.Rmd
draft report to share with collabs

@@
 # Notes
+
+*230816*
+
+- Created a report with preliminary data from Lloyds and Orbis. 
+- Data from Kroodsma was used to identify >16k vessels fishing in places prone to tipping
+- Places were extracted from Smith heatwaves methodology replicated on NOAAs temp data and my dataset on resilience loss
+- > 2000 companies identified, and >700 shareholders, but networks are too sparse
 

### 2023-08-16 marine_tipping.Rmd
draft report to share with collabs

---
title: "Identifying financial actors: marine case"
author: "Juan Rocha"
date: "2023-08-15"
output:
    html_document:
      theme:
        bootswatch: cosmo
        code_font:
            google: Fira Code
      df_print: paged
      code_folding: hide
      toc: true
      toc_float:
        collapsed: true
        smooth_control: true
      toc_depth: 3
      fig_caption: true
      fig_align: center
      highlight: pygments
      self_contained: false
      lib_dir: libs
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(tictoc)
library(patchwork)
library(here)
library(fs)
library(DT)
library(ggnetwork)
library(network)
```

Dear friends,

Here is a short report on the progress of our paper with tentative title: **Identifying financial actors exposed to marine tipping points**. I presented some preliminary results on a conference poster a month ago in Copenhagen. Here I summarize the progress so we can discuss and get input from you on next steps. But first, let's remind ourselves of the research question and data at hand:

> Who is exposed to ecological regime shift risks, and by being exposed, who has the agency or power to invervene and perhaps avoid tipping points?

## Datasets

The analysis builds upon two datasets and knitting a companies data based on publicly accessible databases. The datasets are:

1. Map of areas showing symptoms of resilience loss (Rocha 2022): ~25% of the marine area analyzed shows statistical signatures of resilience loss.

```{r rocha22, fig.cap='Marine areas prone to tipping points', echo = FALSE}
knitr::include_graphics(path = "figures/fig_detection_marine.png")
```

2. Areas prone to marine heatwaves (Smith et al 2021): Marine heatwaves are increasing in frequency and intensity, with socioeconomic consequences due to reduction of marine primary productivity.

![Smith 2021 map on marine heat waves](figures/Smith_2023_MHW.png)

But in order to analyze their map, I replicated their analysis using all data from NOAA on sea surface temperature (Smith only uses 3 decades) and created the following map with probabilities of heatwaves according with their severity:

![Heat waves probability (log units)](figures/mhw_probabilities_log.png)

The Global Fish Watch dataset (Kroodsma et al 2018) was used to identify vessels that have operations in pixels with high probability of heatwaves (a driver of regime shifts), or where resilience loss is detected. Out of >70 000 vessels, 16 878 do operate in high risk areas.

Then I used Lloyds (thanks Frida!) to identify owner companies. Out of the 16 878 vessels, I got companies for 2183 of them. Using Orbis I found 793 shareholders for some of the companies. Already now you can see part of our problem: each step we try to go on the chain (vessel - owner - shareholder) we lose observations (one order of magnitude), thus to what extend are results robust? Let's check them first.

## Results

```{r}
load("data/cleaned_data_230708.Rda")
load("data/boats_owners.Rda")
load("data/marine_shareholders.RData")
## some initial visualizations
countries <- readr::read_csv(
    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGCountry.csv") |> 
    janitor::clean_names()
## correct Namibia
countries[countries$short_name == "Namibia","x2_alpha_code"] <- "NA"

```

Most boats retrieved are indeed fishing boats:

```{r fig.width= 5, fig.height=3, fig.align='center'}
df_boats |> 
    ggplot(aes(lli_vessel_type)) +
    geom_bar() +
    coord_flip()
```

However, with Lloyds I can only recovered `r length(owners)` owners, and from that selection, only `r guos |> filter(!is.na(guo)) |> pull(guo) |> length()` have a global ultimate owner. The table below helps you explore the companies for which we have GUOs and the ones missing. **Do you have any suggestions to _automatically_ extract GUOs? e.g. different spelling companies?**

```{r}
options(DT.options = list(pageLength = 25))
datatable(guos)
```

Working with what we have, China is by far the country with most vessels fishing in marine areas prone to ecological tipping points. However, most of shareholders are based in Europe. Shareholders are mainly corporate companies and individual families. Despite Norway has by far most of reported shareholders, they invest mainly in Norwegian companies (we will see below).

```{r desc, fig.width=8, fig.height=4}
a <- df_boats |> 
    group_by(flag) |> 
    summarize(n = n()) |> 
    arrange(n) |> 
    filter(n> 300) |> 
    mutate(flag = case_when(is.na(flag) ~ "Unknown", TRUE ~ flag)) |> 
    mutate(flag = as_factor(flag)) |> 
    ggplot(aes(n, flag)) +
    geom_col() +
    labs(x = "Number of vessels", y = "Flag", 
         caption = "Only countries with > 300 vessels shown.\n Data source: Lloyds",
         tag = "a") +
    theme_light(base_size = 8)
b <- shrs |> 
    filter(!is.na(name)) |> 
    mutate(country = case_when(is.na(country) ~ "NA", TRUE ~ country)) |> 
    group_by(country) |>     
    summarize(n=n()) |> 
    filter(n>10) |> 
    left_join(select(countries, short_name, country = x2_alpha_code)) |> 
    mutate(short_name = case_when(is.na(short_name) ~ "Unknown", TRUE ~ short_name)) |> 
    arrange(n) |> 
    mutate(short_name = as_factor(short_name)) |> 
    ggplot(aes(n, short_name)) +
    geom_col() +
    labs(x = "Shareholders", y = "Country of origin", caption = "Data source: Orbis",
         tag = "b") +
    theme_light(base_size = 8)

# shareholder types
c <- shrs |> #pull(type) |> unique()
    filter(!is.na(name)) |> 
    mutate(shr_type = case_when(
        is.na(type) ~ "Missing",
        type == "A" ~ "Insurance company",
        type == "B" ~ "Bank",
        type == "C" ~ "Corporate companies",
        type == "D" ~ "Unnamed private shareholders",
        type == "F" ~ "Financial company",
        type == "G" ~ "Wholesale and retail trade",
        type == "H" ~ "Self ownership",
        type == "I" ~ "Individuals or families",
        type == "J" ~ "Foundation or Research Institute",
        type == "Z" ~ "Public",
        type == "L" ~ "Employees/Managers/Directors",
        type == "M" ~ "Professional, scientific and technical activities",
        type == "V" ~ "Venture capital",
        type == "Y" ~ "Hedge fund",
        type == "P" ~ "Private equity firms",
        type == "Q" ~ "Branch",
        type == "W" ~ "Marine vessel",
        type == "S" ~ "Public authorities, States,\n Governments",
        type == "E" ~ "Mutual & Pension Fund"
    )) |> 
    group_by(shr_type) |> 
    summarize(n=n()) |> 
    arrange(n) |> 
    mutate(shr_type = as_factor(shr_type)) |> 
    ggplot(aes(n, shr_type)) +
    geom_col() +
    labs(x = "Number of shareholders", y="Shareholder type", tag = "c",
         caption = "Data source: Orbis") +
    theme_light(base_size = 8)

a + b + c & theme_light(base_size = 8)

```

```{r}
# correct ownerships:
shrs <- shrs |> 
    rename(ownership_direct = owenership_direct) |> 
    mutate(ownership_direct = case_when(
        ownership_direct == "WO" ~ "100",
        ownership_direct == "" ~ NA,
        ownership_direct == "MO" ~ "50",
        ownership_direct == "T" ~ "100",
        ownership_direct == "-" ~ NA,
        TRUE ~ ownership_direct
    )) |> #pull(ownership_direct) |> unique()
    mutate(ownership_total = case_when(
        ownership_total == "n.a." ~ ownership_direct,
        ownership_total == "WO" ~ "100",
        ownership_total == "MO" ~ "50",
        ownership_total == "CTP" ~ ownership_direct,
        TRUE ~ ownership_total
    )) |> #pull(ownership_total) |> unique()
    mutate(ownership_direct = str_remove_all(ownership_direct, "\\>|\\<"),
           ownership_total = str_remove_all(ownership_total, "\\>|\\<")) |> 
    #pull(ownership_total) |> unique()
    mutate(ownership_direct = as.numeric(ownership_direct),
           ownership_total = as.numeric(ownership_total))
```

```{r fig.width=5, fig.height=4}
bip <- shrs |> 
    select(name, company, ownership_total) |> 
    filter(!is.na(name)) |> 
    slice(-338) |>  # problems: the name is the IMO and the shareholder name the same
    network(directed = FALSE, bipartite = TRUE, matrix.type = "edgelist", 
            ignore.eval=FALSE, loops = TRUE)

# plot(bip, edge.lwd = 0.001, edge.col = alpha("gray75", 0.75),
#     vertex.col = "orange", vertex.border = 0, vertex.cex = 0.5)

ggplot(ggnetwork::ggnetwork(bip, arrow.gap = 0.005) ,
       aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(aes(color = ownership_total), size = 0.35, alpha = 1
               #arrow = arrow(length = unit(2, "pt"), type = "closed")
               ) +
    geom_nodes( color = "orange", alpha = 0.4) +
    labs(title = "Bipartite network",
         subtitle = "Shareholders and companies", 
         caption = "Data source: Orbis") + 
    scico::scale_color_scico(
        "Total ownership", palette = "berlin",
        guide = guide_colorbar(barwidth = unit(2,"cm"), barheight = unit(2, "mm")))+
    theme_blank(base_size = 10) +
    theme(legend.position = "bottom")
```

The bipartite network of shareholders connected to companies if we don't have missing data. Please note that we started with over 2700 companies, but not for all there are GUOs, we only have `r shrs |>  filter(!is.na(name)) |> pull(name) |> length()` shareholder - company pairings (links), one of them is a self-link. In other words, I'm not showing the isolates, companies that we identified with Lloyds but for which Orbis did not provided shareholders. The data I'm using to build the network does not distinguish ownership over time, nor name changing. The density of the bipartite network is `r round(network.density(bip),4)`, meaning out of all possible links on the network < 0.1% are realized.

The bipartite network has two one mode projections: a network of shareholders connected to each other if they share investments to the same fishing companies (left), and a network of fishing companies connected to each other if they are invested by the same shareholders (right). As you can see, these networks are also very sparse with very few connections and no "gigant key stone actors" emerging from the data.

```{r fig.width=8, fig.height=4}

a <- shrs |> 
    ggplot(aes(ownership_total)) +
    geom_density() +
    labs(x = "Total ownership (%)", tag = "a") +
    #scale_x_continuous(labels = scales::label_percent()) +
    theme_light()

b <- shrs |> 
    group_by(company) |> 
    summarize(n = n()) |> 
    ggplot(aes(n)) + 
    geom_bar() +
    scale_y_log10() +
    labs(x = 'Number of shareholders', y = "Companies", tag = "b") +
    theme_light()

mat <- bip |> as.sociomatrix() #|> dim()

one <- mat %*% t(mat)
shr_net <- network(one, directed = FALSE)
shr_net %e% "w" <- one
shr_net %v% "outdegree" <- sna::degree(shr_net, gmode = "digraph", cmode = "outdegree")

two <- t(mat) %*% mat
cmp_net <- network(two, directed = FALSE)
cmp_net %e% "w" <- two
cmp_net %v% "outdegree" <- sna::degree(cmp_net, gmode = "digraph", cmode = "outdegree")

# plot(shr_net,edge.lwd = 0.001, edge.col = alpha("gray75", 0.75),
#     vertex.col = "orange", vertex.border = 0, vertex.cex = 0.5)

c <- ggplot(ggnetwork::ggnetwork(shr_net, arrow.gap = 0.005) ,
       aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(aes(color = w), size = 0.15, alpha = .5
               #arrow = arrow(length = unit(2, "pt"), type = "closed")
               ) +
    geom_nodes( color = "orange", alpha = 0.4) +
    scale_size(name = "Outdegree", range = c(0.2,4)) +
    labs(title = "Network of shareholders",
         subtitle = "Shareholders are connected if they have shares on the same fishing companies", 
         caption = "Data source: Orbis") + 
    scale_color_continuous(
        "Companies", breaks = c(1,2,3), 
        guide = guide_colorbar(barwidth = unit(2,"cm"), barheight = unit(2, "mm")))+
    theme_blank(base_size = 10) +
    theme(legend.position = "bottom")

e <- ggplot(ggnetwork::ggnetwork(cmp_net, arrow.gap = 0.005) ,
       aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(aes(color = w), size = 0.15, alpha = .5
               #arrow = arrow(length = unit(2, "pt"), type = "closed")
               ) +
    geom_nodes( color = "orange", alpha = 0.4) +
    scale_size(name = "Outdegree", range = c(0.2,4)) +
    labs(title = "Network of companies",
         subtitle = "Companies are connected if they have shares on the same shareholders", 
         caption = "Data source: Orbis") + 
    scale_color_continuous(
        "Shareholders", 
        guide = guide_colorbar(barwidth = unit(2,"cm"), barheight = unit(2, "mm")))+
    theme_blank(base_size = 10) +
    theme(legend.position = "bottom")

c + e
```

Because we have location for shareholders and vessel flags, we can also project the data in terms of countries. There is a few reasons to focus the analysis at that level: i) there is actually a network to analyze, ii) countries have the mandate to protect their citizens interests, so the can regulate companies and their investments either locally or through bilateral agreements. The network of countries is plotted below:

```{r poster, fig.width=8, fig.height=8}
net <- shrs |> 
    filter(!is.na(name), !is.na(country)) |> 
    select(country, name, company) |> 
    left_join(
        df_boats |> 
            select(flag, reg_owner) |> 
            filter(!is.na(reg_owner)) |> 
            unique(),
        by = c("company" = "reg_owner")
    ) |> 
    mutate(flag = str_remove_all(flag, " \\(Int. Register\\)")) |> 
    group_by(country, flag) |> 
    summarize(n = n()) |># print(n=151)
    left_join(countries |> 
                  select(short_name, x2_alpha_code), 
              by = c("country"= "x2_alpha_code")) |> #print(n = 151)
    mutate(short_name = case_when(
        country == "TW" ~ "Taiwan, China",
        TRUE ~ short_name)) |> 
    mutate(flag = case_when(
        flag == "U.K." ~ "United Kingdom",
        flag == "Republic of Ireland" ~ "Ireland",
        flag == "U.S.A." ~ "United States",
        TRUE ~ flag),
        short_name = case_when(
            short_name == "Dem. People's Rep. Korea" ~ "South Korea",
            TRUE ~ short_name
        )) |> 
    ungroup() |> 
    select(from = short_name, to = flag, n) |>
    filter(!is.na(from)) |> #print(n = 151)
    network(directed = TRUE, bipartite = FALSE, matrix.type = "edgelist", 
            ignore.eval=FALSE, loops = TRUE)

# network.vertex.names(net)[!network.vertex.names(net) %in% countries$short_name]

# plot(
#     net,
#     edge.lwd = 0.001, edge.col = alpha("gray75", 0.75),
#     vertex.col = "orange", vertex.border = 0, vertex.cex = 2)
df_stats <- tibble(
    country = network.vertex.names(net),
    indegree = sna::degree(net, gmode = "digraph", cmode = "indegree"),
    outdegree = sna::degree(net, gmode = "digraph", cmode = "outdegree")
)

df_stats <- df_stats |> 
    left_join(select(countries, country = short_name, iso2 = x2_alpha_code))

w <- as.sociomatrix(net, "n") 
diag(w) <- 0

net %e% "w" <- w
net %v% "loops" <- diag(as.sociomatrix(net, "n") )
net %v% "label" <-  df_stats |> pull(iso2)
library(scico)

d <- ggplot(ggnetwork::ggnetwork(net, arrow.gap = 0.005) ,
       aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(aes(color = w), alpha = 1,
               arrow = arrow(length = unit(2, "pt"), type = "closed")) +
    geom_nodes(aes(size = loops), color = "orange", alpha = 1) +
    geom_text(aes(label = label), size = 2, color = "white") +
    scale_color_scico(
        "Shareholders",palette = "berlin", 
        guide = guide_colorbar(barwidth = unit(4,"cm"), barheight = unit(2, "mm"))) +
    scale_size(range = c(4,10), breaks = c(0, 100, 200)) +
    theme_void() + labs(tag = "d") +
    theme(legend.position = "bottom")

df_stats$loops <- net %v% "loops"

f <- df_stats |> 
    mutate(label = case_when(
        indegree > 5 | outdegree > 5 ~ country
    )) |> 
    ggplot(aes(indegree, outdegree)) +
    geom_point(aes(size = loops), alpha = 0.5, show.legend = FALSE) +
    geom_abline(slope = 1, color = "orange", alpha = 0.5) +
    geom_text(
        aes(label = label), nudge_y = -0.5, size = 2) +
    scale_size(range = c(1,10), breaks = c(0, 100, 200)) +
    theme_light() + labs(tag = "c") +
    theme(legend.position = "bottom")

((a/b/f) | d) + plot_layout(widths = c(1,3))

```

A few observations worth noticing: most ownership is either 50 or 100% contrary to what we observed in the emergent zoonotic disease paper. Bear in mind that I did not filter out ownership over time, so if a company A is own 100% by B in 2019, and in 2022 it is own by C just because B change name to C or C is actually a different company (sell / buy), they both currently appear on the data.

Second, most companies have only a few shareholders, just around 10 actually have 5 shareholders or more. Several of the investments are by shareholders on companies in the same countries, on the network these are loops. They are hard to see in the visualization but I scaled node size to correspond to these loops. Norway is the larger country with loops (> 200 shareholders invest in fishing companies inside Norway). For countries where there is actually foreign investments, Spain, UK and France are hot spots of shareholders, while Russia is a hot spot for receiving investments (see panel C).

Below another representation with the loops:

```{r fig.width=8,fig.height=7, dev.args=list(pointsize=6)}
plot.network(
    net, label = net %v% "label", label.col = "white",
    label.pos = 5, label.cex = 1,
    edge.lwd = 0.001, edge.col = alpha("gray75", 0.75),
    vertex.col = "orange", vertex.border = 0, 
    vertex.cex = 2.5)
```

I ran an exponential random graph model on the country network to explore what influences the likelihood of foreign investment using as predictors World Bank statistics of governance and inequality. The result is below, following the same logic of our EZD paper.

![](figures/poster_ergm.png)

The null model (left) assumes that the only process governing link formation is randomness. The second model (centre) assumes all that matter is the difference between governance indicators across connected countries. And the third model (right) assumes that both differences and the individual attributes are important. The best model is the last one, which predicts that links are more likely to form in countries with higher government efficiency, and the difference in political stability is negative (receiving countries are more unstable than investing ones). In fact, politically stable countries are less likely to form links. All other governance attributes (rule of law, corruption, regulatory quality) do not seem to influence link formation or the likelihood that shareholders makes foreign investments on fishing companies exposed to regime shifts.

## Discussion

1. How do you think we can improve the analysis?
2. What other data sets and questions can we ask?

In previous discussions we thought of using Eikon to get info on debt, who fund companies? In a preliminary analysis with Peruvian vessels and companies we did not find anything in Eikon. Given the sparsity of the current dataset, I doubt we will find much, but happy to discuss.

### 2023-08-28 response.Rmd
draft response review

---
title: "Draft response"
author: "Juan Rocha"
date: "2023-08-15"
output:
    html_document:
      theme:
        bootswatch: cosmo
        code_font:
            google: Fira Code
      df_print: paged
      code_folding: hide
      toc: true
      toc_float:
        collapsed: true
        smooth_control: true
      toc_depth: 3
      fig_caption: true
      fig_align: center
      highlight: pygments
      self_contained: false
      lib_dir: libs
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
# library(tictoc)
library(patchwork)
library(here)
library(fs)
# library(DT)
# library(ggnetwork)
# library(network)
```

## Letter to editor

## Response to reviewers

### Reviewer #1:

I am happy that the authors have effectively and reasonably address all my substantive comments, and I am therefore happy to recommend acceptance. I recognise, however, that the other reviewers raised a number of other issues, and that a final acceptance decision will also require their assessment. I believe that this revised version improves significantly on many of the issues identified by the other reviewers.

> Thank you for your feedback

### Reviewer #2:

**General comments.**

Overall, I think the paper has the potential to open a direction of work and I wish the paper can end up being published. I acknowledge that the authors have carried out a substantial revision of the text. However, I am not entirely satisfied with some of the replies and the revisions made so far.

GC1. The contribution to the existing scientific literature is still not very clear (yes, there is no single one to build on, but it does relate to several ones, e.g. EID, anthropocene, networks, corporate finance). What does the paper brings to these exactly?

> We believe our paper empirically links the risk of emergent zoonotic diseases with companies and their shareholders. Our network mapping will help companies and shareholders to be aware of the risk, while informing countries who can make a difference through their regulatory frameworks. To the best of our knowledge, the link between disesase emergence and corporate operations have not been explicitly made before.

GC2. The answer to question a) (Which specific financial institutions could be said to have influence over changes in such risks through their investments?) is that these are the usual suspects that are find in network analyses of corporate ownership. But it is based on 9 case studies. Can we infer that results could hold for EID in general?

> The influencial actors are likely to be the same. Other studies who rank actors by influence using the full network (Vitali and the refs you suggested) find similar actors (BlackRock, Vanguard, etc). In our case the selection of only 9 cases has to do with making sure the operation of companies do actually increase the risk of zoonotic diseases. So we manually curated companies whose operations imply actions that increase EID such as deforestation or changes in land use in areas that have already been reported to have high risk. Without that manual curation we could err in including many companies whose operations are not changing ecosystems or do not happen in areas of EID risk. But because the global ownership network is relatively stable in its metrics of influence (refs you suggested), the ultimate owners are likely to be similar if we include other sectors of the economy (e.g. banana or cacao production).

GC3. The answer to question b) (In what ways could such financial influence be leveraged to help reduce EID risks globally?) is not very clear to me. The authors have now added a literature review on some aspects of corporate finance and shareholders' activism. They could link the analysis to the ongoing debate on ESG (we saw a big growth in the recent years, but currently a backlash, see e.g. recent Blackrock's view). But, net of this, many passages of the paper remain superficial (e.g. "countries who could work together to leverage financial influence as a means to address EID risks."). Do those country have any incentive to act on EID? Are the authors suggesting a policy proposal? At the moment, I am not sure the paper is provides answers to question b. One options is to play down the claim.

> Thanks for your comment. We see our results useful in three ways. First, for companies, our results make them aware that their operations are putting them at risk of EZDs. So companies can choose to ignore, or relocate some of their operations to low risk areas. Second, from the shareholder perspective, it exposes a different type of risk (EZD) that can be of interst to them. Actors like BlackRock or Vanguard can use our results to i) be aware of the investments that expose them to EZD risk, and ii) decide whether ignore or divest from such companies and re-invest on low risk ones. Now, neither companies producing commodities that increase EZD risk nor their shareholders have any mandate to care about EZDs. Countries do, they have the mandate to protect their citizens and tackle potential emergent problems such as a pandemic. Thus third, our results show which countries are most exposed through the investments that originate in their economies and can apply instruments (policies, taxes, incentives) to avoid EZDs risk. Our paper shows at country level where such efforts are more valuable (which pairs of countries benefit from a bilateral agreement for example). Such instruments have been used in the past to reduce deforestation in the Amazon (e.g. Zero deforestation commitments), or impose sactions (e.g. to Russia for the Ukraine invasion). That is why we focus on countries, they have the mandate to prevent the spreading and emergence of epidemics.

GC4. The authors say the paper opens the way to new work, but it is not clear what is the gap to fill and why it is important to fill it. The paper does not seem to qualify as a comprehensive empirical study. So what do we learn from the 9 cases studies? How generalizable are the results? Is the paper providing a methodology to be applied when more data will be available?

> tbd

GC5. I insist that the fact that the a few US funds appear as most important does not seem EID specific. This is not well addressed in the revision

> Agree, it is not specific to EID. But the valuable information is in reverse. Our results tells the US funds of what investments are exposing the economy to EZDs. That was unknown before our paper.

GC6. The metrics used to assess financial influence are very basic. I had suggested to take into account at least the relative weights, but this was not done. See more below.

> The relative weights have been already taken into consideration when we quantify holdings. We do not use the approach of Vitali and Glattfelder because it requires the full network to quantify the indirect paths. We acknowlege their contribution, but we find it inadequate to answer our research question. More details below.

GC7. About the ERG model, I am still not clear what is the claim made exactly and how is the model helping. See more below.

> Thank you for your comment. We concur that the ERGMs are not sufficiently explained in the text. We have expanded the logic on the text and added a figure to help us explain the logic (e.g. why the ERGMs focus on countries not companies)

GC8. About the network analyses, the results seem similar to what would be found with any sample of 54 firms. Again, it is not clear what do we learn about EID and finance. Maybe we learn that when it comes to finance, EID is similar to other environmental issues, such as climate?

> Your observation is probably true at the level of general features of the network structure (e.g degree distribution, or the distribution of ownership). If we take another issue, the big ultimate owners would have some overlap because it would be a snapshot of the same network. However, if BlackRock wants to divest today from companies that increases the risk of EZD, the list of companies would be different than if they want to divest from companies that causes climate change. Here we provide such list for EZDs.

**Detailed comments**

C1. About influence of financial investors. The metrics used (degree and betweenness) are too basic and inadequate. The discretized score ( i.e. = 1 if the investor has shares in the company, etc.) is too coarse. In my view it is necessary to use metrics that take into account the relative weight that each GUO has in each company.

> We agree. Our analysis is not purely binary, most of our networks are weighted. We appreciate your comment and have attempted to clarify the three different types of networks used. We presented degree and betweeneess as descriptive statistics of our data. We can use instead an eigenvector based metric as Vitali or Glattfelder propose. We do however find some problems with their approach. Before explaining the problems, let us clarify the three levels of networks used.

> First, our raw data is a bipartite network of companies and shareholders $B_{ij}$. For each link we have direct and total percent of ownership, operational revenue, global ultimate owner, and number of employees. However, the data from Orbis has many missing values. For example, in our raw data of 99 companies, 2310 shareholders, 2041 GUO and 4551 links, % of ownership is missing for 58% of observations and operational revenue for 59%. The amount of missing data limits the applicability of Vitali's approach.

> Second, we have one mode networks where companies are connected if they share shareholders (weighted by number of shareholders), and shareholder connected by companies (weighted by number of companies). The third type of network is centered on countries: two countries are connected if a shareholder in country $i$ has an investment on a company operating in country $j$. ERGMs only apply to the third type of network, which is also weighted.

I have already pointed to the methodology developed in Vitali ea. 2011. It is not correct that it cannot be used "because it requires the entire network". It could be applied also to a partial network as the authors have here. So please revise that incorrect statement.

> We believe the advantage of Vitali and Glattfelder approaches is that their approximate influence by combining direct ownership with indirect ownership derived from the network. To take advantage of their method, for indirect ownership one needs the indirect pathways from the global untimate owner to the companies at hand. This is not provided by Orbis when looking up individual companies. One would need to recursively walk up the network from the seed company to the GUO and there is no guarantee in any case of recovering all possible paths. We do not have the complete Orbis dataset, we only investigated companies for which we could associate their location and key commodities to places with high risk of EZD. Without the indirect pathways, we cannot take advantage of Vitali's or Glattfelders algorithms.

> We could try to recursively mine and download the data for all shareholders, but we don't think it will add much to our analysis. It will allow us to estimate Glattfelders influence metric, which uses the shorter paths to account for indirect ownership (avoid double counting and problems reported in the past). But we do not think it is an adequate metric of influence for our research question (expanded below).

> We could still estimate the direct portfolio value for each shareholder following eq.5 in their SM, namely multiplying the weighted matrix $W_{ij}$ with the proportion of ownership (with values on the range 0:1) times the operating revenue $v_j$. We find the approach misleading due to missing values of $v_j$ of the Orbis dataset (59% for our sample). So trying to get a sense of the indirect ownership by recursively quering Orbis is unlikely to solve the problem with missing values. For large scale studies like Glattfelder et al, its not problem because they are interested on the emergent features of the network (e.g. bow tie structure or the distribution of ownership or market capitalization); but for a small N study like ours it can be misleading.

I can provide two simpler and maybe more intuitive quantities. The authors could use:
the Control Index defined in Glattfelder ea. 2009 (Eq. 4) which captures the number of companies "effectively controlled" (meant as synonym of "influenced" there) by each investor.

> Thank you for the suggestion and additional references. We find Glattfelder approach useful to map where the money goes. They quantify influece in monetary terms, in their perspective companies with more money are more powerful. For the purpose of their study it is a reasonable assumption; but not for ours. In our study, being able to influence a company by owning some shares means being able to influence their decisions of what to produce, where, and using which practices. Avoiding EZD from a company board perspective means being able to decide whether the soy exported is produced on a deforestation front or a traditional agricultural landscape. Influence in our study happens at the level of company decisions that increase or decrease the risk of EZDs.

Alternatively, they could use the notion of "portfolio value" defined in Glattfelder ea. 2019, eq 5 page 10 of Supp. Material.
In particular, I recommend the authors to address the following empirical questions:
what is the % of the total market capitalization value of the 54 EID relevant companies that are owned by top 5, 10 most influential investors?
They could also make distribution or a Lorenz curve (see the Glattefelder 2009; Vitali 2011). This will give an idea of how many investors it takes to convince about the relevance of EID in order to get some action at the company board level.

>> to do

C2. I recommend to provide some additional descriptive statistics of the sample studied. E.g. What is the % of number and market capitalization of the 54 firms relative to the firms in the same regions and sector? What is the % of activities that the 54 firms derive from the activities deemed to lead to EID risk?

> To answer the first suggestion we need the complete data: all other firms that exist in the region which we have not collected data for. For the latter, Orbis do not provide such data.

C3. Figure 5 A is not very informative. There are no labels and the color code is not very useful (we know the node in the center are those with high degree). Edges are not visible.

> change figure 5

C4. Figure 5 B: I suppose "mean ownership" is the average share of one investor across the 54 companies. Instead of showing the mean the figure could show all the values.

> change for distribution of ownership.

C5. I am not clear about the likelihood statements obtained with the ERG model. Passage paper page 11 "By combining our data with known indicators of "good governance" from the World Bank, and income inequality metrics (Gini) from the World Inequality database, we note that financial institutions seem to avoid investing in companies operating in countries with high corruption and poor regulatory quality, and that government effectiveness increases the likelihood of investments (Figure 6C)."

I still do not understand the statement exactly. In order to estimate that likelihood I presume one would also need the probability of randomly investing in those countries, based on a given number of links and countries/sectors/firms where these links could be assigned. But the sample only contains the 54 EID sensitive firms. So, I miss what is the null model here.

> Thanks for your comment, it points out the lack of clarity in our writing. When one fits an exponential random graph model, one compares the observed networks with random networks typically generated with a Monte Carlo approach. The most basic null model is the Erdos-Reny which is equivalent of only fitting the probability of a link at random (the term edge in Fig 6). More sophisticated models are created based on hypotheses or expectations derived from theory......

C6 Supp info page page 32: The scientific quality of the section 4 is insufficient. I do not understand how to replicate the computations. Examples:
"Formally, we assess if the probability of link y in the observed network Y depends on a network configuration (η_A) e.g., a link or node level parameter, and gA(y) is one if the network configuration is present, or zero otherwise; compared against random graphs with similar features (A)." I suppose one of the features of the nodes would be the governance? The text is not clear. The definition of "network configuration" is not provided. A link level is certainly not a network configuration.
"… compared against random graphs with similar features (A). The summation is over all random graphs' statistics and the parameter κ is a normalising parameters that ensure a proper probability distribution" : are these random graph still among the 54 firms? If so, it is not clear to me what null model this gives. I am happy to be convinced that the model allows to assess likelihood to invest in high governance countries, but I at the moment I fail to see why.

C7: Passage: "Four financial institutions (Dimensional, Vanguard, BlackRock, and the Government of Norway) have ownership in >40 companies in total, and in three of the selected case study regions (North America, East Asia and Europe)." I suppose this is in line with a random allocation, just because those financial institutions have shares in thousands of listed firms.

C8. Passage "While these patterns of ownership are not surprising and the result of a broader shift in asset management globally (49) …" These patterns are, in general, in line with the findings of 47. Then specifically, regarding the three US funds, in line with 49.

C9: "the specific patterns of cross-national ownership depend on the case study region of interest (Figure 4). For example, the importance of investors from Asia and countries like Japan is more prominent in case study regions Southeast Asia compared to Africa and South America." I recommend to expand on this aspect a bit more with some statistics. The charts in Supp Info seem to suggest some regional focus of the investors.

References:

Glattfelder, J.B. and Battiston, S., 2009. Backbone of complex networks of corporations: The flow of control. PHYSICAL REVIEW E Phys Rev E, 80, p.036104.

Glattfelder, J. and Battiston, S., 2019. The architecture of power: Patterns of disruption and stability in the global ownership network. Available at SSRN 3314648.

### 2023-08-28 05-investors.R

@@ -406,6 +406,8 @@ inv_net <- case_df |>
 
 inv_net %v% "outdegree" <- degree(inv_net, gmode = "digraph", cmode = "outdegree")
 inv_net %v% "indegree" <- degree(inv_net, gmode = "digraph", cmode = "indegree")
+## the lines above are unnecessary, the net is bipartite, in R always undirected so in and out degree are ==
+
 inv_net %v% "bigshark" <- ifelse(
     inv_net %v% "outdegree" >= 20,
     inv_net %v% "vertex.names", NA

### 2023-08-28 07-reviewer_requests.R

@@ -1,0 +1,289 @@
+# Extract key investors per case study
+library(tidyverse)
+library(sna)
+library(ggnetwork)
+library(sf)
+library(spData)
+library(scico)
+library(ggnewscale)
+library(ggrepel)
+library(patchwork)
+
+#### load data ####
+load("data/investors_cleaned.RData")
+load("data/casestudies.RData")
+load("data/shr_class.RData")
+
+dat ## companies - shareholders dataset
+df1 # company - commodities dataset
+df0 # case studies classification
+
+dat <- dat |>
+    ## Filtering ownership > 0.01 %: reduces obs from 4609 to 3621
+    #filter(ownership > 0.01) |> 
+    mutate(holdings = ownership * 0.05)
+
+#### Cleaning steps ####
+# Exclude nodes that are not countries:
+no_country <- c("Central Europe", "Eastern Europe", "Middle East", "Southern Europe", "Europe", "Fejér", "Taiwan")
+case_df <-  df1 %>% unique() |> 
+    select(company, commodity, casestudy, country, origin, type) %>% 
+    #group_by(origin, country, casestudy) %>% 
+    ## standardize names to combine later with wgi dataset
+    mutate(
+        origin = case_when(
+            origin == "Croacia" ~ "Croatia",
+            origin == "Hong Kong" ~ "Hong Kong SAR, China" ,
+            origin == "Ivory Coast" ~ "Cote d'Ivoire" ,
+            origin == "New Zeland" ~ "New Zealand",
+            origin == "Russia" ~ "Russian Federation",
+            origin == "Yemen" ~ "Yemen, Rep.",
+            origin == "South Korea" ~ "Korea, Rep." ,
+            origin == "UK" ~ "United Kingdom"  ,
+            origin == "USA" ~ "United States", 
+            origin == "Slovakia" ~ "Slovak Republic" ,
+            origin == "Congo" ~ "Congo, Rep.",
+            origin == "Republic of Sao Tome" ~ "Sao Tome and Principe" ,
+            origin == "Venezuela" ~ "Venezuela, RB" ,
+            origin == "Laos" ~ "Lao PDR",
+            origin == "Bosnia" ~ "Bosnia and Herzegovina" ,
+            origin == "Egypt" ~ "Egypt, Arab Rep.", 
+            origin == "Macedonia" ~ "North Macedonia" , 
+            origin == "Republic of Congo" ~ "Congo, Rep.",
+            TRUE ~ origin),
+        country = case_when(
[…]

### 2023-09-07 response.Rmd
response to reviewers: improving graphs

@@
 output:
+    word_document:
+      df_print: paged
+      code_folding: hide
+      toc: false
     html_document:
@@
 > We agree.
-Our analysis is not purely binary, most of our networks are weighted.
+Our analysis is not purely binary, most of our networks are in fact weighted.
 We appreciate your comment and have attempted to clarify the three different types of networks used.
-We presented degree and betweeneess as descriptive statistics of our data.
+We presented degree and betweeneess as descriptive statistics of our data (on one mode projections).
 We can use instead an eigenvector based metric as Vitali or Glattfelder propose.
@@
 The amount of missing data limits the applicability of Vitali's approach.
+The matrix $B_{ij}$ is printed below weighted by ownership.
+Being a directed bipartite matrix, eigenvector based centrality fails because it is directed and acyclic.
+
+<img src = "figures/sm_ownership_matrix.png">
 
 > Second, we have one mode networks where companies are connected if they share shareholders (weighted by number of shareholders), and shareholder connected by companies (weighted by number of companies).
-The third type of network is centered on countries: two countries are connected if a shareholder in country $i$ has an investment on a company operating in country $j$.
+We can call the one mode networks $A_ij$ for companies, and $S_ij$ for shareholders.
+The degree and betweenness of the original manuscript was computed on these networks, they are not weighted by ownership, so Vitali's or Glattfelder approaches do not apply.
+
+>The third type of network is centered on countries ($C_ij$): two countries are connected if a shareholder in country $i$ has an investment on a company operating in country $j$.
 ERGMs only apply to the third type of network, which is also weighted.
@@
 > We believe the advantage of Vitali and Glattfelder approaches is that their approximate influence by combining direct ownership with indirect ownership derived from the network.
-To take advantage of their method, for indirect ownership one needs the indirect pathways from the global untimate owner to the companies at hand.
+To take advantage of their method, for indirect ownership one needs the indirect pathways from the global untimate owner to the companies at hand -- at least the shorter path.
 This is not provided by Orbis when looking up individual companies.
@@
 We do not have the complete Orbis dataset, we only investigated companies for which we could associate their location and key commodities to places with high risk of EZD.
-Without the indirect pathways, we cannot take advantage of Vitali's or Glattfelders algorithms.
+Without the (shorter) indirect pathways, we cannot take advantage of Vitali's or Glattfelders algorithms.
 
@@
 It will allow us to estimate Glattfelders influence metric, which uses the shorter paths to account for indirect ownership (avoid double counting and problems reported in the past).
-But we do not think it is an adequate metric of influence for our research question (expanded below).
+But we do not think it is an adequate metric of influence for our research question.
+A further explanation is provided below, but a practical reason is missing values.
 
-> We could still estimate the direct portfolio value for each shareholder following eq.5 in their SM, namely multiplying the weighted matrix $W_{ij}$ with the proportion of ownership (with values on the range 0:1) times the operating revenue $v_j$.
+<img src = "figures/fig_reviewer.png">
+
+> We estimated the direct portfolio value for each shareholder following eq.5 in their SM, namely multiplying the weighted matrix $W_{ij}$ with the proportion of ownership (with values on the range 0:1) times the operating revenue $v_j$.
 We find the approach misleading due to missing values of $v_j$ of the Orbis dataset (59% for our sample).
@@
 bow tie structure or the distribution of ownership or market capitalization); but for a small N study like ours it can be misleading.
+It is not discussed in Glattfelder, but if one carefully check their SM, one notices that they do not have either operational revenues for all links, or key variables like market capitalization are only available for 30% or less of the links for the different years they report (and they do have the full data!)
+
+<img src="figures/fig_reviewer2.png">
+
+> We implemented eigenvector based metrics of centrality that take into consideration the weighted structure of the network.
+We report now power as reported in Glattfelder, and alpha centrality, which is a modified version of the power Bonacich centrality suited for directed and weighted networks.
+Thanks very much for the suggestion!
+
+> Bonacich, P. and Lloyd, P. (2001). “Eigenvector-like measures of centrality for asymmetric relations” Social Networks, 23, 191-201.
 
@@
 
+<img src="figures/fig5_key_actors.png">
+
+>
+
 Alternatively, they could use the notion of "portfolio value" defined in Glattfelder ea. 2019, eq 5 page 10 of Supp. Material.
@@
 
->> to do
+> As explained above, the portfolio value $p_i = \sum_{j \in \Gamma (i)} W_{ij} v_j$ can be misleading due to missing values. We in any case calculated it and reported in a new version of Fig 5.
 
@@
 
-> change figure 5
+> We have changed fig 5
 
@@
 
-> change for distribution of ownership.
+> We included the figure with all values in SM, the bipartite matrix with ownership depicted above.
 
@@
 > Thanks for your comment, it points out the lack of clarity in our writing.
+The sample for which there is country data available is 54 companies, 1819 shareholders ($B_ij$ has 1873 nodes), but the network analyzed is on $C_ij$, so we are investigating what features of the countries where shareholder investing, or companies operating affect the likelihood or investing (creating a link).
 When one fits an exponential random graph model, one compares the observed networks with random networks typically generated with a Monte Carlo approach.
-The most basic null model is the Erdos-Reny which is equivalent of only fitting the probability of a link at random (the term edge in Fig 6).
-More sophisticated models are created based on hypotheses or expectations derived from theory......
+The most basic null model is the Erdos-Renyi which is equivalent of only fitting the probability of a link at random (the term edge in Fig 6).
+More sophisticated models are created based on hypotheses or expectations derived from theory.
+In sociology for example to study the formation of friendship or conflicts, one often look at closed or open triangles, or transitive triads in the network lingo.
+But such motifs do not make sense in our investment problem.
+Previous work suggests that features such as corruption or weak regulatory frameworks can increase the likelihood of foreign investment for industries whose activities relate to resource exploitation.
 
@@
 
+> Again apologies for the lack of clarity.
+As mentioned on the response above, the ERGMs were only performed on the third class of networks reported in our paper, namely $C_{ij}$.
+Code for replication of our computations is available at https://github.com/juanrocha/finance_tipping
+
 C7: Passage: "Four financial institutions (Dimensional, Vanguard, BlackRock, and the Government of Norway) have ownership in >40 companies in total, and in three of the selected case study regions (North America, East Asia and Europe)." I suppose this is in line with a random allocation, just because those financial institutions have shares in thousands of listed firms.
+
+> We agree, these large financial actors have a highly diversified portfolio and chances are they will pop up in any case study.
+We believe a key finding is that we identify their investments that increases EZD, it would be quite easy and unexpensive for them to divest on that "handful" of companies (rather small % of their portfolio) to similar companies wihtout the associated EZD risk.
 
@@
 
+>> I didn't understood this comment
+
 C9: "the specific patterns of cross-national ownership depend on the case study region of interest (Figure 4).
@@
 The charts in Supp Info seem to suggest some regional focus of the investors.
+
+>> I believe Paula and Bianca working on this
 

### 2023-09-07 networks.Rmd
Improving network models and prelim graphs for paper
response to reviewers: improving graphs

@@
 date: "Last update `r Sys.Date()`"
-
 output: html_notebook
+editor_options: 
+  chunk_output_type: inline
 ---
@@
 
+dat <- map2(
+    .x = dat, .y=fls, .f = function(x,y) {
+        x$file_name <- y; 
+        return(x)}
+    )
+
 dat <- dat |> 
   bind_rows()
+
+# dat |> 
+#     mutate(file_name = str_remove(file_name, "data/Paulas_files/Financial_actors/")) |> 
+#     select(file_name, company) |> 
+#     unique()
 

### 2023-09-07 02-actors_descriptive_stats.R

@@ -143,6 +143,7 @@ dat |>
 ## but a reasonable approach is to keep the most recent record only; or the max
 ## if reported on the same date
 dat |> 
+    unique() |> 
     group_by(company, shareholder) |> 
     add_count() |> 
     filter(n > 1) |> 
@@ -150,7 +151,8 @@ dat |>
     print(n = 45) 
     
 
-dat <- dat |> 
+dat <-  dat |> 
+    unique() |> 
     mutate(shareholder = str_to_title(shareholder),
            guo_orbis = str_to_title(guo_orbis),
            guo_final = str_to_title(guo_final))|> 
@@ -159,7 +161,7 @@ dat <- dat |>
     mutate(ownership = pmax(direct_percent, total_percent, na.rm = TRUE),
            is_max = (ownership == max(ownership, na.rm = TRUE)))
 
-dat |> filter(count >1) |>arrange(desc(shareholder)) |> print(n = 45)
+dat |> filter(count >1) |> arrange(desc(shareholder)) |> print(n = 45)
 dat |> filter(count == 1, is.na(is_max)) # there is 298 records with no ownership (NA in direct and total %)
 
 dat <- dat |> 
@@ -177,7 +179,7 @@ dat |>
 dat |> filter(company == "Bio Pappel S.a.b") ## It seems that it double counts when accounting for individual persons
 dat |> filter(company == "Th Plantations Berhad") |> #print(n=32)
     filter(direct_percent> 70 | total_percent > 70) |> 
-    select(shareholder) ## It seem32)here was a big move of shares ~70% from one owner to another. Since we don't know who sold, we only keep the most recent data for this case.\
+    select(shareholder) ## It seems there was a big move of shares ~70% from one owner to another. Since we don't know who sold, we only keep the most recent data for this case.\
 ## three
 dat |> filter(company == "Nine Dragons Paper (Holdings) Limited") 
     
@@ -197,6 +199,13 @@ dat <- dat |>
 dat |> 
     ungroup() |> 
     group_by(company, guo_final) |> 
+    summarize(total_own = sum(ownership)) |> 
+    arrange(desc(total_own)) ## Nothing above 100%
+
+## J230906: Adding Alice test
+dat |> 
+    ungroup() |> 
+    group_by(company, shareholder) |> 
     summarize(total_own = sum(ownership)) |> 
     arrange(desc(total_own)) ## Nothing above 100%
 

### 2023-09-07 05-investors.R

@@ -127,6 +127,7 @@ unique(case_df$shareholder) %in% shr_class$shareholder |> sum()
 
 # file seave upon Paula and Bianca request:
 #save(case_df, file = "data/230823_companies-case_study.Rda")
+#write_csv(case_df, file = "data/230823_cases_dataframe.csv")
 
 case_df |> 
     group_by(shareholder, casestudy) |> 

### 2023-09-07 07-reviewer_requests.R

@@ -26,6 +26,8 @@ dat <- dat |>
 #### Cleaning steps ####
 # Exclude nodes that are not countries:
 no_country <- c("Central Europe", "Eastern Europe", "Middle East", "Southern Europe", "Europe", "Fejér", "Taiwan")
+
+
 case_df <-  df1 %>% unique() |> 
     select(company, commodity, casestudy, country, origin, type) %>% 
     #group_by(origin, country, casestudy) %>% 
@@ -98,6 +100,8 @@ case_df
 
 case_df$company [!case_df$company %in% dat$company] |> unique() # 4 in case_df but not in dat
 dat$company[!dat$company %in% case_df$company] |> unique() # 45 in dat but not in case_df
+
+unique(dat$company)[!unique(dat$company) %in% unique(df1$company)]
 
 case_df |> pull(company) |> unique()  # 57 companies
 dat |> pull(company) |> unique() # 99 companies
@@ -191,50 +195,131 @@ bip_net <- case_df |> ungroup() |>
     )) |> #pull(company) |> unique() |> length()
     select(shareholder, company, ownership, holdings) |> unique() |> 
     filter(!is.na(shareholder)) |> 
+    filter(!shareholder %in% c("Public", "Self Owned")) |> 
     network(directed = FALSE, bipartite = TRUE, matrix.type = "edgelist", 
             ignore.eval=FALSE, multiple = FALSE)
 
 bib_mat <- bip_net |> as.sociomatrix.sna()
-dim(bib_mat) # 1821 shareholders, 54 companies
+dim(bib_mat) # 1819 shareholders, 54 companies
 
+## bipartite network with all 99 companies regardless of case study
+
+net2 <- dat |> 
+    select(shareholder, company, ownership, holdings, op_revenue) |> 
+    unique() |> 
+    # check missing values
+    # filter(is.na(op_revenue)) # 1858 / 4540 links
+    filter(!shareholder %in% c("Public", "Self Owned")) |> 
+    mutate(pow = (ownership/100)*op_revenue) |> 
+    igraph::graph_from_data_frame(directed = TRUE)
+    # network(directed = TRUE, bipartite = FALSE, matrix.type = "edgelist", 
+    #         ignore.eval=FALSE, multiple = FALSE)
+
+net2
 
 #### Revisions ####
 
 # reviewer wants eigen based metrics of centrality, some of them easier to calculate in weighted form in igraph
[…]

### 2023-09-07 calculating_hhi.R

@@ -1,0 +1,134 @@
+###### Zoonotic disease risk paper
+###### Computing Herfindahl-Hirschman (HH) index of ownership concentration for companies and case studies
+###### Script written by Alice Dauriach, last updated 2023-09-06
+
+### Preparing
+#install.packages("dplyr")
+library("dplyr")
+
+`%notin%` <- Negate(`%in%`)
+
+setwd("/home/alice/Documents/GEDB/")
+df <- as_tibble(read.csv("230823_cases_dataframe.csv"))
+
+
+### Checking sum of ownership for each company
+test1 <- aggregate(df$ownership, by=list(company=df$company, casestudy=df$casestudy), FUN=sum)
+
+#There seems to be a serious problem with how the ownership figure is calculated. When a shareholder appears twice in Orbis, with a direct ownership figure AND a total ownership figure, our previous method was to take only the larger of these two figures. But here, instead of taking the maximum, we retain both rows with both ownership figures. The "is_max" column is TRUE at every row. This means double counting of many ownership stakes, which leads to the total ownership per company often exceeding 100%, and which leads to a severe bias which overestimates the stake of shareholders which appear twice in Orbis (with a "direct" and a "total" ownership - usually given by different sources of information to Orbis, which simply reports both figures). Please have a look in your other calculations of the ownership stakes in the network to see whether you solve this problem at a later stage.
+
+#I also noticed that some ownership links between shareholders were missing (i.e., the shareholders are listed as having different GUOs when in fact they have the same one). I do a manual check of all the largest shareholders (visual check trying to spot shareholders with similar names or similar share percentages in the same company with different sources of information).
+
+
+### Manual check of all shareholders
+
+#Ballarpur Industries: 108% ownership but I can't find an easy-to-spot duplicate. The main owner is indirect with 49% - keeping only direct owners would lead to a very misleading HH index... 
[…]

### 2023-10-21 Notes.Rmd
notes reading Hobday 2023

@@
 
+*[Hobday et al 2023 Nature](https://doi.org/10.1038/d41586-023-02730-2)*
+
+_"Worryingly, the climate crisis could eventu- ally cause oceans to reach a permanent heat- wave state relative to historical baselines11, and some regions might no longer support certain species and ecosystems.
+The eco- systems that emerge might not operate and respond to warmer waters in ways that can be anticipated12."_
+
 *Carmine 2020 OneEarth*

### 2023-10-23 16-cleaning_orbis.R
identifying additional data from carmine et al that could be useful

@@ -7,6 +7,8 @@ library(patchwork)
 load("data/boats_owners.Rda")
 load("data/marine_shareholders.RData")
 
+load("data/marine_shareholders_missing.RData")
+
 df_boats
 
 owners <- df_boats |>
@@ -156,10 +158,11 @@ guos <- bind_rows(guos)
 revs <- bind_rows(revs)
 
 
-guos$company <- owners[1:2181] # the last two are missing for now
-revs$company <- owners[1:2181]
+guos$company <- guos_missing #owners[1:2181] # the last two are missing for now
+revs$company <- guos_missing #owners[1:2181]
 
-shrs <- map2(shrs, owners[1:2181], function(x,y) {x$company <- y; return(x)})
+shrs <- map2(shrs, guos_missing, # owners[1:2181], 
+             function(x,y) {x$company <- y; return(x)})
 shrs <- bind_rows(shrs)
 
 shrs |> filter(!is.na(name)) |> 
@@ -176,8 +179,37 @@ guos |> filter(is.na(orbis_id)) # 1189
 ## save the data for now and try to recover more from Orbis later with advise from 
 ## the team
 
-
+# first file for the first round of mining data
 save(shrs, revs, owners, guos, file = "data/cleaned_data_230708.Rda")
+# second file for the second round of mining data
+save(shrs, revs, guos, owners, file = "data/cleaned_data_231023.Rda")
+
+# merge them
+# first change names of the second round objects
+guos2 <- guos
+revs2 <- revs
+shrs2 <- shrs
+
+# now load the old objects
+load("data/cleaned_data_230708.Rda")
+
+## merge dropping missing values, you can always recover the missing owners with 
+## owners %in% df-of-interest$company
+guos <- bind_rows(
+    guos |> filter(!is.na(orbis_id)), guos2 |> filter(!is.na(orbis_id)))
+revs <- bind_rows(revs |> filter(!is.na(latest_year)), revs2 |> filter(!is.na(latest_year)))
+shrs <- bind_rows(shrs |> filter(!is.na(name)), shrs2 |> filter(!is.na(name)))
+
+# now save again, file with new date contains the most recent combined.
+
+# recover the list of missing companies:
+df_missing <- tibble(
+    company = guos2 |> filter(is.na(orbis_id)) |> pull(company)
+)
+
+write_csv(df_missing, file = "data/missing_owners.csv")
+
+
 
 ## some initial visualizations
 countries <- readr::read_csv(

### 2023-10-23 17-alternative_names.R

@@ -22,3 +22,28 @@ guos_missing <- guos |>
 
 ### Now with gous_missing run 15-shareholders.R again
 
+
+#### Compare with Carmine's dataset ####
+car_dat <- read_csv(file = "~/Documents/Projects/high-seas-corporations/data/03_output/HS_fishing_vessels_with_corp_actors.csv")
+car_dat <- readxl::read_xlsx(path = "~/Documents/Projects/high-seas-corporations/data/02_processed/corp_info_HS_fishing_vessels.xlsx", sheet = 1)
+
+car_dat |> names()
+
+car_dat |> 
+    select(year, MMSI, best_flag, best_vessel_class, gfw_owner, owner:last_col())
+
+
+
+missing_mmsi <- df_boats |> 
+    filter(is.na(reg_owner)) |> 
+    pull(mmsi)
+
+(missing_mmsi %in% car_dat$MMSI) |> sum() # 2244
+(df_boats |> filter(!is.na(reg_owner)) |> pull(mmsi) %in% car_dat$MMSI) |> sum() #840
+
+# filter only boats that appear on our dataset of vessels fishing on risk areas
+car_dat <- car_dat |> 
+    filter(MMSI %in% missing_mmsi) |> 
+    select(MMSI, owner, gfw_owner)
+
+### Now with car_dat run 15-shareholders.R again

### 2023-12-14 marine_tipping.Rmd
update with new data

@@
 ```{r}
-load("data/cleaned_data_230708.Rda")
+load("data/cleaned_data_231023.Rda")
 load("data/boats_owners.Rda")
@@
         by = c("company" = "reg_owner")
-    ) |> 
+    ) |> filter(!is.na(flag)) |> 
     mutate(flag = str_remove_all(flag, " \\(Int. Register\\)")) |> 

### 2023-12-14 04-ergms.R

@@ -264,9 +264,9 @@ fits <- map2(fits, fit_name, function(x,y) {x$model <- y; return(x)})
 fig5c <- fits |> 
     bind_rows() |> 
     mutate(p_value = case_when(
-        p.value < 0.05 ~ "p < 0.05",
-        p.value > 0.05 & p.value <= 0.1 ~ "p < 0.1",
-        p.value > 0.1 ~ "p > 0.1")) |> 
+        p.value < 0.001 ~ "p < 0.001",
+        p.value > 0.001 & p.value <= 0.05 ~ "p < 0.05",
+        p.value > 0.05 ~ "p > 0.05")) |> 
     mutate(term = case_when(
         term == "edges" ~ "Edges",
         term == "diff.t-h.corruption" ~ "Difference in corruption",
@@ -708,6 +708,25 @@ df_attr <- df_attr |>
 
 shr_net %v% "community" <- as.numeric(df_attr2$comm)
 
+
+### Pull tables for editors:
+case_df |>  # this is the shareholder network
+    group_by(shr_country, country) |> 
+    summarize(n = n()) |> 
+    left_join(df_attr |> select(country = nodes, comp_country = isoa2)) |> 
+    select(shr_country, comp_country, n) |> 
+    filter(!is.na(shr_country)) |> 
+    left_join(df_attr2 |> select(shr_country = nodes, comm)) |> 
+    filter(shr_country == "NO") |> 
+    arrange(desc(n))
+    write_csv(file = "data/table_fig_sm_shareholders.csv")
+
+net_edgelist |> 
+    left_join(df_attr |> select(origin = nodes, comm2)) |> 
+    filter(origin == "Norway") |> 
+    arrange(desc(channel))
+    write_csv(file = "data/table_fig_5_companies-communities.csv")
+
 ## maps
 library(sf)
 library(spData)
@@ -741,7 +760,7 @@ world_map
 #world_map/(fig5b+fig5c) + plot_layout(heights = c(1.5,1))
 ggsave(
     plot = (world_map / fig5b) / fig5c + plot_layout(heights = c(1.5,1,1)),
-    filename = "figures/fig6_shr_net_map.pdf",
+    filename = "figures/fig4_shr_net_map.pdf",
     device = "pdf", width = 7, height = 5, dpi = 400, bg = "white"
 )
 
@@ -845,3 +864,19 @@ ggsave(
     width = 7, height = 5,
[…]

### 2023-12-14 15-shareholders.R

@@ -9,9 +9,12 @@ owners <- df_boats |>
     filter(!is.na(reg_owner)) |> 
     pull(reg_owner) |> unique()
 
+df_missing <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1TWdrM4ybY_9xYRfWcVL2KqBNTfw-Tyat6mEnzSXKCGU/edit#gid=491876521")
+
+
 ## Initialize remote driver
 d <- rsDriver(
-    port = 4440L, browser = "firefox", extraCapabilities = list(acceptInsecureCerts = TRUE)) # should open a chrome
+    port = 4448L, browser = "firefox", extraCapabilities = list(acceptInsecureCerts = TRUE)) # should open a chrome
 remDr <- d[["client"]]
 #remDr$open()
 tic()
@@ -40,8 +43,16 @@ remDr$goBack()
 # of restarting search again 
 # guos_missing comes from script 17
 owners <- guos_missing # only run for second batch, trying to recover companies missed on the first round
+
+# car_dat comes from script 17, additional companies from Carmine et al dataset
+owners <- car_owners
+
+# df_missing comes from google drive, they are orbis ids recovered manually by Bianca
+owners <- df_missing$orbis_id
+
+remDr$refresh()
 tic()
-for (i in 1:length(owners)){ #seq_along(dat$company)
+for (i in 603:length(owners)){ #seq_along(dat$company)
     tic()
     ## Search one company:
     srch <- remDr$findElement("id", "search")
@@ -64,7 +75,8 @@ for (i in 1:length(owners)){ #seq_along(dat$company)
         "xpath",
         '//*[contains(concat( " ", @class, " " ), concat( " ", "name", " " ))]')
     
-    if(length(opts) == 0) next ## If the company is not found, go to next iteration
+    ## If the company is not found, go to next iteration
+    if(length(opts) == 0 |  opts[[1]]$getElementText() == "") next 
     
     opts[[1]]$getElementText()
     opts[[1]]$clickElement() ## need to interveene here manually for first time
@@ -80,7 +92,7 @@ for (i in 1:length(owners)){ #seq_along(dat$company)
     tbls <- html |> html_elements("table") 
     idx <- which(tbls |> html_attr("class") == "ETBL contactInformation" )
     info <- tbls[idx] |> html_table()
-    Sys.sleep(3)
+    Sys.sleep(10)
     ## Extract revenues:
[…]

### 2023-12-14 16-cleaning_orbis.R

@@ -8,7 +8,8 @@ load("data/boats_owners.Rda")
 load("data/marine_shareholders.RData")
 
 load("data/marine_shareholders_missing.RData")
-
+load("data/marine_shareholders_Carmine.RData")
+load("data/marine_shareholders_missing_completed.RData")
 df_boats
 
 owners <- df_boats |>
@@ -158,10 +159,10 @@ guos <- bind_rows(guos)
 revs <- bind_rows(revs)
 
 
-guos$company <- guos_missing #owners[1:2181] # the last two are missing for now
-revs$company <- guos_missing #owners[1:2181]
+guos$company <- owners[1:1387] #owners[1:2181] # the last two are missing for now
+revs$company <- owners[1:1387] #owners[1:2181]
 
-shrs <- map2(shrs, guos_missing, # owners[1:2181], 
+shrs <- map2(shrs,  owners[1:1387], 
              function(x,y) {x$company <- y; return(x)})
 shrs <- bind_rows(shrs)
 
@@ -182,8 +183,9 @@ guos |> filter(is.na(orbis_id)) # 1189
 # first file for the first round of mining data
 save(shrs, revs, owners, guos, file = "data/cleaned_data_230708.Rda")
 # second file for the second round of mining data
-save(shrs, revs, guos, owners, file = "data/cleaned_data_231023.Rda")
+save(shrs, revs, guos, owners, file = "data/cleaned_data_231214.Rda")
 
+## owners is shorter now but it can be recoverd in full length from guos$comapany
 # merge them
 # first change names of the second round objects
 guos2 <- guos
@@ -191,7 +193,7 @@ revs2 <- revs
 shrs2 <- shrs
 
 # now load the old objects
-load("data/cleaned_data_230708.Rda")
+load("data/cleaned_data_231214.Rda")
 
 ## merge dropping missing values, you can always recover the missing owners with 
 ## owners %in% df-of-interest$company
@@ -199,6 +201,7 @@ guos <- bind_rows(
     guos |> filter(!is.na(orbis_id)), guos2 |> filter(!is.na(orbis_id)))
 revs <- bind_rows(revs |> filter(!is.na(latest_year)), revs2 |> filter(!is.na(latest_year)))
 shrs <- bind_rows(shrs |> filter(!is.na(name)), shrs2 |> filter(!is.na(name)))
+
 
 # now save again, file with new date contains the most recent combined.
 

### 2023-12-14 17-alternative_names.R

@@ -30,7 +30,7 @@ car_dat <- readxl::read_xlsx(path = "~/Documents/Projects/high-seas-corporations
 car_dat |> names()
 
 car_dat |> 
-    select(year, MMSI, best_flag, best_vessel_class, gfw_owner, owner:last_col())
+    select(year, MMSI, best_flag, gfw_owner, owner:last_col())
 
 
 
@@ -47,3 +47,15 @@ car_dat <- car_dat |>
     select(MMSI, owner, gfw_owner)
 
 ### Now with car_dat run 15-shareholders.R again
+car_owners <- car_dat |> 
+    select(owner) |> 
+    unique() |> 
+    filter(owner != 'n/a') |> 
+    mutate(owner = str_remove(owner, pattern = "\\(.*")) |> 
+    mutate(owner = str_remove(
+        owner, pattern = "LTD|LTDA|IMP|EXP|LLC|S\\.L\\.|YK$|\\.\\,|\\.\\, \\.|\\,$|\\, \\.$")) |>
+    mutate(owner = str_trim(owner, "both")) |> unique() |> #print(n=100)
+    pull(owner) 
+
+car_dat |> 
+    filter(str_detect(owner, "OKAY"))

### 2024-02-27 Notes.Rmd
update with new data
marine tipping draft

@@
 # Notes
+
+**240220**
+
+- All data from Carmine is now added to the database.
+Bianca manually annotated vessels for whom we could not find orbis ID automatically (N=411).
+This added observations to our GUO, revenues and shareholders.
+However, from Orbis although the company exist, there is still missing 374 companies shareholder data, 286 companies revenues data, and 143 companies GUOs data.
+All data is now combined and saved in `data/cleaned_data_231214.Rda`
+
+**240216**
+
+- A sample of ~10k mine names were downloaded from Orbis over Xmas break.
+That took ~10days of computing time.
+The problem is many mines are not easily matchable with records in Orbis.
+For a few of them one can be sure due to the economic activity field in Orbis (but not always recorded), or because the mine name contains the product they exploit: diamonds, gold, etc.
+However, for many of them we don't know if the company info downloaded is actually a mine.
+- Idea: one could circumvent this by hard coding the verification: it is mine if the econ activity is recorded, or if the name contains the metal.
+Other options are if the company is based on the country where the coordinates from Maus are recorded...
+buiding up a confidence score or something similar.
+Once solved, we can go ahead and download more.
+Bear in mind it takes long time to do so.
+- Marine paper: I need to download the missing Carmine companies and then merge with complete dataset, so scripts 15 and 16 with missing data corrected by Bianca.
+
+**231213**
+
+- Discussion? stranded assets - link to literature and who is exposed to them
+- High net worth individuals: merging the marine data, EID, fossil fuels, mining.
 

### 2024-02-27 marine_tipping.Rmd
marine tipping draft

@@
 
-Then I used Lloyds (thanks Frida!) to identify owner companies.
-Out of the 16 878 vessels, I got companies for 2183 of them.
-Using Orbis I found 793 shareholders for some of the companies.
-Already now you can see part of our problem: each step we try to go on the chain (vessel - owner - shareholder) we lose observations (one order of magnitude), thus to what extend are results robust?
-Let's check them first.
-
-## Results
-
 ```{r}
-load("data/cleaned_data_231023.Rda")
+load("data/cleaned_data_231214.Rda")
 load("data/boats_owners.Rda")
@@
 
+Then I used Lloyds (thanks Frida!) to identify owner companies.
+Out of the 16 878 vessels, I got `r length(guos)` owners, some of them individual people, others companies.
+Using Orbis I found `r shrs$company |> unique() |> length()` companies and `r shrs$name |> unique() |> length()` unique shareholders with `r nrow(shrs)` investment links to the companies.
+This is after playing around with different strategies to complement the dataset including the data published by Carmine.
+
+## Results
+
 Most boats retrieved are indeed fishing boats:
@@
     geom_bar() +
-    coord_flip()
+    coord_flip() +
+    labs(x = "Vessel type", y = "Count", caption = "Data source: Lloyds")
 ```
@@
 ```{r}
-options(DT.options = list(pageLength = 25))
+options(DT.options = list(pageLength = 10))
 datatable(guos)
@@
 
+# ggsave(filename = "marine_stats.png", path = "figures/",
+#        plot = a + b + c & theme_light(base_size = 6),
+#        device = "png", width = 7, height = 3)
 ```
@@
     filter(!is.na(name)) |> 
-    slice(-338) |>  # problems: the name is the IMO and the shareholder name the same
+    filter(name != company) |>  
+    unique() |> 
+    # problems: the name is the IMO and the shareholder name the same
     network(directed = FALSE, bipartite = TRUE, matrix.type = "edgelist", 
@@
        aes(x = x, y = y, xend = xend, yend = yend)) +
-    geom_edges(aes(color = ownership_total), size = 0.35, alpha = 1
+    geom_edges(aes(color = ownership_total), linewidth = 0.35, alpha = 1
                #arrow = arrow(length = unit(2, "pt"), type = "closed")
@@
 The data I'm using to build the network does not distinguish ownership over time, nor name changing.
-The density of the bipartite network is `r round(network.density(bip),4)`, meaning out of all possible links on the network < 0.1% are realized.
+The density of the bipartite network is `r network.density(bip)`, meaning out of all possible links on the network < 0.001% are realized.
 
 The bipartite network has two one mode projections: a network of shareholders connected to each other if they share investments to the same fishing companies (left), and a network of fishing companies connected to each other if they are invested by the same shareholders (right).
-As you can see, these networks are also very sparse with very few connections and no "gigant key stone actors" emerging from the data.
+As you can see, these networks are also very sparse with very few connections and no "giant key stone actors" emerging from the data.
 
@@
     #scale_x_continuous(labels = scales::label_percent()) +
-    theme_light()
+    theme_light(base_size = 8)
 
@@
     labs(x = 'Number of shareholders', y = "Companies", tag = "b") +
-    theme_light()
+    theme_light(base_size = 8)
 
@@
        aes(x = x, y = y, xend = xend, yend = yend)) +
-    geom_edges(aes(color = w), size = 0.15, alpha = .5
+    geom_edges(aes(color = w), linewidth = 0.15, alpha = .5
                #arrow = arrow(length = unit(2, "pt"), type = "closed")
@@
     scale_size(range = c(4,10), breaks = c(0, 100, 200)) +
-    theme_void() + labs(tag = "d") +
+    theme_void(base_size = 7) + labs(tag = "d") +
     theme(legend.position = "bottom")
@@
     scale_size(range = c(1,10), breaks = c(0, 100, 200)) +
-    theme_light() + labs(tag = "c") +
+    theme_light(base_size = 7) + labs(tag = "c") +
     theme(legend.position = "bottom")
@@
 ((a/b/f) | d) + plot_layout(widths = c(1,3))
+
+ggsave(
+    path = "figures/", filename = "marine_nets.png", device = "png",
+    width = 7, height = 5, dpi = 500,
+    plot = ((a/b/f) | d) + plot_layout(widths = c(1,3))
+)
 
@@
 
-Second, most companies have only a few shareholders, just around 10 actually have 5 shareholders or more.
+Second, most companies have only a few shareholders.
 Several of the investments are by shareholders on companies in the same countries, on the network these are loops.
@@
 
+**Global Ultimate Owners**
+
+```{r fig.width=8, fig.height=8}
+load("data/cleaned_data_231214.Rda")
+
+guos <- guos |>
+    mutate(class = case_when(
+        str_detect(guo, "Public") ~ "Public",
+        str_detect(guo, "Private") ~ "Private",
+        str_detect(guo, "Formerly") ~ "Former public",
+        .default = guo
+    )) |> # pull(class) |> table()
+    mutate(
+        ## remove all the cases
+        guo = str_remove(guo, "Private The Global Ultimate Owner of this controlled subsidiary is "), 
+        guo = str_remove(guo, "Formerly publicly quoted The Global Ultimate Owner of this controlled subsidiary is "),
+        guo = str_remove(guo, "Private This entity is a marine vessel of "),
+        guo = str_remove(guo, "Publicly quoted The Global Ultimate Owner of this controlled subsidiary is "),
+        guo = str_remove(guo, "Private This entity was a branch of "),
+        guo = str_remove(guo, "Private This entity is a branch of "), 
+        guo = str_remove(guo, "Private This entity is a foreign company of "),
+        ) |> #pull(guo) |> unique() #|> 
+    mutate(
+        guo = case_when(
+            guo =="Private This entity is a marine vessel." ~ NA,
+            guo == "Private The Global Ultimate Owner of this company could not be identified."  ~ NA,
+            guo == "Private This entity is a branch." ~ NA,
+            guo == "Publicly quoted This company is Independent (but not the Global Ultimate Owner of a Corporate Group)"  ~ NA,
+            guo == "Publicly quoted This company is a Single location" ~ NA, 
+            guo == "Private" ~ NA,
+            guo == "Private This entity is a foreign company." ~ NA,
+            guo == "Private This company is Independent (but not the Global Ultimate Owner of a Corporate Group)" ~NA,
+            guo == "Publicly quoted This company is the Global Ultimate Owner of the corporate group"  ~ orbis_name,
+            guo == "Private This company is the Global Ultimate Owner of the corporate group"  ~ orbis_name,
+            guo == "Formerly publicly quoted This company is the Global Ultimate Owner of the corporate group" ~ orbis_name,
+            guo == "Private This company is a Single location" ~ NA,
+            .default = guo
+        )
+    )  |> 
+    mutate(person = str_starts(guo, pattern = "MR|MRS|MS")
+    )
+
+a <- guos |> 
+    ggplot(aes(class)) +
+    geom_bar() +
+    coord_flip()
+
+b <- guos |> 
+    ggplot(aes(person)) + 
+    geom_bar() +
+    coord_flip()
+
+c <- guos |> 
+    filter(!is.na(guo), person == FALSE) |> 
+    group_by(guo) |> 
+    summarize(n = n()) |>
+    arrange(n) |> 
+    slice_tail(n = 25) |> 
+    mutate(guo = as_factor(guo)) |> 
+    ggplot(aes(n, guo)) +
+    geom_col()
+
+(a + b) / c & theme_light(base_size = 8)
+
+```
+
 ## Discussion
 
-1. How do you think we can improve the analysis?
-2. What other data sets and questions can we ask?
+1. What to do with GUOs? Do we include them on the analysis?
 
@@
 Given the sparsity of the current dataset, I doubt we will find much, but happy to discuss.
+I also tried with the Datastream (part of Refinitiv) platform thanks to IVL access.
+I did not find much info on debt, the companies are too small / unknown.

### 2024-02-27 paper/manuscript.Rmd
marine tipping draft

--- 2023-02-23 manuscript.Rmd
@@
 author: |
-    | \small Juan C. Rocha^1^, Jean-Baptiste Jouffray^1^, Paula A. Sánchez^1,2^, Victor Galaz^1^
+    | \small Juan C. Rocha^1^, Jean-Baptiste Jouffray^1^, Frida Bengtsson^1^, Bianca-Ioana Voicu^1^,  Paula A. Sánchez^1,2^, Victor Galaz^1^
     | \footnotesize ^1^Stockholm Resilience Centre, Stockholm University
@@
 csl: /Users/juanrocha/Documents/styles/apa.csl
-#bibliography: refs.bib
+bibliography: refs.bib
 citecolor: blue
@@
 - Author order tbd
-- Target journal: Nature family?, GEC, Finance journal? ~ 4000w, 4 figs, 40refs
-- Word budget
+- Target journal: Nature family?, SciAdv, GEC, Finance journal? ~ 4000w, 4 figs, 40refs
+
+## Outline
+
 - Intro: 1000w
+    -  Marine ecosystems are expected to undergo large biological reorganizations
+    - Exposure to regime shifts are mediated by pressures from climate change, fishing pressure, or pollution.
+    - Companies and financial actors are part of the problem but can also be part of the solutions.
+    - RQ: **Which financial actors are exposed to regime shift risks, and thus have the agency to avoid them or can play a role in navigating transformations?**
 - Method: 800w
+    - Identifying areas prone to regime shifts:
+        - High probability of marine heatwaves
+        - High probability or regime shifts, singals of proximity to tipping points
+    - Identify vessels who fish in these areas
+    - Identify companies who operate these vessels (owners). Identify shareholders who have shares on these companies. Identify global ultimate owners. 
+    - Network statistical models to explore what influence the likelihood of investment.
 - Results: 1000w
+    - The network of ownership and shareholder investments is sparse lacking a backbone or common financial giants
+    - Despite the sparsity of the ownership network, there are important flows of investments across national borders.
+    - Probability of investments: ergms
 - Discussion: 800w
@@
 
+- Identifying areas prone to regime shifts:
+    - High probability of marine heatwaves
+    - High probability or regime shifts, singals of proximity to tipping points
+- Identify vessels who fish in these areas
+- Identify companies who operate these vessels (owners). Identify shareholders who have shares on these companies. Identify global ultimate owners. 
+- Network statistical models to explore what influence the likelihood of investment.
+
 _Case studies:_ To that end, we first identified areas of the oceans exposed to potential regime shifts.
@@
 _Companies_: Based on the geographic exposure to regime shifts, we then identified fishing companies operating in these regions.
-We used permits and licences issued in Alaska by the National Oceanic and Atmospheric Administration in the United States ([NOAA](https://www.fisheries.noaa.gov/alaska/commercial-fishing/permits-and-licenses-issued-alaska)) to list companies with permits to fish Alaskan pollock, one of the most profitable resources in the region.
-We used fishing permits issued by the _Servicio nacional de Pesca y Acuicultura_ or the National Fishing and Aquaculture authority of Chile, to companies with the purpose of harvesting anchovies and sardines.
-
-- check with Paula, we should have Peru as well
-- what to do with tuna? we only have O
 
@@
 
+- The network of ownership and shareholder investments is sparse lacking a backbone or common financial giants (SM nets figures).
+    - contrary to previous results of SeaBOS, and financial giants: why?
+Ownership structure: few owners who own 1/2 or the entirety of the company, many family and small scale businesses.
+Market conglomeration probably emerge at the level of intermediaries, but from the data, that is not the case for the captains who make decisions on where to fish.
+    - Yet, most vessels are Chinesse, followed by Japan and USA. Most shareholders are based in Norway but they do invest mostly on Norwegian fishing companies.
+
+- Despite the sparsity of the ownership network, there are important flows of investments across national borders. 
+    - International investments are dominated by Spain, UK and France as countries where investments originate, while Russia is a country who disproportionately receives investments.
+
 # Discussion
 
+- Contrast to other papers of key stone actors: fisheries does not seem to be highly concentrated, at least not at the level of decisions that expose to risk of regime shifts -- where to go fishing and when.
+
 - Limitation: publicly listed companies, and legal fisheries.
+    - We do not take care of timing: a company can change owners during the time period.
+We do not control for it because chances are the previous owners reinvest in fishing related activities, there is no way to track investments.
 
@@
 ## Acknowledgements
+
+Grants XYZ
 

### 2024-02-27 paper/manuscript.qmd
marine tipping draft

---
title: "Identifying financial actors exposed to marine tipping points"
format: pdf
editor: visual
date: "2024-02-27"
author: 
  - name: Juan C. Rocha
    orcid: 0000-0003-2322-5459
    email: juan.rocha@su.se
    affiliation:
      - name: Stockholm Resilience Centre, Stockholm University
abstract: >
  Here is the abstract
  multiple lines
keywords:
  - Tipping points
  - Financial actors
  - Regime shifts
---

## Outline

-   Introduction \[1000w\]

-   Method \[800w\]

-   Results \[1000w\]

-   Discussion \[800w\]

-   Conclusion \[400w\]

## Quarto

Quarto enables you to weave together content and executable code into a finished document.
To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code.
You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

### 2024-02-27 04-ergms.R

@@ -727,6 +727,10 @@ net_edgelist |>
     arrange(desc(channel))
     write_csv(file = "data/table_fig_5_companies-communities.csv")
 
+df_attr |> 
+    select(nodes, starts_with("comm")) |> 
+    write_csv("data/table_for_bianca.csv")
+
 ## maps
 library(sf)
 library(spData)

### 2024-02-27 15-shareholders.R

@@ -9,17 +9,20 @@ owners <- df_boats |>
     filter(!is.na(reg_owner)) |> 
     pull(reg_owner) |> unique()
 
-df_missing <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1TWdrM4ybY_9xYRfWcVL2KqBNTfw-Tyat6mEnzSXKCGU/edit#gid=491876521")
+df_missing <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1TWdrM4ybY_9xYRfWcVL2KqBNTfw-Tyat6mEnzSXKCGU/edit#gid=322700622", sheet = 2)
+
+df_missing <- df_missing |> 
+    filter(!is.na(OrbisID))
 
 
 ## Initialize remote driver
 d <- rsDriver(
-    port = 4448L, browser = "firefox", extraCapabilities = list(acceptInsecureCerts = TRUE)) # should open a chrome
+    port = 4444L, browser = "firefox", extraCapabilities = list(acceptInsecureCerts = TRUE)) # should open a chrome
 remDr <- d[["client"]]
 #remDr$open()
 tic()
 remDr$navigate(url = "https://ezp.sub.su.se/login?url=https://orbis4.bvdinfo.com/ip")
-toc() # >40s
+toc() # >6s
 ## Note: working with VPN skips authentification steps! but takes longer to load
 Sys.sleep(30)
 
@@ -48,11 +51,14 @@ owners <- guos_missing # only run for second batch, trying to recover companies 
 owners <- car_owners
 
 # df_missing comes from google drive, they are orbis ids recovered manually by Bianca
-owners <- df_missing$orbis_id
+owners <- df_missing$OrbisID
+
+# df for mining: 29k
+owners <- mines
 
 remDr$refresh()
-tic()
-for (i in 603:length(owners)){ #seq_along(dat$company)
+
+for (i in 411:length(owners)){ #seq_along(dat$company)
     tic()
     ## Search one company:
     srch <- remDr$findElement("id", "search")
@@ -76,7 +82,7 @@ for (i in 603:length(owners)){ #seq_along(dat$company)
         '//*[contains(concat( " ", @class, " " ), concat( " ", "name", " " ))]')
     
     ## If the company is not found, go to next iteration
-    if(length(opts) == 0 |  opts[[1]]$getElementText() == "") next 
+    if(length(opts) == 0) next 
     
     opts[[1]]$getElementText()
[…]

### 2024-02-27 16-cleaning_orbis.R

@@ -10,6 +10,7 @@ load("data/marine_shareholders.RData")
 load("data/marine_shareholders_missing.RData")
 load("data/marine_shareholders_Carmine.RData")
 load("data/marine_shareholders_missing_completed.RData")
+load("data/carmine_shareholders.RData") # this is the missing values of carmine manually coded by Bianca
 df_boats
 
 owners <- df_boats |>
@@ -158,11 +159,11 @@ toc() #25s
 guos <- bind_rows(guos)
 revs <- bind_rows(revs)
 
+# 411 is the number of owners recovered from Carmine (annotated by Bianca)
+guos$company <- owners[1:411] #owners[1:2181] # the last two are missing for now
+revs$company <- owners[1:411] #owners[1:2181]
 
-guos$company <- owners[1:1387] #owners[1:2181] # the last two are missing for now
-revs$company <- owners[1:1387] #owners[1:2181]
-
-shrs <- map2(shrs,  owners[1:1387], 
+shrs <- map2(shrs,  owners[1:411], 
              function(x,y) {x$company <- y; return(x)})
 shrs <- bind_rows(shrs)
 
@@ -207,10 +208,10 @@ shrs <- bind_rows(shrs |> filter(!is.na(name)), shrs2 |> filter(!is.na(name)))
 
 # recover the list of missing companies:
 df_missing <- tibble(
-    company = guos2 |> filter(is.na(orbis_id)) |> pull(company)
+    company = guos |> filter(is.na(orbis_id)) |> pull(company)
 )
 
-write_csv(df_missing, file = "data/missing_owners.csv")
+write_csv(df_missing, file = "data/missing_owners_carmine.csv")
 
 
 
@@ -344,6 +345,7 @@ library(network)
 bip <- shrs |> 
     select(name, company, ownership_total) |> 
     filter(!is.na(name)) |> 
+    #filter(name != company) |> 
     slice(-338) |>  # problems: the name is the IMO and the shareholder name the same
     network(directed = FALSE, bipartite = TRUE, matrix.type = "edgelist", 
             ignore.eval=FALSE, loops = TRUE)

### 2024-02-27 20-mines.R

@@ -1,0 +1,38 @@
+library(tidyverse)
+library(sf)
+
+dat <- read_sf(dsn = "data/mines_sp_old.gpkg")
+
+
+dat |> 
+    ggplot() +
+    geom_sf()
+
+dat # 34 820 obs
+
+## Option 1: download all mines and figure out intersecitons with resilience loss
+## or loss of functional integrity later.
+## Opetion 2: do the intersection first and only download the mines of interest.
+## If Option 1, we can also study a bit the issue of concentration.
+
+dat |> pull(mine) |> unique() |> length() # 29 718 mines
+dat |> pull(known_as) |> unique() |> length() # 18 268 aka.
+
+## reduce dataset to unique mines:
+df_dat <- dat |> 
+    as_tibble() |> 
+    select(mine, known_as) |> 
+    unique() 
+
+df_dat |> print(n=200)
+df_dat <- df_dat |> filter(!is.na(mine))
+mines <- df_dat |>  pull(mine) |> unique() 
+
+#### Orbis ####
+
+# Orbis script run, data object:
+load("data/mining_shareholders.RData")
+
+#### Explore shareholders ####
+
+guo_list[[1]]

### 2024-11-15 11-cleaning_data.R
moving all files of the marine paper to its own repo

@@ -1,84 +1,0 @@
-library(tidyverse)
-
-
-load("data/bolivia_orbis.RData")
-dat <- read_csv("data/fishing_companies_alaskan-pollock.csv")
-
-dat
-
-## Recovering IDs
-guo_list[[1]][1][[1]][2,1]
-
-dat$ids <-  guo_list |> 
-    map(function(x) x[[1]][[1]][2]) |> 
-    map(function(x) str_remove_all(x,pattern = "\n|\t|Orbis ID: ")) |> 
-    map(function(x) str_remove_all(x, "ActiveBvD ID: |Active \\(dormant\\)BvD ID: "))
-
-dat <- dat |> 
-    unnest(ids, keep_empty = TRUE) |> 
-    mutate(ids = str_replace(ids, "[:space:]", "_")) |> 
-    separate(col = ids, into = c("BvD_id", "orbis_id"), sep = "_")
-
-## We also have company address if needed
-## Recover GUO:
-
-guo_list[[1]][[1]][2,3]
-
-dat$guo <- guo_list |> 
-    map(function(x) x[[1]][2,3]) |> 
-    # recover GUOs
-    map(function(x) str_remove_all(x, pattern = "Private The Global Ultimate Owner of this controlled subsidiary is ")) |> 
-    map(function(x) str_replace(x, "Private This company is the Global Ultimate Owner of the corporate group", "self-GUO")) |> 
-    map(function(x) str_replace_all(x, "Private This company is Independent \\(but not the Global Ultimate Owner of a Corporate Group\\)", "independent")) |> 
-    map(function(x) str_replace_all(x, "Private This company is a Single location", "single")) |> 
-    map(function(x) str_remove_all(x, "Private This entity is a branch of "))
-
-dat <- dat |> 
-    unnest(guo, keep_empty = TRUE)
-
-## Recover revenues
-revenues[[1]][[1]][1,2] # units
-revenues[[1]][[1]][3,2] # revenue
-revenues[[1]][[1]][4,2] # employees
-
-dat$rev_units <- revenues |> map(function(x) x[[1]][2][[1]][1])
-dat$revenue <- revenues |> map(function(x) x[[1]][2][[1]][3])
-dat$employees <- revenues |> map(function(x) x[[1]][2][[1]][4])
-
-dat <- dat |> 
-    unnest(cols = rev_units:employees, keep_empty = TRUE) |> 
[…]

### 2024-11-15 12-GFW_data.R

@@ -1,135 +1,0 @@
-## Exploring the global fish watch data
-
-library(tidyverse)
-library(here)
-library(fs)
-library(sf)
-library(raster)
-
-
-dat <- read_csv("~/Documents/Projects/DATA/GlobalFishWatch/9c3d6bc0-a3c1-11ed-a61d-7ba19957371b/fishing-vessels-v2.csv")
-load("data/mhw_exposure.Rda")
-load("~/Documents/Projects/ESDL_earlyadopter/ESDL/Results/210301_delta_detected_ChlorA_log.RData")
-
-dat |> names()
-dat |> skimr::skim()
-dat |> pull(registries_listed) |> unique()
-dat |> pull(vessel_class_registry) |> unique() ## identifies tuna as target
-dat
-
-## create boxes for each pixel. 
-pxls <- pxls |> 
-    mutate(xmin = lon - 0.125, xmax = lon + 0.125, ymin = lat - 0.125, ymax = lat + 0.125)
-pxls
-
-pxls |> 
-    ggplot(aes(lon, lat)) +
-    geom_tile(aes(fill = mp_severe))
-
-## Create a mask with pixels where boats IDs need to be recovered.
-## they are in different projections
-
-df_delta_detected |> 
-    filter(n_ews >=3) |> 
-    ggplot(aes(lon,lat)) +
-    geom_tile(aes(fill=n_ews))
-
-# pxls |> 
-#     full_join(df_delta_detected |> 
-#                   filter(n_ews >= 3) |> 
-#                   select(lon, lat, n_ews))
-## This is to get the original CRS from the NOAA SST data
-fls <- dir_ls("~/Documents/Projects/DATA/OISST/")
-sst <- raster(fls[1], varname = "sst")
-crs(sst)
-x <- st_crs("WGS84")
-
-deltas <- rasterFromXYZ(df_delta_detected |> 
-                            ungroup() |> 
-                            dplyr::select(lon, lat, n_ews), 
-                        crs = x$proj4string)
-
-pxls2 <- rasterFromXYZ(pxls |> dplyr::select(lon,lat, mp_extreme, mp_severe),
-                       crs = crs(sst)) |> 
-    # rotate changes from 0-360 longitude to -180:180, compatible with other data
-    rotate()
-
-deltas <- extend(deltas, pxls2)
-
-plot(deltas)
[…]

### 2024-11-15 13-Lloyds.R

@@ -1,101 +1,0 @@
-library(tidyverse)
-library(RSelenium)
-library(keyring)
-library(rvest)
-library(tictoc)
-
-
-load("data/boats.Rda")
-#load("data/vessels/vessels-6001-7000.RData")
-df_boats
-
-
-## Initialize remote driver
-d <- rsDriver(
-    port = 4445L, browser = "firefox") # should open a chrome
-remDr <- d[["client"]]
-#remDr$open()
-tic()
-remDr$navigate(url = "https://www.seasearcher.com")
-toc() #
-
-btn <- remDr$findElement("xpath",'//*[contains(concat( " ", @class, " " ), concat( " ", "lli-flex-row--justify-centered", " " ))]')
-btn$clickElement()
-
-usr <- remDr$findElement("id", "thePage:siteTemplate:j_id27:username")
-usr$sendKeysToElement(list("frida.bengtsson@su.se"))
-
-
-pwd <- remDr$findElement("id", 'thePage:siteTemplate:j_id27:password')
-pwd$sendKeysToElement(list(keyring::key_get("maritime", "frida.bengtsson@su.se"), key = "enter"))
-Sys.sleep(6)
-
-btn <- remDr$findElement("css", '.lli-nav__link')
-btn$clickElement()
-
-# owners <- list()
-# features <- list()
-
-remDr$refresh()
-Sys.sleep(5)
-
-j = 16000# last iteration in the last saved file.
-
-tic()
-for (i in 761:875){
-    
-    srch <- remDr$findElement("class", "lli-searchform__input")
-    srch$clickElement()
-    # delete any left over from last iteration
-    walk(1:9, ~srch$sendKeysToElement(list(key = "backspace")))
-    
-    srch$sendKeysToElement(list(df_boats$boats[[i + j]] |> as.character()))
-    # gets the options from the list that appears on the search pane
-    Sys.sleep(18)
-    
-    opts <- remDr$findElement("class","lli-typeahead-options__option")
-    
-    ## Test if the vessel is found, if not go to next
-    if("No Results Found" == opts$getElementText() |> unlist()) {
-        walk(1:9, ~srch$sendKeysToElement(list(key = "backspace")))
-        next
-        }
-           
-    opts$clickElement() # click the first one)
-
-    Sys.sleep(15)
-    
-    ## Get html of all info on vessel:
-    html <- remDr$getPageSource() |> 
-        unlist() |> 
-        read_html()
[…]

### 2024-11-15 14-cleaning_lloyds.R

@@ -1,103 +1,0 @@
-library(tidyverse)
-library(fs)
-library(tictoc)
-
-fls <- dir_ls("data/vessels/")
-fls
-
-load(fls[1])
-load("data/boats.Rda")
-# Each file has features, and owners
-
-head(features)
-
-features[[1]] |> print(n=56)
-features |> map(.f = nrow) |> unlist() |> table()
-features[[1000]] |> print(n=120)
-
-## Features to extract:
-## LLI NO: Lloyds ID; IMO; Flag; LLI Vessel Type; Status; Reg. Owner; built; Name;
-## MMSI; 
-
-key_features <- c("LLI NO:", "IMO:", "Flag:", "LLI Vessel Type:", "Status:", 
-                  "Reg. Owner:", "Built:", "Name:", 'MMSI:')
-
-
-## Start for loop here
-out <- list() # output list
-tic()
-for (i in seq_along(fls)){
-    load(fls[i])
-    ## select features
-    features <- features |> 
-        map(.f = function(x) if(!is.null(x)) filter(x, feature %in% key_features))
-    # extract the null queries
-    not_ok <- map(features, is.null) |> unlist()
-    empty_df <- map(features, nrow) |> unlist() == 0
-    tic()
-    out[[i]]  <- features[!not_ok][!empty_df] |> #head() |> 
-        map(.f = function(x) {
-            x |> unique() |> # remove duplicated MMSI or IMOs
-                ## remove the duplication of flag
-                mutate(values = str_remove_all(
-                    values, pattern = "Since: [:alnum:]* \\d{2} \\w{3} \\d{4}")) |>
-                mutate(values = str_remove_all(
-                    values, pattern = "Since: \\d{2} \\w{3} \\d{4}")) |> 
-                mutate(values = str_remove_all(
-                    values, pattern = "Since: In the [:alnum:]* of \\d{2} \\w{3} \\d{4}")) |> 
-                unique() |> 
-                # makes missing values explicit
-                mutate(values = case_when(values == "" ~ NA, TRUE ~ values)) |> 
-                pivot_wider(names_from = feature, values_from = values) |> 
-                janitor::clean_names()
-        }) |> bind_rows()
-    print(glue::glue("File ", i, " completed, it took: \n"))
-    toc() #24s
-}
-toc() # 396s | 6.7mins
-
[…]

### 2024-11-15 15-shareholders.R

@@ -1,188 +1,0 @@
-library(tidyverse)
-library(RSelenium)
-library(keyring)
-library(rvest)
-library(tictoc)
-
-load("data/boats_owners.Rda")
-owners <- df_boats |>
-    filter(!is.na(reg_owner)) |> 
-    pull(reg_owner) |> unique()
-
-df_missing <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1TWdrM4ybY_9xYRfWcVL2KqBNTfw-Tyat6mEnzSXKCGU/edit#gid=322700622", sheet = 2)
-
-df_missing <- df_missing |> 
-    filter(!is.na(OrbisID))
-
-
-## Initialize remote driver
-d <- rsDriver(
-    port = 4444L, browser = "firefox", extraCapabilities = list(acceptInsecureCerts = TRUE)) # should open a chrome
-remDr <- d[["client"]]
-#remDr$open()
-tic()
-remDr$navigate(url = "https://ezp.sub.su.se/login?url=https://orbis4.bvdinfo.com/ip")
-toc() # >6s
-## Note: working with VPN skips authentification steps! but takes longer to load
-Sys.sleep(30)
-
-## Authentication
-# me <- remDr$findElement("id", "username")
-# # me$clearElement()
-# me$sendKeysToElement(list("juro7132"))
-# pswd <- remDr$findElement("xpath", '//*[(@id = "password")]')
-# pswd$clearElement()
-# pswd$sendKeysToElement(list(keyring::key_get("sub", "juro7132"), key = "enter"))
-
-#### for loop should start here
-## create list to catch results
-# revenues <- list()
-# guo_list <- list()
-# shr_list <- list()
-
-## re-start the search: here just for quick restart
-remDr$goBack()
-# You need to do the search manually (2-3 loops) until it gives you the option
-# of restarting search again 
-# guos_missing comes from script 17
-owners <- guos_missing # only run for second batch, trying to recover companies missed on the first round
-
-# car_dat comes from script 17, additional companies from Carmine et al dataset
-owners <- car_owners
-
-# df_missing comes from google drive, they are orbis ids recovered manually by Bianca
-owners <- df_missing$OrbisID
-
-# df for mining: 29k
-owners <- mines
-
-remDr$refresh()
-
-for (i in 411:length(owners)){ #seq_along(dat$company)
-    tic()
-    ## Search one company:
-    srch <- remDr$findElement("id", "search")
-    # delete any left over from last iteration
[…]

### 2024-11-15 16-cleaning_orbis.R

@@ -1,668 +1,0 @@
-library(tidyverse)
-library(tictoc)
-library(patchwork)
-# load datasets
-
-
-load("data/boats_owners.Rda")
-load("data/marine_shareholders.RData")
-
-load("data/marine_shareholders_missing.RData")
-load("data/marine_shareholders_Carmine.RData")
-load("data/marine_shareholders_missing_completed.RData")
-load("data/carmine_shareholders.RData") # this is the missing values of carmine manually coded by Bianca
-df_boats
-
-owners <- df_boats |>
-    filter(!is.na(reg_owner)) |> 
-    pull(reg_owner) |> unique()
-
-guo_list |> length()
-revenues |> length()
-shr_list |> length()
-
-guo_list[[1]]
-revenues[[1]]
-shr_list[[1]]
-
-## First I need to clean the list object and coerce to df
-guo_list[[1]]
-
-guo_list[[1]][[1]]$X1[2] |> 
-    str_remove_all("\\n|\\t") |> 
-    #str_extract("Orbis ID: \\d{1,10}") |> 
-    str_extract("BvD ID: \\w{1,12}")
-
-
-#### Create functions ####
-extract_guo <- function(x){
-    
-    if(is.null(x))
-        return(out <- tibble(
-            orbis_id = NA,
-            BvD_id = NA,
-            orbis_name = NA,
-            location = NA,
-            guo = NA))
-    
-    else
-        orbis_id <- x[[1]]$X1[2] |> 
-            str_remove_all("\\n|\\t") |> 
-            str_extract("Orbis ID: \\d{1,10}")
-    
-    BvD_id <- x[[1]]$X1[2] |> 
-        str_remove_all("\\n|\\t") |> 
-        str_extract("BvD ID: \\w{1,12}")
-    
-    txt <- x[[1]]$X2[1] |> 
-        str_split(pattern = " \n\t\t\t\t\n\t\t\t\t") |> 
-        unlist()
-    
-    location <- txt[2]
-    orbis_name <- txt[1] |> str_remove("0\n\t\t\t\t")
-    
-    guo <- x[[1]]$X3[2]
-    
-    out <- tibble(
-        orbis_id = orbis_id,
-        BvD_id = BvD_id,
-        orbis_name = orbis_name,
-        location = location,
-        guo = guo)
-    
-    
-    return(out)
-}
-
-extract_guo(guo_list[[110]])
-
[…]

### 2024-11-15 17-alternative_names.R

@@ -1,61 +1,0 @@
-# many of the companies identified in Lloys do not match a company in Orbis
-# probably due to different spellings on both databases. Here we attempt to 
-# match companies with alternative names
-
-library(tidyverse)
-
-
-load("data/boats_owners.Rda")
-load( "data/cleaned_data_230708.Rda")
-
-guos
-owners
-revs
-
-# 000 means the company is dissolved
-guos_missing <- guos |> 
-    filter(is.na(orbis_id)) |> 
-    mutate(company = str_remove_all(
-        company, "S.A.|Company|Limited|LLC|H/f|AS|Incorporated|Inc")) |> 
-    mutate(company = str_trim(company, "both")) |> 
-    pull(company)
-
-### Now with gous_missing run 15-shareholders.R again
-
-
-#### Compare with Carmine's dataset ####
-car_dat <- read_csv(file = "~/Documents/Projects/high-seas-corporations/data/03_output/HS_fishing_vessels_with_corp_actors.csv")
-car_dat <- readxl::read_xlsx(path = "~/Documents/Projects/high-seas-corporations/data/02_processed/corp_info_HS_fishing_vessels.xlsx", sheet = 1)
-
-car_dat |> names()
-
-car_dat |> 
-    select(year, MMSI, best_flag, gfw_owner, owner:last_col())
-
-
-
-missing_mmsi <- df_boats |> 
-    filter(is.na(reg_owner)) |> 
-    pull(mmsi)
-
-(missing_mmsi %in% car_dat$MMSI) |> sum() # 2244
-(df_boats |> filter(!is.na(reg_owner)) |> pull(mmsi) %in% car_dat$MMSI) |> sum() #840
-
-# filter only boats that appear on our dataset of vessels fishing on risk areas
-car_dat <- car_dat |> 
-    filter(MMSI %in% missing_mmsi) |> 
-    select(MMSI, owner, gfw_owner)
-
-### Now with car_dat run 15-shareholders.R again
-car_owners <- car_dat |> 
-    select(owner) |> 
-    unique() |> 
-    filter(owner != 'n/a') |> 
-    mutate(owner = str_remove(owner, pattern = "\\(.*")) |> 
-    mutate(owner = str_remove(
-        owner, pattern = "LTD|LTDA|IMP|EXP|LLC|S\\.L\\.|YK$|\\.\\,|\\.\\, \\.|\\,$|\\, \\.$")) |>
[…]

### 2024-11-15 18-combine.R

@@ -1,9 +1,0 @@
-library(tidyverse)
-
-
-load("data/cleaned_data_231023.Rda")
-
-
-guos # 1556 obs
-revs # 965 obs
-shrs # 1294 obs

### 2024-11-15 99-orbis.R

@@ -1,214 +1,0 @@
-library(tidyverse)
-library(RSelenium)
-library(keyring)
-library(rvest)
-library(tictoc)
-
-
-## todo:
-## - implement the search in such a way that always picks up the first company (most likely result)
-## - check before going into shareholders if there are shareholders on the first result page: solved
-## - If not, create an error / message and record that as response: solved | it does not create an error message but return a df saying there is no companies.
-## - if multiple shareholders, loop through pages
-## - Create a loop or walking function that go through all of them (safely?)
-## 
-## - download the sardina-anchoveta data and tuna; for tuna use the open ocean project
-
-dat <- read_csv(file = "data/fishing_companies_alaskan-pollock.csv") # pollock
-dat <- read_csv(file = "data/fishing_comps_sardina-anchoveta.csv") # sardine-anchovies
-load("data/shr_class.RData")
-
-## Bolivian companies
-dat <- readxl::read_xlsx(path = "data/bolivian-soy-which-are-the-biggest-exporting-companies-in-2020.xlsx", sheet = 1) |> 
-    rename(company = exporter_group)
-dat <- dat |> 
-    # very useful reprex: all caracters from the begining until one finds -i
-    mutate(company = str_remove(company, pattern = "^.*?(?=- )") |> 
-               str_remove("- "))
-
-
-## Initialize remote driver
-d <- rsDriver(
-    port = 4445L, browser = "firefox", extraCapabilities = list(acceptInsecureCerts = TRUE)) # should open a chrome
-remDr <- d[["client"]]
-#remDr$open()
-tic()
-remDr$navigate(url = "https://ezp.sub.su.se/login?url=https://orbis4.bvdinfo.com/ip")
-toc() # >40s
-## Note: working with VPN skips authentification steps! but takes longer to load
-Sys.sleep(30)
-
-## Authentication
-# me <- remDr$findElement("id", "username")
-# # me$clearElement()
-# me$sendKeysToElement(list("juro7132"))
-# pswd <- remDr$findElement("xpath", '//*[(@id = "password")]')
-# pswd$clearElement()
-# pswd$sendKeysToElement(list(keyring::key_get("sub", "juro7132"), key = "enter"))
-
-#### for loop should start here
-## create list to catch results
-revenues <- list()
-guo_list <- list()
-shr_list <- list()
-
-## re-start the search: here just for quick restart
-remDr$goBack()
[…]

### 2024-11-15 99_noaa.R

@@ -1,25 +1,0 @@
-library(tidyverse)
-library(fs)
-library(here)
-
-
-## Not very informative
-dat <- read_csv(file = "/Users/juanrocha/Downloads/Nearshore Fish Atlas of Alaska_ID_17274.csv")
-dat |> 
-    names()
-
-dat <- read_csv("/Users/juanrocha/Downloads/goa2015_2019.csv")
-
-## from Alaskan pollock permits:
-## files ending in cp are catcher processors, endin gin cv are catcher vessels
-fls <- dir_ls("data/Alaska_pollock/")
-
-dat <- fls |> str_subset("afa") |> 
-    map(.f = function(x) read_csv(x, skip=1) |> janitor::clean_names()) |> 
-    map(select, owner_name) |> 
-    bind_rows() |> 
-    unique()
-dat
-
-read_csv("data/fishing_companies_alaskan-pollock.csv")
-

### 2024-11-15 download_sst.R

@@ -1,49 +1,0 @@
-library(tidyverse)
-library(tictoc)
-## Some of this code follows the steps of this tutorial to download NOAA SST data
-## https://robwschlegel.github.io/heatwaveR/articles/OISST_preparation.html
-## also available under the heatwaveR package
-
-
-# First we tell R where the data are on the interwebs
-OISST_base_url <- "https://www.ncei.noaa.gov/data/sea-surface-temperature-optimum-interpolation/v2.1/access/avhrr/"
-# Note that one may go to this URL in any web browser to manually inspect the files
-
-# Now we create a data.frame that contains all of the dates we want to download
-# NB: In order to change the dates download changes the dates in the following line
-OISST_dates <- tibble(t = seq(as.Date("1981-09-01"), as.Date("2022-12-31"), by = "day"))
-
-# To finish up this step we add some text to those dates so they match the OISST file names
-OISST_files <- OISST_dates %>% 
-    dplyr::mutate(t_day = base::gsub("-", "", t),
-                  t_month = base::substr(t_day, 1, 6),
-                  t_year = lubridate::year(t),
-                  file_name = base::paste0(OISST_base_url, t_month, "/", "oisst-avhrr-v02r01.", t_day ,".nc"))
-
-# This function will go about downloading each day of data as a NetCDF file
-# Note that this will download files into a 'DATA/OISST' folder in the root directory
-# If this folder does not exist it will create it
-# If it does not automatically create the folder it will need to be done manually
-# The folder that is created must be a new folder with no other files in it
-# A possible bug with netCDF files in R is they won't load correctly from 
-# existing folders with other file types in them
-# This function will also check if the file has been previously downloaded
-# If it has it will not download it again
-OISST_url_daily_dl <- function(target_URL){
-    base::dir.create("~/Documents/Projects/DATA/OISST", showWarnings = F)
-    file_name <- base::paste0("~/Documents/Projects/DATA/OISST/",base::sapply(base::strsplit(target_URL, split = "/"), "[[", 10))
-    if(!base::file.exists(file_name)) utils::download.file(url = target_URL, method = "libcurl", destfile = file_name)
-}
-
-# The more cores used, the faster the data may be downloaded
-# It is best practice to not use all of the cores on one's machine
-doParallel::registerDoParallel(cores = 10)
-
-# And with that we are clear for take off
-tic()
-base::system.time(plyr::l_ply(OISST_files$file_name, .fun = OISST_url_daily_dl, .parallel = T))
[…]

### 2024-11-15 heatwaves.R

@@ -1,169 +1,0 @@
-library(tidyverse)
-library(heatwaveR)
-library(fs)
-library(here)
-library(tidync)
-library(tictoc)
-library(doParallel)
-registerDoParallel(cores = 10)
-#library(data.table)
-
-#### Heatwaves on big ncdf file ####
-# J230301: `processing_sst.Rmd` was used to combine all .nc files from NOAA into
-# a unique file that can be queried. Instead of reading files multiple times one
-# reads the big one once and call slices at the time.
-# Following code from the package vignette also available at:
-# https://robwschlegel.github.io/heatwaveR/articles/Download_SST_v2.html
-
-nc_file <- dir_ls("data/") |> str_subset("combined.nc") # one file
-datadir <- "data/processed_MHW" # processed marine heatwaves
-
-## Function to read data (from tutorial)
-OISST_load <- function(file_name, lon1, lon2) {
-    OISST_dat <- tidync(file_name) %>%
-        hyper_filter(lon = between(lon, lon1, lon2)) %>%
-        hyper_tibble(select_var = "sst", force = TRUE, drop = TRUE) %>%
-        dplyr::select(-zlev) %>%
-        dplyr::rename(t = time, temp = sst) %>%
-        mutate(t = as.Date(t, origin = "1978-01-01"))
-    return(OISST_dat)
-    rm(OISST_dat)
-}
-
-# Event detection function (from tutorial)
-event_only <- function(df) {
-    # first calculate the climatologies
-    clim <- ts2clm(data = df, climatologyPeriod = c("1989-01-01", "2019-12-31"))
-    # then the events
-    event <- detect_event(data = clim, categories = TRUE, climatology = TRUE)
-    rm(clim)
-    # return only the event metric dataframe of results
-    return(event$event)
-    rm(event)
-}
-
-# Define the slices
-# 5° longitude slices seem to work fine on
-# my MacBook Pro with 32Gb RAM and 12 cores
-slice_df <- tibble(lon1 = seq(0, 355, 5),
-                   lon2 = seq(5, 360, 5))
-
-# test on one slide
-# tic()
-# sst <- OISST_load(nc_file, lon1 = slice_df$lon1[1], lon2 = slice_df$lon2[1]) |> 
-#     filter(t < "2020-02-01") ## added line to address problem below
-# toc() # 60s, including filter 74s
-
-## I get an error with the following:
-## Error in do.ply(i) :
[…]

### 2024-11-15 map_heatwaves.R

@@ -1,115 +1,0 @@
-library(tidyverse)
-library(fs)
-library(here)
-library(tictoc)
-
-
-fls <- dir_ls("data/processed_MHW/")
-
-## function to summarize each file
-mhw_summary <- function(file){
-    dat <- read_csv(file)
-    x <- dat |> 
-        group_by(lon, lat) |> 
-        summarize(
-            mean_intensity = mean(intensity_mean),
-            n_events = n(),
-            tot_days = sum(duration),
-            mp_moderate = mean(p_moderate, na.rm = TRUE),
-            mp_strong = mean(p_strong, na.rm = TRUE),
-            mp_severe = mean (p_severe, na.rm = TRUE),
-            mp_extreme = mean(p_extreme, na.rm = TRUE))
-    return(x)
-    rm(dat, x)
-}
-# create a function that slice only one year to check if results align with other papers
-mhw_year <- function(file, date1, date2){
-    dat <- read_csv(file) 
-    x <- dat |> 
-        filter(date_start >= date1, date_end <= date2) |> 
-        group_by(lon, lat, category) |>
-        add_column(n = 1) |> 
-        summarize(tot = sum(n)) |> 
-        pivot_wider(names_from = category, values_from = tot)
-    return(x)
-    rm(dat, x)
-}
-
-mhw_categories <- function(file){
-    dat <- read_csv(file)
-    x <- dat |> 
-        group_by(lon, lat, category) |>
-        add_column(n = 1) |> 
-        summarize(tot = sum(n)) |> 
-        pivot_wider(names_from = category, values_from = tot)
-    return(x)
-    rm(dat, x)
-}
-# create a function that extracts the proportion of the events above / below the 
-# categories thresholds.
-dat <- read_csv(fls[1])
-
-
-tic()
-out <- map(fls, mhw_summary) #, date1 = "2016-01-01", date2 = "2016-12-31"
-toc() # 1.37s
-
-out <- out |>
-    bind_rows()
-# 
-out |>
-    pivot_longer(cols = starts_with("I"), names_to = "category", values_to = "events" ) |>
-    group_by(lon,lat, category) |>
-    summarize(total_events = sum(events, na.rm = TRUE)) |> 
-    #filter(category != "I Moderate") |>
-    ggplot(aes(lon, lat)) +
-    geom_tile(aes(fill = total_events)) +
[…]

### 2024-11-15 refs.R

@@ -1,0 +1,10 @@
+library(tidyverse)
+
+txt <- read_lines(file = "paper/manuscript.Rmd")
+
+txt |> str_subset("\\[*\\]")
+
+refs
+
+
+system("curl -LH 'Accept: application/x-bibtex' https://www.doi.org/10.1126/science.1156401")

### 2024-11-15 Notes.Rmd
update notes and gitignore

@@
 # Notes
+
+**241112**
+
+- MHW did not have strong clear effects on single species models (some sp decline while other thrive).
+But sp can move, primary producers don't move.
+MHW so far have been contained to limited areas, but projections suggest larger scale possibly longer than the ability of fish to move away.
+-
+
+**sometime**
+
+- Make sure to read for the paper:
+    - Malin Pinsky paper on MHW [X]
+    - Latest Kroodsma paper on fishing fleets []
+
+**240321**
+
+- The lists of owners have numbers because at some point we decided to search for OrbisIDs instead of company names.
+The company names results were ambiguous and Bianca did some manual work to match ambiguous names to IDs that were of companies fishing.
+That is why we have some owners coded as orbis ID instad of compnay names.
+- `guos` is not unique because the same company with the same `orbis_id` can have multiple GUOs over time. We do not distinguish them.
 
